### 2013-01-11 DGE_paper.tex
initial commit
added two tables

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Genome-wide transcriptome variation in chicken line 6 and 7 in response to Marek's disease infection}
}
% Insert Author names, affiliations and corresponding author email.
\\
Author1$^{1}$, 
Author2$^{2}$, 
Author3$^{3,\ast}$
\\
\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Gene models summary}
The number of genes and isoforms is shown in Table \ref{tab:gene_models}.

\subsection*{Differential gene expression}
The number of differentially expressed genes is shown in Table \ref{tab:dge}.

\subsection*{Differential exon usage}

\subsection*{Polymorphisms in differentially expressed genes/exons}

\subsection*{Functional analysis}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{template}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}

\begin{table}[!ht]
\caption{
\bf{Gene models summary}}
\begin{tabular}{ccc}
\hline
Method & Gene & Isoform \\
\hline
Assembly & 25,290 & 54,044 \\
Cufflinks & 21,345 & 36,218 \\
Combined & 24,980 & 46,613 \\
\hline
\end{tabular}
\begin{flushleft}Number of genes and isoforms from gene models
built from \emph{de novo} assembly and Cufflinks.
Combination of both methods decreased the number of genes
and isoforms by merging fragmented transcripts to
form more complete gene models.
\end{flushleft}
\label{tab:gene_models}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Differentially expressed genes}}
\begin{tabular}{cccccc}
\hline
Sample & Up-regulated & Down-regulated & Total & Refseq & Ensembl\\
\hline
Line 6 & 1,043 & 595 & 1,638 & 984 (60.0\%) & 1,252 (76.4\%) \\
Line 7 & 1,976 & 1,124 & 3,100 & 2,114 (68.2\%) & 2,574 (83.0\%) \\
Both & 681 & 236 & 917 & 664 (72.4) & 771 (84.1\%) \\
\hline
\end{tabular}
\begin{flushleft}
FDR $< 0.1$
\end{flushleft}
\label{tab:dge}
\end{table}
\end{document}

### 2014-01-03 DGE_paper.tex
update DGE_paper.tex

@@
 {\Large
-\textbf{Genome-wide transcriptome variation in chicken line 6 and 7 in response to Marek's disease infection}
+\textbf{Transcriptome variation in response to Marek's disease virus early infection}
 }
@@
 \\
-Author1$^{1}$, 
-Author2$^{2}$, 
-Author3$^{3,\ast}$
+Likit Preeyanon$^{1}$
+C. Titus Brown$^{1,2}$
+Hans H. Cheng$^{3,\ast}$
 \\
-\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
+\bf{1} Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI, USA
 \\
-\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
+\bf{2} Computer Science and Engineering, Michigan State University, East Lansing, MI, USA
 \\
-\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
+\bf{3} USDA, ARS, Avian Disease and Oncology Laboratory, East Lansing, MI, USA
 \\
-$\ast$ E-mail: Corresponding author@institute.edu
+$\ast$ E-mail: Corresponding hans.cheng@ars.usda.gov
 \end{flushleft}
@@
 \section*{Abstract}
+Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
+Different MHC alleles have been associated with susceptibility and resistance to the disease.
+However, chickens line 6 and 7 show significant phenotypic differences when chanllenged with
+MDV even with the same MHC allele. Therefore, the major unanswered question is what genetic factors
+underlie the susceptibility and resistance of line 6 and 7.
+In this study, we identified differentially expressed genes and isoforms in chicken line 6 and 7
+in response to Marek's disease infection. Results from pathway analysis and functional analysis
+show that changes in expression of genes and isoforms may contribute to resistance and
+susceptibility of the disease. We also identified many single nucleotide polymorphisms (SNPs) between
+line 6 and 7 that may cause alteration of alternative splicing, which in turn results in allele
+specific expression.
 
@@
 
+Marek's disease is an economically significant chicken disease that affects a poultry industry worldwide.
+Total of \$2 billion loss has been reported from outbreaks.
+The disease is caused by highly oncogenic Marek's disease virus (MDV) which causes T-cell lymphoma.
+Vaccination is effective in reducing incidence of tumor formation; however, current MD vaccines do not
+prevent infection or horizontal spread of the virus.
+Improper use of vaccines is a key factor that drives evolution of highly virulent strains in vaccinated
+flocks.
+***Need to add intro about breed selection against infection***
+
+Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
+the disease. For example, chickens with MHC allele B$^{21}$ is highly resistant in contrast to chickens with
+B$^{16}$ allele.
+However, chickens line 6 and 7 , even with common allele (B$^6$), exhibit different phenotypic response to
+MD.
+The major unanswered question is what genetic factors contribute to susceptibility and resistance of
+the disease.
+***Need to add intro about gene and isoforms in immune system***
+In this study, we identified several annotated and unannotated genes and isoforms that are differentially
+expressed in response to MD infection. Pathway analysis and gene ontology suggest that these
+genes may have important roles in immune response to MD.
+
 % Results and Discussion can be combined.
@@
 
-\subsection*{Gene models summary}
+\subsection*{Custom Gene Annotations}
+A large number of reads mapped to intergenic and intronic regions (table 1, supp.)
+regions of Ensembl annotations (Galgal4.72)
+indicates that many genes and isoforms are not included in the annotations.
+Therefore, to extensively study gene and isoform expression, we employed \emph{de novo}
+transcriptome assembly as well as genome-based assembly to build custom gene annotations (see methods).
 The number of genes and isoforms is shown in Table \ref{tab:gene_models}.
+Genes were identified by aligning against all vertebrate transcripts in GenBank database.
+We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
+as well as genes found in other organisms (Table \ref{}).
 
-\subsection*{Differential gene expression}
-The number of differentially expressed genes is shown in Table \ref{tab:dge}.
+\subsection*{Differential Gene Expression}
+Many previous studies have reported many differentially expressed (DE) genes between line 6 and line 7
+from Microarray and RNA-Seq experiments from different stages of infection.
+Some known differentially expressed genes were identified in this study.
+For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.5 fold in susceptible
+chicken with MHC allele B\textsuperscript{19}\cite{}. It was also found to be down-regulated
+about three fold in line 7.
 
-\subsection*{Differential exon usage}
+Even though both lines shared a significant number of genes that were differentially expressed,
+approximately, twice as many genes in line 7 as in line 6 were differentially expressed (Table \ref{tab:dge}).
+Interestingly, some genes that were differentially expressed in both lines were
+regulated in opposite directions.
+As shown in table \ref{tab:opposite}, LYG2, SFTPA1, LL and SERPINB10 were up-regulated in line 7 but down-regulated in line 6.
+In contrast, TAF1D and LOC424145 were up-regulated in line 6 but down-regulated in line 7.
+LYG2, for instance, was up-regulated by 2.57 folds in line 7 but down-regulated by 1.7 folds in line 6.
+This gene is involved in a defense against bacteria.
 
+Pathway analysis shows that genes in line 7 are significantly enriched (p $< 0.05$) in pathways involved in
+immune system such as Toll-like receptor signaling, intestinal immune network for IgA production,
+and Phagosome pathway, which are not significantly enriched in line 6 (fig. \ref{line67_kegg}).
+Figure \ref{KEGG_phagosome} and \ref{KEGG_IgA} shows a diagram of phagosome and intestinal immune network
+for IgA production repectively. Genes from line 6 and line 7 are highlighted with yellow and red color.
+Common genes are highlighted with orange color.
+These two pathways show that both line 6 and line 7 regulated MHC class I gene in response to infection,
+but only line 7 regulated MHC class II gene.
+It has been shown that Marek's disease virus (MDV), unlike other herpes viruses, up-regulates MHC class II gene
+during infection. Moreover, regulation is believed to be mediated directly by MDV infection\cite{Niikura:2007fq}.
+However, from these results, it appears that only MHC class II gene in line 7 was up-regulated.
+
+Furthermore, differentially expressed genes in line 7 are enriched in biological processes involved in
+both adaptive and innate immune responses, whereas
+differentially expressed genes in line 6 are only enriched in innate immune responses (fig. \ref{line67_go}).
+This might correspond to the fact that MHC class II gene was upregulated in line 7.
+High level of MHC class II molecules may help recruit more T cells, which in turn enhance activation of
+the adaptive immune response.
+
+\subsection*{Differential Exon Usage}
+
+\subsection*{Functional Analysis}
 \subsection*{Polymorphisms in differentially expressed genes/exons}
-
-\subsection*{Functional analysis}
 
@@
 \section*{Materials and Methods}
+\subsection{Gene Models Construction}
+With the lack of high quality gene models for chickens. We employed two different approaches to
+construct gene models from RNA-Seq reads.
+First, short reads were quality trimmed with conditri/2.1\cite{} \texttt{(-cutfirst 10)}
+and assembled using Velvet/1.21\cite{} and Oases/0.2.06\cite{} to obtain long transcripts.
+Assembly was done with k value ranges from 21 to 31 for both local and global assembly
+(described in Gimme paper\cite{}).
+Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean\cite{}.
+All transcripts were then aligned to chicken reference genome (galGal4, with short
+contigs removed) with BLAT\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
+Filtered alignments from BLAT were then used to produced gene models using Gimme.
+
+Second, short reads were aligned to reference genome using Tophat/2.0.5\cite{} and mapping output were used to
+build gene models by Cufflinks2\cite{} with default parameters.
+Both set of gene models were combined using Gimme to build more comprehensive models
+to be used in this study.
+
+\subsection{Differential Gene Expression and Gene Ontology}
+
+To identify DE genes, we obtained read counts using multiBamCov command from BEDTools/12.13.1\cite{}
+for all datasets.
+Then DE gene were identified by DESeq/1.10.1\cite{S:2010fu} from Bioconductor.
+Data from single- and paired-end datasets from the same line were treated as biological replicates.
+To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq/1.10.0\cite{}.
+P-values were corrected by Benjamini-Hochberg multiple testing correction.
+Genes, patways and GO terms with corrected P-value $<0.05$ were considered significant.
+
+\subsection{Differential Splicing Event}
+Gene models were converted to alternative splicing models using a Python script.
+In order to increase sensitivity, read counts from single- and paired-end samples were combined and treated
+as single-end reads for splicing event analysis with MISO/0.4.9\cite{Katz:2010iv}.
+Splicing events with Bayes factor $>10$ and $\Delta\Psi>0.20$ were considered significant.
+Read coverages and $\Psi$ distributions were plotted using Sashimi plot\cite{Katz:2013vx}.
+
+\subsection{Variant Calling and \emph{In Silico} Splicing Analysis}
+Variants were called using mpileup command from SAMTools/0.1.18\cite{} and BCFTools\cite{}.
+Only variants with quality score $\ge20$ were used for mutation analyses.
+Exon enhancers and suppressors were predicted using Human Splicing Enhancer web portal\cite{}.
+Human default parameter settings were used in all analyses.
+The following regulatory sequences were used to determined the effect of variants:
+HSF integrated matrices for serine/arginine-rich proteins (SRp40, SC35, SF2/ASF, SF2/ASF,
+IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE), RESCUE-ESE hexamer (RESCUE-ESE),
+putative 8-mer ESEs (PESEs) and putative 8-mers exonic splicing silencers (PESSs),
+exon-identity and intron-identiry elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding
+motifs, and Fas exonic splicing silencers.
 
@@
 
-%\section*{References}
+% \section*{References}
 % The bibtex filename
-\bibliography{template}
+\bibliography{ref}{}
 
 \section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=7in]{line67_KEGG_cleveland.pdf}
+\end{center}
+\caption{
+{\bf Enriched KEGG pathways.} Rest of figure 2  caption.  Caption 
+should be left justified, as specified by the options to the caption 
+package.
+}
+\label{line67_kegg}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=7in]{line67_GOseq_GO_cleveland.pdf}
+\end{center}
+\caption{
+{\bf Enriched Gene Ontology (BP) Terms.} Rest of figure 2  caption.  Caption 
+should be left justified, as specified by the options to the caption 
+package.
+}
+\label{line67_go}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=6in]{line67_KEGG_gga04145.png}
+\end{center}
+\caption{
+{\bf Enriched KEGG pathways in line 7.}  Rest of figure 2  caption.  Caption 
+should be left justified, as specified by the options to the caption 
+package.
+}
+\label{KEGG_phagosome}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=6in]{line67_KEGG_gga04672.png}
+\end{center}
+\caption{
+{\bf Intestinal network for IgA production pathway.}  Rest of figure 2  caption.  Caption 
+should be left justified, as specified by the options to the caption 
+package.
+}
+\label{KEGG_IgA}
+\end{figure}
 
@@
 \hline
-Line 6 & 1,043 & 595 & 1,638 & 984 (60.0\%) & 1,252 (76.4\%) \\
+Line 6 & 1,043 & 595 & 1,188 & 984 (60.0\%) & 1,252 (76.4\%) \\
 Line 7 & 1,976 & 1,124 & 3,100 & 2,114 (68.2\%) & 2,574 (83.0\%) \\
@@
 \end{table}
+
+\begin{table}[!ht]
+\caption{
+\bf{Genes regulated in opposite direction}}
+\begin{tabular}{cccccc}
+\hline
+Gene & Entrez ID & Fold Change & \\
+ & & Line 6 & line 7 \\
+\hline
+TAF1D & 419002 & +1.08 & -1.13 \\
+LYG2 & 395708 & -1.70 & +2.57 \\
+SFTPA1 & 395308 & -5.12 & +3.51 \\
+LOC424145 & 424145 & +1.57 & -1.68 \\
+LL & 423630 & -3.30 & +8.31 \\
+SERPINB10 & 395715 & -1.12 & +0.87 \\
+\hline
+\end{tabular}
+\begin{flushleft}
+    (-) down-regulated, (+) up-regulated
+\end{flushleft}
+\label{tab:opposite}
+\end{table}
 \end{document}

### 2014-01-27 DGE_paper.tex
minor update

@@
 flocks.
+
 ***Need to add intro about breed selection against infection***
@@
 the disease.
+
 ***Need to add intro about gene and isoforms in immune system***
+
 In this study, we identified several annotated and unannotated genes and isoforms that are differentially
@@
 indicates that many genes and isoforms are not included in the annotations.
-Therefore, to extensively study gene and isoform expression, we employed \emph{de novo}
-transcriptome assembly as well as genome-based assembly to build custom gene annotations (see methods).
-The number of genes and isoforms is shown in Table \ref{tab:gene_models}.
-Genes were identified by aligning against all vertebrate transcripts in GenBank database.
+To extensively study gene and isoform expression, we employed \emph{de novo}
+transcriptome assembly and a reference-guided assembly to build custom gene annotations (see methods).
+The number of genes and isoforms is summarized in Table \ref{tab:gene_models}.
+Genes were identified by aligning against non-redundant protein database from NCBI.
 We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
@@
 
+\subsection*{Gene network analysis}
 Pathway analysis shows that genes in line 7 are significantly enriched (p $< 0.05$) in pathways involved in
@@
 \subsection{Gene Models Construction}
-With the lack of high quality gene models for chickens. We employed two different approaches to
+Due to lack of complete gene models for chickens, we employed two methods to
 construct gene models from RNA-Seq reads.
@@
 Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean\cite{}.
-All transcripts were then aligned to chicken reference genome (galGal4, with short
-contigs removed) with BLAT\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
-Filtered alignments from BLAT were then used to produced gene models using Gimme.
+All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
+chromosomes removed) with BLAT\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
+Filtered alignments from BLAT were then used to produced gene models using Gimme\cite{}.
 
-Second, short reads were aligned to reference genome using Tophat/2.0.5\cite{} and mapping output were used to
-build gene models by Cufflinks2\cite{} with default parameters.
-Both set of gene models were combined using Gimme to build more comprehensive models
-to be used in this study.
+Second, short reads were aligned to reference genome using Tophat/2.0.5\cite{} and 
+gene models were built by Cufflinks2\cite{} with default parameters.
+A combination of gene models from both set of gene models were used in this study.
 

### 2014-06-01 paper.tex
updated

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Transcriptome variation in response to Marek's disease virus early infection}
}
% Insert Author names, affiliations and corresponding author email.
\\
Likit Preeyanon$^{1}$
C. Titus Brown$^{1,2}$
Hans H. Cheng$^{3,\ast}$
\\
\bf{1} Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI, USA
\\
\bf{2} Computer Science and Engineering, Michigan State University, East Lansing, MI, USA
\\
\bf{3} USDA, ARS, Avian Disease and Oncology Laboratory, East Lansing, MI, USA
\\
$\ast$ E-mail: Corresponding hans.cheng@ars.usda.gov
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
Different MHC alleles have been associated with susceptibility and resistance to the disease.
However, chickens line 6 and 7 show significant phenotypic differences when chanllenged with
MDV even with the same MHC allele. Therefore, the major unanswered question is what genetic factors
underlie the susceptibility and resistance of line 6 and 7.
In this study, we identified differentially expressed genes and isoforms in chicken line 6 and 7
in response to Marek's disease infection. Results from pathway analysis and functional analysis
show that changes in expression of genes and isoforms may contribute to resistance and
susceptibility of the disease. We also identified many single nucleotide polymorphisms (SNPs) between
line 6 and 7 that may cause alteration of alternative splicing, which in turn results in allele
specific expression.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

Marek's disease is an economically significant chicken disease that affects a poultry industry worldwide.
Total of \$2 billion loss has been reported from outbreaks~\cite{}.
The disease is caused by highly oncogenic Marek's disease virus (MDV) which causes T-cell lymphoma.
Vaccination is effective in reducing incidence of tumor formation; however, current MD vaccines do not
prevent infection or horizontal spread of the virus.
Improper use of vaccines is a key factor that drives evolution of highly virulent strains in vaccinated
flocks.

***Need to add intro about breed selection against infection***

Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
the disease. For example, chickens with MHC allele B$^{21}$ is highly resistant in contrast to chickens with
B$^{16}$ allele.
However, chickens line 6 and 7 , even with common allele (B$^6$), exhibit different phenotypic response to
MD.
The major unanswered question is what genetic factors contribute to susceptibility and resistance of
the disease.

Although gene expression has been shown to have a high correlation with phenotypes in many studies~\cite{}.
Studies have shown that alternative splice forms play a significant role in many biological events and
isoform expression levels can provide better signatures for some diseases~\cite{zhang2013isoform}.
Regulation of alternative splicing by enhancers (ESE) and silencers (ESS) and their corresponding
binding motifs have been identified in some organisms including human.
Abberation of those binding sites have been shown to be associated with increasing susceptibility and severity of diseases~\cite{}.
We hypothesized that SNPs between line 6 and 7, which are located in exons, are responsible for differential
expression of some isoforms between chicken line 6 and line 7.
To investigate those SNPs, we compared exon sequences of genes identified to have differential exon usage
from both lines and then predicted their effects on binding motifs.

% Results and Discussion can be combined.
\section*{Results and discussion}

\subsection*{Custom Gene Annotations}
A large number of reads mapped to intergenic and intronic regions (table 1, supp.)
regions of Ensembl annotations release 73.
indicates that many genes and isoforms are not included in the annotations.
To extensively study gene and isoform expression, we employed \textit{de novo}
transcriptome assembly and a reference-guided assembly method to build
custom gene annotations (see methods).
The number of genes and isoforms is summarized in Table~\ref{tab:gene_models}.
Genes were identified by aligning against non-redundant protein database from NCBI.
We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
as well as genes found in other organisms (Table~\ref{}).

\subsection*{Differential gene expression in response to MDV infection}

Many genes were found to be differentially expressed (DE) between control and infected
chicks in line 6 and line 7.
While the number of unique downregulated genes in both lines are approximately equal,
the number of unique upregulated genes in line 7 is much greater than in line 6 as shown
in figure~\ref{degenes_venn}.

Some DE genes have been reported from previous Microarray studies to be differentially expressed
in response to MDV infection.
For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
chicken with MHC allele B\textsuperscript{19} at 4 d.p.i.
It was also found to be down-regulated about 3-fold in line 7 (susceptible) in this study.
Similary, \textit{GMZA} gene, which has been reported to be upregulated across genetically different chickens,
was also found to be highly upregulated.
In contrast, some genes that have been reported to be highly expressed in resistant chickens were
downregulated in both lines. Those genes are \textit{AMIGO2, MMP13} and \textit{CLEC3B}, which were
found to be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
Several genes involved in innate immune response were differentially expressed in resistant chickens:
\textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5 was upregulated.
Other immune genes reported to be highly expressed in susceptible chickens including
\textit{AVD, ART1, NOS2, CXCL13L2, MX1} and \textit{SOCS1}~\cite{kaiser2003differential},
were found to be highly upregulated in both lines.
However, some discrepancies may be due to the difference in sensitivity between
Microarray and RNA-Seq methods.

Important cytokines such as \textit{IFN$\gamma$} and \textit{IFN$\beta$} were found to be highly
upregulated in infected chickens in both line 6 and 7 as well as \textit{INF$\alpha$3} (table~\ref{}).
However, expressions of genes encoding their corresponding receptors
were not different in line 6, but upregulated in line 7.
Keiser et al reported that expressions of \textit{IL6} and \textit{IL18} were different between
line 6 and 7.
Both genes were only upregulated in line 7 in early stage of infection (3-5 d.p.i).
At 10 d.p.i, the expression of \textit{IL6} was different from the control chickens in both lines;
whereas, the level of \textit{IL18} was only different in susceptible chickens.
Our results show the same pattern of expression of both cytokines.

In contrary, \textit{IL15} was reported no differences in expression between lines in the same study.
However, in this study, it was only upregulated in line 7 compared to uninfected control chickens.
Expression of \textit{IL15} is induced by \textit{TLR9}, which binds to non-methylated CpG
residues present in genomes of many DNA viruses, including herpes simplex virus.
This cytokine auto-regulates the expression of \textit{CD40}, which is
a transmembrane receptor required for activation of macrophages by CD4 T cells.
As a result, \textit{CD40} was only upregulated in line 7.

Interestingly, some genes that were differentially expressed in both lines were
regulated in the opposite direction (table~\ref{tab:opposite}).
Among genes downregulated in line 6 but upregulated in line 7,
\textit{LL} (lung lectin) and \textit{SFTPA1}, which encode lectin proteins, are part of
lectin-complement pathway.
\textit{LIMS1} is involved in cell differentiation and proliferation and
\textit{PPARG} is a suppressor of \textit{NFK$\beta$}-mediated proinflamatory response.
On the other hand, nearly all genes upregulated in line 6 but downregulated in line 7
are involved in cell survival such as mRNAs splicing, cell growth, and protein systhesis,
except CD7 whose function is involved in T cell-B cell interaction.
This difference suggests that even at this stage of infection in line 6, the lytic phase 
could be controlled or subsided because the viruses are driven to undergo latent infection early.
Therefore, genes important for recovery of cell damages are upregulated.
In comparison, the lytic phase in line 7 may still continue and as a result, genes involved in
immune responses are still upregulated.

\subsection*{Functional analysis of differntial-expressed genes}

To determine pathways that were perturbed during the infection, data were
analysed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{}.
Significantly perturbed pathways (adjusted p-value $< 0.1$) from both lines that involved in
immune response include Toll-like receptor signaling, cytokine-cytokine receptor interaction,
intestinal immune network for IgA production and cell adhesion molecules (CAMs).
Some other pathways important in response to viral infection and only significantly
enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway pathway,
NOD-like receptor signaling pathway and lysosome.
Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed genes in phagosome pathway.
Although MHC I gene (\textit{BF1}) was differentially expressed in both lines, other genes involved in
expressing newly synthesized MHC I were differentially expressed in line 7.
This suggests that at the stage of infection, new MHC I molecules were actively produced and expressed to
present antigen.
Furthermore, gene ontology analysis of biological processes (GO:BP) shows that categories involved in
both adaptive and innate immune responses were enriched in line 7.
On the other hand, only categories involved in innate immune responses were enriched in line 6.
In addition, the apoptosis pathway was significantly enriched in line 7, which suggests that the programmed
cell death could be induced by CTL response to eliminate the viruses.

At this stage of infection, our results suggest that lytic infection of MDV stimulated
both innate and the adaptive immune response, which led to activation of T cells in line 7.
Only activated T cells are infected by MDV, therefore, the lytic phase could facilitate the spread of
the viruses by enhancing expansion of activated T cells.
In contrast, the immune response in line 6 could have been able to drive the viruses to latency phase early
without eliciting adaptive immune response.
Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell contact between B cells and
T cells during antigen presentation or B cells activation by helper T cells.
Therefore, it is beneficial for the host to restrain such contact.
Limiting of T cells activation in line 6 could attribute to several mechanisms.
The innate immune response could be highly effective and could activate a strong adaptive immune response that
rapidly controls viral replication and forces the viruses into latency phase
Alternatively, the innate immune resposne itself could be highly effective in limiting viral replication.
Conversely, the innate immune response in line 6 could be inefficient in eliciting an adaptive immune
response. Since MDVs cannot replicate in APCs~\cite{}, the replication of MDVs is naturally limited.

\subsection*{Differential exon usage (DEU)}

Immune system is isoform-rich and many genes express different isoforms with distinctive functions in response to
stimuli such as stress, chemicals and infection.
Changes in expression of splice forms of immune related genes have been reported to be associated with increase
susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
Studying differential isoform expression could therefore shed lights into an inherent different between 
line 6 and 7 in that confers resistance and susceptibility to MD.
Microarray technology has been used to study isoform expression in several studies, but
its sensitiviy for detection of structurally similar isoforms is low and known or predicted annotations
are required~\cite{kane2000assessment}.
Although RNA-Seq method can provide reliable estimate of exon expressions compared to Microarray~\cite{pan2008deep}
and is not constrained to those limitations, studying isoform
expression using RNA-Seq is still not straightforward because of short read lengths.
Reads from current Illumina technology are generally not long enough to span across all exons in an isoform.
In most cases, only exons in close proximity are covered by the same read, which makes it
difficult to accurately predict a full structure of the isoform.
In addition, some genes are fused due to overlapping untranslated regions (UTRs),
which can also result in erroneous predicted isoform structures.

Due to those issues, it is not feasible to accurately estimate exression of isoforms, especially when
gene annotation is constructed from \textit{de novo} assembly~\cite{trapnell2013differential}.
To avoid these issues, we chose to study exon expression instead of isoform expression.
Using MISO with exon-centric method, only reads spanning across a few exons are used and only exons
involved in the splicing event are examined.
The expression of exon inclusion is calculated as percent spliced in (Psi or $\Psi$), which can be used to
infer the portion of transcripts that include the exon in each sample~\cite{Katz:2010iv}.
In this study, we investigated the three most common alternative splicing events in vertebrates, which are skipped
exons (SE), an alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.

\subsubsection*{DEU genes in response to MDV}

List of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
in infected chickens is shown in table~\ref{tab:line67i_diff_line67u}.
Genes can be categorized roughly into four groups based on the pattern of $\Psi$ across control and
infected birds in both lines.

Group I includes genes that $\Psi$s were up or down regulated in infected chickens in line 6 only.
This group include \textit{BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
associated C2H2-type zinc finger protein encoding gene, which functions as a tumour-suppressor
in T-cell lymphoma in human.
According to homologous alignments on UCSC genome browser, a splice form with the skipped exon
is similar to mouse \textit{BCL11B isoform b}.
The skipped exon was expressed 30\% in infected line 6 chickens; whereas it was rarely expressed
(4-7\%) in the control line 6 and both groups in line 7.
The skipped exon was not found to encode any known protein domain, however, it is in the middle of
two adjacent C2H2-type finger protein domains.
Another gene related to B and T cell lymphoma is \textit{SIK2} (salt-inducible kinase 2).
This gene has been reported to have a negative effect on T cell lymphoma by limiting the transcription of
HTLV-1 virus~\cite{tang2013lkb1}.
Some other genes are important in pre-mRNA splicing including \textit{TRA2A}, \textit{SRSF6} and \textit{GEMIN6}.
\textit{TRA2A} (transformer-2 protein homolog alpha) encodes RNA recognition motif (RRM),
the skipped exon is not found to encode any known protein domain.
\textit{SRSF6} (SR splicing factor 6) encodes a nuclear protein that belongs to the splicing factor protein family.
\textit{GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
These genes could play a significant role in regulating inclusion of alternative exons.
A gene in this group that might be important for innate immune responses is \textit{RAC3} (Ras-related C3
botulinum toxin substrate 3).
This gene encodes small GTPases, belonging to Ras family, that regulates a wide variety of cellular events
including cell growth, cytoskeletal reorganization and the activation of protein kinases.
The role of small GTPases in immune responses is discussed further below.

In group II, $\Psi$ values were relatively stable in control and infected chickens in line 6 and line 7.
Genes that could play an important role in immune responses are \textit{ITGB2} and \textit{HCK}.
\textit{ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of \textit{LFA-1} and \textit{CR3} receptors.
\textit{LFA-1} plays an important role in adhesion of lymphocytes with other cells.
\textit{CR3} binds to a vast array of ligands and molecules including complement C3bi, microbial proteins,
ICAM-1 and -2, ECM proteins and coagulation proteins~\cite{}.
It plays a significant role in neutrophils and monocytes activation including phagocytosis, adhesion and
migration~\cite{}.
Mutations in CD18 gene has been associated with type 1 leukocyte adhesion deficiency (LAD-1),
an autosomal-recessive inherited disease found in few families.
The disease is characterized by impair of lymphocytes in adherent-dependent functions, lack of accumulation
to the site of infection and recurrent bacterial and fungal infection~\cite{}.
\textit{HCK} transmits signal from cell surface receptors such as \textit{FCGR1A, FCGR2A, IL2, IL6,
IL18} and integrins (\textit{ITGB1, ITGB2}).
Based on our gene model, the skipped exon of \textit{ITGB2} is located in the 5$\prime$ UTR;
whereas, the skipped exon of \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).
Some other genes in this group are involved in cell rearrangement and cytokinesis that could be
regulated by signals from integrins including \textit{DYLNT1,DYNLL2, SEPT11, PFN2} and \textit{ZDHHC7}.
The role of \textit{ITGB2} in innate immune responses will be discuss further in the next section.

Group III includes genes that exhibit differential isoform expression only in infected chickens in line 7.
A number of genes in this group encode proteins that are parts of spliceosome: \textit{SRSF3}, \textit{HNRNPDL},
\textit{SFSWAP}, \textit{THOC1}, \textit{RNPC3} and \textit{SRSF5}.
Two genes involved in cell-cell contact regulated by integrins are also in this group.
\textit{PPP1R12A} is a mysosis phosphatase that regulates the interaction of actin and myosin downstream of GTPase
Rho proteins.
\textit{PODXL} encodes protein that functions in integrin-dependent manner to increase cell-cell contact.
Connections of these two genes and integrins will be discussed in the next section.

The last group only has one gene, \textit{GOSR1}.
The $\Psi$ value of this gene was less than 0.20 cutoff in control and infected chickens in line 6, but it is
significantly different between infected chickens in line 6 and 7.
However, there is a significant difference between control and infected chickens in line 7.
This gene encodes traffiking membrane protein important for transporting proteins from \textit{cis-} to \textit{trans-}
golgi network.

\subsubsection*{Integrins and actin cytoskeleton pathway}

Integrins are a special kind of receptors that transmit signals bidirectionally across cell membrane.
They are heterodimeric composed of an $\alpha$ (large) and a $\beta$ (small) subunits.
$\beta_{2}$ (CD18) subunit is exclusively expressed on lymphocytes and APCs as a component of \textit{LFA-1}
and \textit{CR3} receptors.
Upon binding, \textit{LFA-1} and \textit{CR3}.
As discussed above, \textit{RAC3, HCK, DYLNT1, DYNLL2, SEPT11, PFN2, ZDHHC7, PODXL} and \textit{PPP1R12A}
could be activated by the integrins because of their roles in cell rearrangement.
Interestingly, we found that in human, \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A} are in the regulation
of actin cytoskeleton pathway (figure~\ref{actin_pathway}).
Moreover, at least two of these genes are co-present in other three pathways (table~\ref{tab:integrin}) that
involved in immune responses.
This strongly suggests that these four genes could contribute to inherent difference in immune response between
line 6 and 7 even though their expressions were not significantly different between control and infected groups.

\subsubsection*{Splicing factors}

The second group includes genes encode protein compoments of spliceosome or are involved in pre-mRNA splicing.
Those genes are \textit{TRA2A, SRSF6, GEMIN6, SRSF3, HNRNPDL, SFSWAP, THOC1, THOC2, RNPC3} and \textit{SRSF5}.
Even though alternative splicing is regulated by splicesome, it is difficult to elucidate the effect of alternative
splice forms of spliceosome subunits on alternative splicing regulation.
This is in part due to the fact that the spliceosome is the most complex protein structures composed of many different
subunits each of which could have isoforms differing in function or activity~\cite{}.

\subsubsection*{Protein domains}
To understand the role of the alternative splice forms on the function of these genes, transcripts were
translated to protein sequences by ESTscan~\cite{}.
Protein sequences were then searched for annotated protein domains using InterProSearch~\cite{}.
Besides \textit{ITGB2}, other genes have alternative exons located in coding regions and could potentially affect
functional protein domains in some ways.
The exon with alternative 3$\prime$ splice site of \textit{RAC3} encode part of a protein domain identified as small GTPase of Ras subfamily
(ProSiteProfiles:PS5142 and SMART:SM00173).
Rac3 is highly homologous to Rac1 and has been reported to possess the ability of promoting membrane ruffling, transformation, activation of
c-Jun transcriptional activity and a co-activator of NF$\kappa$B (verbatim)~\cite{,werbajh2000rac}.
Activated Rac also regulates production of superoxide in neutrophils and macrophages.

Interestingly, the alternative exon of \textit{PFN2} seems to disrupt the coding sequence that encode profilin domain (Pfam:PF00235).
The profilin domain is essential for almost all orgamisms and its functions include regulating actin polymerization, controlling
complex network of molecular interaction and transmitting signals from small-GTPase pathway.
It also binds to Rac effector molecules and a number of other ligands~\cite{witke2004role}.
The skipped exon of \textit{PPP1R12A} encode part of a protein domain annotated as protein phosphatase 1 (PIRSF:PIRSF038141).

\subsubsection*{Prediction of \textit{cis}-regulatory elements in alternative splicing exons}

Mutations that disrupt or create exonic splicing enhancers could alter splicing patterns leading to aberrant alternative splicing.
In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing.
A number of disease-associated single-nucleotide polymorphims in coding regions (cSNPs) that affect exon splicing enhancers (ESEs) and
silencers (ESSs) have been characterized~\cite{blencowe2000exonic, wang2007splicing}.

To investigate the possibility that differential isoforms expression is caused by SNPs in coding exons, we obtained exon sequences from line 6
and used Human Splicing Enhancers (HSE)~\cite{} to determine whether SNPs from line 7 could alter predicted ESEs or ESSs.
SNPs in exons of \textit{ITGB2, RAC3, PFN2} and \textit{PPP1R12A} between line 6 and 7 are listed in table~\ref{}.

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection{Gene Models Construction}
Due to lack of complete gene models for chickens, we employed two methods to
construct gene models from RNA-Seq reads.
First, short reads were quality trimmed with conditri/2.1~\cite{} \texttt{(-cutfirst 10)}
and assembled using Velvet/1.21~\cite{} and Oases/0.2.06~\cite{} to obtain long transcripts.
Assembly was done with k value ranges from 21 to 31 for both local and global assembly
(described in Gimme paper~\cite{}).
Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean~\cite{}.
All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
chromosomes removed) with BLAT~\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
Filtered alignments from BLAT were then used to produced gene models using Gimme~\cite{}.

Second, short reads were aligned to reference genome using Tophat/2.0.5~\cite{} and 
gene models were built by Cufflinks2~\cite{} with default parameters.
A combination of gene models from both set of gene models were used in this study.

\subsection{Differential Gene Expression and Gene Ontology}

To identify DE genes, we obtained read counts using multiBamCov command from BEDTools/12.13.1~\cite{}
for all datasets.
Then DE gene were identified by DESeq/1.10.1~\cite{S:2010fu} from Bioconductor.
Data from single- and paired-end datasets from the same line were treated as biological replicates.
To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq/1.10.0~\cite{}.
P-values were corrected by Benjamini-Hochberg multiple testing correction.
Genes, patways and GO terms with corrected P-value $<0.05$ were considered significant.

\subsection{Differential Splicing Event}
Gene models were converted to alternative splicing models using a Python script.
In order to increase sensitivity, read counts from single- and paired-end samples were combined and treated
as single-end reads for splicing event analysis with MISO/0.4.9~\cite{Katz:2010iv}.
Splicing events with Bayes factor $>10$ and $\Delta\Psi>0.20$ were considered significant.
Read coverages and $\Psi$ distributions were plotted using Sashimi plot~\cite{Katz:2013vx}.

\subsection{Variant Calling and \textit{In Silico} Splicing Analysis}
Variants were called using mpileup command from SAMTools/0.1.18~\cite{} and BCFTools~\cite{}.
Only variants with quality score $\ge20$ were used for mutation analyses.
Exon enhancers and suppressors were predicted using Human Splicing Enhancer web portal~\cite{}.
Human default parameter settings were used in all analyses.
The following regulatory sequences were used to determined the effect of variants:
HSF integrated matrices for serine/arginine-rich proteins (SRp40, SC35, SF2/ASF, SF2/ASF,
IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE), RESCUE-ESE hexamer (RESCUE-ESE),
putative 8-mer ESEs (PESEs) and putative 8-mers exonic splicing silencers (PESSs),
exon-identity and intron-identiry elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding
motifs, and Fas exonic splicing silencers.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

% \section*{References}
% The bibtex filename
\bibliography{ref}{}

\section*{Figure Legends}
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=7in]{line67_KEGG_cleveland.pdf}
\end{center}
\caption{
{\bf Enriched KEGG pathways.} Rest of figure 2  caption.  Caption 
should be left justified, as specified by the options to the caption 
package.
}
\label{line67_kegg}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=6in]{gga04145_degenes_multi.png}
\end{center}
\caption{
{\bf Enriched KEGG pathways in line 7.}  Rest of figure 2  caption.  Caption 
should be left justified, as specified by the options to the caption 
package.
}
\label{kegg_phagosome}
\end{figure}

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[width=6in]{degenes_venn.pdf}
    \end{center}
    \caption{
        {\bf Differential-expressed genes in response to MDV infection}
    }
    \label{degenes_venn}
\end{figure}

\section*{Tables}

\begin{table}[!ht]
\caption{
\bf{Gene models summary}}
\begin{tabular}{ccc}
\hline
Method & Gene & Isoform \\
\hline
Assembly & 25,290 & 54,044 \\
Cufflinks & 21,345 & 36,218 \\
Combined & 24,980 & 46,613 \\
\hline
\end{tabular}
\begin{flushleft}Number of genes and isoforms from gene models
built from \textit{de novo} assembly and Cufflinks.
Combination of both methods decreased the number of genes
and isoforms by merging fragmented transcripts to
form more complete gene models.
\end{flushleft}
\label{tab:gene_models}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Genes regulated in opposite direction}}
\begin{tabular}{cccccc}
\hline
Gene & Entrez ID & Fold Change & \\
 & & Line 6 & line 7 \\
\hline
TAF1D & 419002 & +1.08 & -1.13 \\
LYG2 & 395708 & -1.70 & +2.57 \\
SFTPA1 & 395308 & -5.12 & +3.51 \\
LOC424145 & 424145 & +1.57 & -1.68 \\
LL & 423630 & -3.30 & +8.31 \\
SERPINB10 & 395715 & -1.12 & +0.87 \\
\hline
\end{tabular}
\begin{flushleft}
    (-) down-regulated, (+) up-regulated
\end{flushleft}
\label{tab:opposite}
\end{table}

\begin{table}[!ht]
\caption{
\bf{DEU between Line 6 and Line 7 in infected birds}}
\begin{tabular}{cccccccc}
\hline
& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
\hline
SE & chr2:13729.ev2 & ENSGALG00000010973 & \textbf{TRA2A} & 0.45 & \textbf{0.75} & 0.58 & 0.50 \\
SE & chr5:634.ev3 & ENSG00000127152* & BCL11B & 0.07 & \textbf{0.31} & 0.06 & 0.04 \\
SE & chr5:22858.ev1 & ENSGALG00000011127 & BCL11B & 0.07 & \textbf{0.30} & 0.06 & 0.04 \\
SE & chr2:13065.ev1 & ENSGALG00000013137 & INO80C & 0.14 & \textbf{0.35} & 0.96 & 0.86 \\
SE & chr20:14995.ev1 & ENSG00000124193* & SRSF6 & 0.43 & \textbf{0.72} & 0.54 & 0.34 \\
A5SS & chr7:24049.ev1 & ENSGALG00000009824 & \textbf{C7H2ORF77} & 0.50 & \textbf{0.73} & 0.32 & 0.38 \\
A5SS & chr3:18098.ev1 & ENSGALG00000013821 & \textbf{GEMIN6} & 0.84 & \textbf{0.61} & 0.81 & 0.85 \\
A3SS & chr1:6450.ev1 & ENSGALG00000027665 & \textbf{SYNGR1} & 0.46 & \textbf{0.22} & 0.68 & 0.60 \\
A3SS & chr2:13270.ev1 & ENSGALG00000026498 & \textbf{Unknown} & 0.12 & \textbf{0.71} & 0.10 & 0.34 \\
A3SS & chr19:12399.ev1 & ENSGALG00000005685 & \textbf{KSR1} & 0.23 & \textbf{0.56} & 0.28 & 0.35 \\
A3SS & chr1:6437.ev1 & ENSGALG00000012050 & \textbf{TNRC6B} & 0.57 & \textbf{0.39} & 0.95 & 0.93 \\
A3SS & chr18:11800.ev1 & ENSGALG00000002859 & \textbf{RAC3} & 0.31 & \textbf{0.15} & 0.33 & 0.39 \\
\hline
SE & chr17:11370.ev2 & ENSGALG00000004971 & URM1 & \textbf{0.07} & \textbf{0.03} & 0.18 & 0.23 \\
SE & chr2:12495.ev3 & ENSGALG00000005582 & \textbf{KLHL18} & \textbf{0.44} & \textbf{0.27} & 0.56 & 0.52 \\
SE & chr7:24327.ev1 & ENSGALG00000007511 & \textbf{ITGB2} & \textbf{0.17} & \textbf{0.22} & 0.02 & 0.01 \\
SE & chr2:12984.ev1 & ENSGALG00000012809 & ECI2 & \textbf{0.15} & \textbf{0.29} & 0.58 & 0.49 \\
SE & chr8:25247.ev1 & ENSGALG00000028790 & DNASE2B & \textbf{0.38} & \textbf{0.50} & 0.29 & 0.26 \\
SE & chr1:4653.ev1 & ENSGALG00000011682 & CNOT4 & \textbf{0.57} & \textbf{0.63} & 0.40 & 0.40 \\
SE & chr20:14907.ev3 & ENSGALG00000006522 & \textbf{HCK} & \textbf{0.46} & \textbf{0.59} & 0.99 & 0.97 \\
SE & chr4:21539.ev1 & ENSGALG00000015709 & \textbf{TACC3} & \textbf{0.87} & \textbf{0.94} & 0.76 & 0.72 \\
SE & chr2:32201.ev1 & ENSGALG00000012258 & GOLGA4 & \textbf{0.33} & \textbf{0.34} & 0.51 & 0.61 \\
SE & chr3:19470.ev1 & ENSGALG00000019979 & DYNLT1 & \textbf{0.23} & \textbf{0.22} & 0.02 & 0.02 \\
SE & chr19:12391.ev1 & ENSGALG00000005522 & \textbf{DYNLL2} & \textbf{0.01} & \textbf{0.02} & 0.19 & 0.25 \\
A5SS & chr6:23812.ev1 & ENSG00000107651* & \textbf{SEC23IP} & \textbf{0.02} & \textbf{0.12} & 0.50 & 0.43 \\
A5SS & chr2:12775.ev1 & ENSGALG00000011488 & \textbf{CMTM7} & \textbf{0.55} & \textbf{0.67} & 0.37 & 0.41 \\
A3SS & chr8:25257.ev1 & ENSGALG00000008939 & \textbf{FUBP1} & \textbf{0.58} & \textbf{0.74} & 0.41 & 0.46 \\
A3SS & chr9:25750.ev1 & ENSGALG00000010410 & \textbf{PFN2} & \textbf{0.71} & \textbf{0.78} & 0.53 & 0.50 \\
A3SS & chr4:21230.ev1 & ENSGALG00000011476 & \textbf{SEPT11} & \textbf{0.22} & \textbf{0.14} & 0.40 & 0.42 \\
A3SS & chr4:20185.ev1 & ENSGALG00000008507 & THOC2 & \textbf{0.48} & \textbf{0.47} & 0.31 & 0.22 \\
A3SS & chr11:8703.ev1 & ENSGALG00000020987 & \textbf{ZDHHC7} & \textbf{0.42} & \textbf{0.23} & 0.57 & 0.55 \\
A3SS & chr4:21136.ev1 & ENSGALG00000027908 & \textbf{LOC42228} & \textbf{0.72} & \textbf{0.61} & 0.87 & 0.91 \\
\hline
SE & chr12:8987.ev1 & ENSGALG00000008320 & EDEM1 & 0.96 & 0.99 & 0.91 & \textbf{0.72} \\
SE & chr4:20411.ev1 & ENSGALG00000023199 & HNRPDL & 0.39 & 0.40 & 0.30 & \textbf{0.18} \\
SE & chrZ:26257.ev1 & ENSGALG00000001745 & PSTPIP2 & 0.06 & 0.04 & 0.13 & \textbf{0.26} \\
SE & chr1:6316.ev3 & ENSG00000058272* & PPP1R12A & 0.99 & 0.97 & 0.96 & \textbf{0.77} \\
SE & chr1:4478.ev2 & ENSGALG00000009029 & TSPAN12 & 0.09 & 0.15 & 0.20 & \textbf{0.46} \\
SE & chr4:20075.ev1 & ENSGALG00000006157 & DDX26B & 0.67 & 0.61 & 0.58 & \textbf{0.84} \\
SE & chr6:23515.ev1 & ENSGALG00000003861 & HERC4 & 0.31 & 0.37 & 0.45 & \textbf{0.06} \\
SE & chr26:16840.ev1 & ENSGALG00000000533 & SRSF3 & 0.36 & 0.38 & 0.30 & \textbf{0.16} \\
SE & chr11:8507.ev2 & ENSGALG00000000904 & \textbf{C11H16ORF57} & 0.91 & 0.98 & 0.84 & \textbf{0.78} \\
SE & chr1:4323.ev1 & ENSGALG00000006409 & PODXL & 0.21 & 0.34 & 0.26 & \textbf{0.13} \\
SE & chrZ:26582.ev2 & ENSGALG00000014642 & LOC374195 & 0.60 & 0.57 & 0.70 & \textbf{0.81} \\
SE & chr6:23833.ev1 & ENSG00000175029* & CTBP2 & 0.38 & 0.38 & 0.23 & \textbf{0.12} \\
A5SS & chr15:10720.ev1 & ENSGALG00000002487 & \textbf{SFSWAP} & 0.58 & 0.73 & 0.55 & \textbf{0.41} \\
A5SS & chr7:24350.ev1 & ENSGALG00000008038 & \textbf{SF3B1} & 0.41 & 0.57 & 0.55 & \textbf{0.31} \\
A3SS & chr2:13147.ev1 & ENSGALG00000014915 & \textbf{THOC1} & 0.33 & 0.48 & 0.33 & \textbf{0.23} \\
A3SS & chr23:15908.ev1 & ENSGALG00000000720 & LOC419563 & 0.12 & 0.06 & 0.17 & \textbf{0.30} \\
A3SS & chr8:25157.ev1 & ENSGALG00000005162 & RNPC3 & 0.45 & 0.64 & 0.58 & \textbf{0.33} \\
A3SS & chrZ:27197.ev1 & ENSGALG00000000189 & \textbf{YTHDC2} & 0.44 & 0.59 & 0.42 & \textbf{0.32} \\
A3SS & chr23:15983.ev1 & ENSG00000163875* & \textbf{MEAF6} & 0.28 & 0.57 & 0.40 & \textbf{0.29} \\
A3SS & chr5:21970.ev1 & ENSGALG00000009421 & \textbf{SRSF5} & 0.55 & 0.72 & 0.54 & \textbf{0.39} \\
\hline
SE & chr27:17351.ev1 & ENSGALG00000001107 & GOSR2 & 0.73 & 0.92 & 0.38 & 0.60 \\
\hline
\end{tabular}
\begin{flushleft}
    *human homologs, Un=uninfected, Inf=infected
\end{flushleft}
\label{tab:line67i_diff_line67u}
\end{table}

\begin{table}[!ht]
\caption{
\bf{DEU Genes in Spliceosome pathway}}
\begin{tabular}{cccccc}
\hline
Gene ID &  Ensembl & Symbol  & $-log2$FC* & \\
        & & & Line 6 & Line 7 \\
\hline
chr26:16840 & ENSGALG00000000533 & SRSF3 & 0.43 & 0.18 \\
chr2:13147 & ENSGALG00000014915 & THOC1 & -- & 0.21 \\
chr5:21970 & ENSGALG00000009421 & SRSF5 & -- & 0.18 \\
chr2:13729 & ENSGALG00000010973 & TRA2A & -- & -- \\
chr4:20185 & ENSGALG00000008507 & THOC2 & -- & -- \\
chr7:24350 & ENSGALG00000008038 & SF3B1 & -- & -- \\
\hline
\end{tabular}
\begin{flushleft}
    * FDR $<$ 0.05
\end{flushleft}
\label{tab:spliceosome}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Pathways containing \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A}}}
\begin{tabular}{ccc}
\hline
Pathway ID &  Description & Gene \\
\hline
hsa04810 & Regulation of actin cytoskeleton & \textit{RAC3, ITGB2, PPP1R12A, PFN2} \\
hsa04015 & RAP1 signaling pathway & \textit{RAC3, ITGB2, PFN2} \\
hsa04650 & Natural killer cells cytotoxicity & \textit{RAC3, ITGB2} \\
hsa05416 & Viral myocarditis & \textit{RAC3, ITGB2} \\
hsa04510 & Focal adhesion & \textit{RAC3, PPP1R12A} \\
\hline
\end{tabular}
\begin{flushleft}
\end{flushleft}
\label{tab:integrin}
\end{table}

\begin{table}[!ht]
\caption{
\bf{SNPs in exons}}
\begin{tabular}{ccccccc}
\hline
Gene &  Chromosome & Position & Reference & Alternative & & Strand \\
&  & & & Line 6 & Line 7 \\
\hline
ITGB2 & 7 & 7183696 & C & \textit{T} & C & - \\
PFN2 & 9 & 23221934 & -  - & & \textit{AA} & + \\
\hline
\end{tabular}
\begin{flushleft}
\end{flushleft}
\label{tab:spliceosome}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Results from human splicing finder}}
\begin{tabular}{|l|p{1cm}|p{3cm}|l|l|l|l|}
\hline
Gene &  cDNA Position & Linked SR protein & Type & Reference Motif & Mutant Motif & Variation \\
\hline
PFN2 & 2068 & Tra2-$\beta$ & ESE & AAAAT (81.02) & AAAAa & +16.19\% \\
 & 2069 & Tra2-$\beta$ & ESE & & AAAaa (94.14) & New site \\
 & 2070 & Tra2-$\beta$ & ESE & & AAAaaT (81.02) & New site \\
 & 2066 & & ESS$^{2}$ & & ACAAAAaa (38.13) & New site \\
 & 2067 & & ESS$^{2}$ & & CAAAAaaT (28.85) & New site \\
\hline
ITGB2 & 20 & SC35 & ESE$^{1}$ & TGCTCATG (78.19) & & Site broken \\
 & 22 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (91.15) & +18.03\% \\
 & 22 & SF2/ASF(IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (89.23) & +15.53\% \\
 & 22 & SF2/ASF, SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (91.15) & +22.27\% \\
 & 22 & SF2/ASF, SF2/ASF & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (89.23) & +19.69\% \\
 & 24 & SRp55 & ESE$^{1}$ & & CACGGA (79.30) & New site \\
 & 26 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & & CGGAGAT (80.00) & New site \\
 & 26 & SF2/ASF & ESE$^{1}$ & & CGGAGAT (75.36) & New site \\
 & 21 & & ESS$^{3}$ & & GCTCACGG (63.35) & New site (-5.59) \\
 & 23 & & ESS$^{3}$ & TCATGGAG (61.41) & & Site broken (3.16) \\
 & 23 & & ESS$^{3}$ & ATGGAGAT (64.83) & ACGGAGAT & -5.16\% \\
 & 24 & & ESS$^{4}$ & CATGGA (65.48) & CACGGA (65.48) & 0\% \\
\hline

\end{tabular}
\begin{flushleft}
    $^{1}$ESE Finder matrices for SRp40, SC35, SF2/ASF and SRp55 proteins.
    $^{2}$Predicted PESS Octamers from Zhang \& Chasin.
    $^{3}$Silencer motif from Sironi et al.
    $^{4}$hnRNP motif.
\end{flushleft}
\label{tab:spliceosome}
\end{table}

\end{document}

### 2014-06-02 paper.tex
updated

@@
 underlie the susceptibility and resistance of line 6 and 7.
-In this study, we identified differentially expressed genes and isoforms in chicken line 6 and 7
-in response to Marek's disease infection. Results from pathway analysis and functional analysis
-show that changes in expression of genes and isoforms may contribute to resistance and
-susceptibility of the disease. We also identified many single nucleotide polymorphisms (SNPs) between
-line 6 and 7 that may cause alteration of alternative splicing, which in turn results in allele
-specific expression.
+In this study, we identified differentially expressed genes and isoforms in response to Marek's disease infection
+from chickens line 6 and 7 and performed pathway and functional analysis.
+Results show that changes in expression of genes and isoforms involved in innate immune responses in line 6 may contribute
+to resistance of the disease. We also identified many single nucleotide polymorphisms (SNPs) between
+line 6 and 7 that may cause aberrant alternative splicing, which results in isoform with distictive functions.
 
@@
 
-***Need to add intro about breed selection against infection***
-
 Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
@@
 binding motifs have been identified in some organisms including human.
-Abberation of those binding sites have been shown to be associated with increasing susceptibility and severity of diseases~\cite{}.
+Mutations that disrupt or create exonic splicing enhancers could alter splicing patterns leading to aberrant alternative splicing.
+In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing.
+A number of disease-associated single-nucleotide polymorphims in coding regions (cSNPs) that affect exon splicing enhancers (ESEs) and
+silencers (ESSs) have been characterized~\cite{blencowe2000exonic, wang2007splicing}.
 We hypothesized that SNPs between line 6 and 7, which are located in exons, are responsible for differential
 expression of some isoforms between chicken line 6 and line 7.
-To investigate those SNPs, we compared exon sequences of genes identified to have differential exon usage
-from both lines and then predicted their effects on binding motifs.
+We reported some genes that could be regulated by \textit{cis-}regulatory elements that may also play a role in
+immune reponses.
 
@@
 
-\subsection*{Custom Gene Annotations}
-A large number of reads mapped to intergenic and intronic regions (table 1, supp.)
-regions of Ensembl annotations release 73.
-indicates that many genes and isoforms are not included in the annotations.
-To extensively study gene and isoform expression, we employed \textit{de novo}
-transcriptome assembly and a reference-guided assembly method to build
-custom gene annotations (see methods).
-The number of genes and isoforms is summarized in Table~\ref{tab:gene_models}.
-Genes were identified by aligning against non-redundant protein database from NCBI.
-We found several genes that not annotated in either RefSeq or Ensembl, including chicken specific genes
-as well as genes found in other organisms (Table~\ref{}).
-
 \subsection*{Differential gene expression in response to MDV infection}
 
+To extensively study gene and isoform expression, we incorporated \textit{Ensembl} gene models
+release 73 with \textit{de novo} and a reference-guided transcriptome assembly to build custom
+gene models.
+The models, therefore, include both \textit{Ensembl} annotated transcripts
+and putative transcripts and genes.
+The advantage of using custom gene models is it allows an investigation of unannotated genes and
+isoforms, which is necessary for in-depth transcriptome study of an immune system.
+
 Many genes were found to be differentially expressed (DE) between control and infected
-chicks in line 6 and line 7.
+chickens in line 6 and line 7.
 While the number of unique downregulated genes in both lines are approximately equal,
@@
 in figure~\ref{degenes_venn}.
-
-Some DE genes have been reported from previous Microarray studies to be differentially expressed
-in response to MDV infection.
-For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
-chicken with MHC allele B\textsuperscript{19} at 4 d.p.i.
-It was also found to be down-regulated about 3-fold in line 7 (susceptible) in this study.
-Similary, \textit{GMZA} gene, which has been reported to be upregulated across genetically different chickens,
-was also found to be highly upregulated.
-In contrast, some genes that have been reported to be highly expressed in resistant chickens were
-downregulated in both lines. Those genes are \textit{AMIGO2, MMP13} and \textit{CLEC3B}, which were
-found to be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
-Several genes involved in innate immune response were differentially expressed in resistant chickens:
-\textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5 was upregulated.
-Other immune genes reported to be highly expressed in susceptible chickens including
-\textit{AVD, ART1, NOS2, CXCL13L2, MX1} and \textit{SOCS1}~\cite{kaiser2003differential},
-were found to be highly upregulated in both lines.
-However, some discrepancies may be due to the difference in sensitivity between
-Microarray and RNA-Seq methods.
-
-Important cytokines such as \textit{IFN$\gamma$} and \textit{IFN$\beta$} were found to be highly
-upregulated in infected chickens in both line 6 and 7 as well as \textit{INF$\alpha$3} (table~\ref{}).
-However, expressions of genes encoding their corresponding receptors
-were not different in line 6, but upregulated in line 7.
-Keiser et al reported that expressions of \textit{IL6} and \textit{IL18} were different between
-line 6 and 7.
-Both genes were only upregulated in line 7 in early stage of infection (3-5 d.p.i).
-At 10 d.p.i, the expression of \textit{IL6} was different from the control chickens in both lines;
-whereas, the level of \textit{IL18} was only different in susceptible chickens.
-Our results show the same pattern of expression of both cytokines.
-
-In contrary, \textit{IL15} was reported no differences in expression between lines in the same study.
-However, in this study, it was only upregulated in line 7 compared to uninfected control chickens.
-Expression of \textit{IL15} is induced by \textit{TLR9}, which binds to non-methylated CpG
-residues present in genomes of many DNA viruses, including herpes simplex virus.
-This cytokine auto-regulates the expression of \textit{CD40}, which is
-a transmembrane receptor required for activation of macrophages by CD4 T cells.
-As a result, \textit{CD40} was only upregulated in line 7.
-
 Interestingly, some genes that were differentially expressed in both lines were
@@
 \textit{LIMS1} is involved in cell differentiation and proliferation and
-\textit{PPARG} is a suppressor of \textit{NFK$\beta$}-mediated proinflamatory response.
+\textit{PPARG} is a suppressor of \textit{NF$\kappa$B}-mediated proinflamatory response.
 On the other hand, nearly all genes upregulated in line 6 but downregulated in line 7
@@
 immune responses are still upregulated.
+In addition, type I interferon (\textit{IFN-$\gamma$} and \textit{IFN-$\beta$}) as well as \textit{INF-$\alpha$3} were found to be highly
+upregulated in infected chickens in both line 6 and 7 (table~\ref{tab:cytokines}).
+However, expressions of genes encoding their corresponding receptors
+were not different in line 6, but upregulated in line 7.
+This could also reflect the ongoing immune responses in line 7.
+
+Some DE genes have been reported from previous Microarray studies were also found to be
+differentially expressed in this study.
+For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
+chicken with MHC allele B\textsuperscript{19} at 4 d.p.i~\cite{sarson2008transcriptional}.
+It was also found to be down-regulated about 3-fold in line 7 (susceptible).
+Similary, \textit{GMZA} gene reported to be upregulated across genetically different chickens
+(B$^{19}$, B$^{21}$ alleles), was also found to be highly upregulated.
+In contrast, some genes that have been reported to be highly expressed in resistant chickens were
+downregulated in both lines. Those genes are \textit{AMIGO2, MMP13} and \textit{CLEC3B}, which were
+found to be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
+% Several genes involved in innate immune response were differentially expressed in resistant chickens:
+% \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5 was upregulated.
+Other immune genes reported to be highly expressed in susceptible chickens (line 7$_{2}$) including
+\textit{AVD, ART1, NOS2, CXCL13L2, MX1} and \textit{SOCS1}~\cite{smith2011systems}
+were also found to be highly upregulated in both line 6 and 7.
+However, our results show the similar expression pattern of \textit{IL6} and \textit{IL18}, which were
+only upregulated in line 7 in early stage of infection (3-5 d.p.i).
+
+In contrary, \textit{IL15} has been reported non differential-expressed between control and infected chickens in
+line 6$_{1}$ and 7$_{2}$~\cite{kaiser2003differential}; however,
+it was only upregulated in line 7 in our study.
+Expression of \textit{IL15} is induced by \textit{TLR9}, which binds to non-methylated CpG
+residues present in genomes of many DNA viruses, including herpes simplex virus.
+This cytokine auto-regulates the expression of \textit{CD40}, which is
+a transmembrane receptor required for activation of macrophages by CD4 T cells.
+Consequently, \textit{CD40} was only upregulated in line 7 (data not shown).
 
@@
 To determine pathways that were perturbed during the infection, data were
-analysed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{}.
-Significantly perturbed pathways (adjusted p-value $< 0.1$) from both lines that involved in
-immune response include Toll-like receptor signaling, cytokine-cytokine receptor interaction,
+analysed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{young2010method}.
+Significantly perturbed pathways (FDR $< 0.1$) from both lines that involved in
+immune response include Toll-like receptor signaling pathway, cytokine-cytokine receptor interaction,
 intestinal immune network for IgA production and cell adhesion molecules (CAMs).
 Some other pathways important in response to viral infection and only significantly
-enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway pathway,
+enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway,
 NOD-like receptor signaling pathway and lysosome.
@@
 Although MHC I gene (\textit{BF1}) was differentially expressed in both lines, other genes involved in
-expressing newly synthesized MHC I were differentially expressed in line 7.
-This suggests that at the stage of infection, new MHC I molecules were actively produced and expressed to
-present antigen.
+expressing newly synthesized MHC I were only upregulated in line 7
+suggesting that new MHC I molecules were actively produced and expressed on cell surface.
 Furthermore, gene ontology analysis of biological processes (GO:BP) shows that categories involved in
-both adaptive and innate immune responses were enriched in line 7.
+both adaptive and innate immune responses were enriched in line 7 (supplementary material).
 On the other hand, only categories involved in innate immune responses were enriched in line 6.
-In addition, the apoptosis pathway was significantly enriched in line 7, which suggests that the programmed
+In addition, enrichment of apoptosis pathway in line 7 indicates that the programmed
 cell death could be induced by CTL response to eliminate the viruses.
 
-At this stage of infection, our results suggest that lytic infection of MDV stimulated
-both innate and the adaptive immune response, which led to activation of T cells in line 7.
+At this stage of infection, our results suggest that lytic infection of MDV stimulates
+both innate and the adaptive immune response, which leads to activation of T cells in line 7.
 Only activated T cells are infected by MDV, therefore, the lytic phase could facilitate the spread of
 the viruses by enhancing expansion of activated T cells.
-In contrast, the immune response in line 6 could have been able to drive the viruses to latency phase early
-without eliciting adaptive immune response.
+In contrast, the immune response in line 6 could be able to drive the viruses to latency phase early
+and only activate the adaptive immune response in a very short duration.
 Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell contact between B cells and
-T cells during antigen presentation or B cells activation by helper T cells.
+T cells during antigen presentation or B cells activation by T helper cells.
 Therefore, it is beneficial for the host to restrain such contact.
-Limiting of T cells activation in line 6 could attribute to several mechanisms.
-The innate immune response could be highly effective and could activate a strong adaptive immune response that
-rapidly controls viral replication and forces the viruses into latency phase
-Alternatively, the innate immune resposne itself could be highly effective in limiting viral replication.
-Conversely, the innate immune response in line 6 could be inefficient in eliciting an adaptive immune
-response. Since MDVs cannot replicate in APCs~\cite{}, the replication of MDVs is naturally limited.
+However, it is not clear how chickens in line 6 control the lytic infection of MDV.
+% Two mechanisms have been speculated to contribute to MD resistance.
+% First innate immune responses could be highly effective and could activate
+% strong adaptive immune responses that rapidly control viral replication and force the viruses to undergo latent phase.
+% Second, innate immune resposnes itself could be highly effective in limiting viral replication~\cite{smith2011systems}.
 
@@
 susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
-Studying differential isoform expression could therefore shed lights into an inherent different between 
-line 6 and 7 in that confers resistance and susceptibility to MD.
-Microarray technology has been used to study isoform expression in several studies, but
+Studying differential isoform expression could therefore shed lights into inherent differences between 
+line 6 and 7 that confer resistance or susceptibility to MD.
+
+In the past decades, microarray technology has been used to study gene and isoform expressions in many studies, but
 its sensitiviy for detection of structurally similar isoforms is low and known or predicted annotations
-are required~\cite{kane2000assessment}.
-Although RNA-Seq method can provide reliable estimate of exon expressions compared to Microarray~\cite{pan2008deep}
+are required to design probes~\cite{kane2000assessment}.
+Although RNA-Seq method can provide a reliable estimate of an exon expression compared to microarray~\cite{pan2008deep}
 and is not constrained to those limitations, studying isoform
-expression using RNA-Seq is still not straightforward because of short read lengths.
+expressions using RNA-Seq is still not straightforward because of short read lengths.
 Reads from current Illumina technology are generally not long enough to span across all exons in an isoform.
@@
 Using MISO with exon-centric method, only reads spanning across a few exons are used and only exons
-involved in the splicing event are examined.
+involved in a splicing event are examined.
 The expression of exon inclusion is calculated as percent spliced in (Psi or $\Psi$), which can be used to
@@
 List of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
-in infected chickens is shown in table~\ref{tab:line67i_diff_line67u}.
+in infected chickens is shown in table~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
+and~\ref{tab:line67i_diff_line67u_three}.
 Genes can be categorized roughly into four groups based on the pattern of $\Psi$ across control and
@@
 Two genes involved in cell-cell contact regulated by integrins are also in this group.
-\textit{PPP1R12A} is a mysosis phosphatase that regulates the interaction of actin and myosin downstream of GTPase
+\textit{PPP1R12A} is a myosin phosphatase that regulates the interaction of actin and myosin downstream of GTPase
 Rho proteins.
@@
 significantly different between infected chickens in line 6 and 7.
-However, there is a significant difference between control and infected chickens in line 7.
-This gene encodes traffiking membrane protein important for transporting proteins from \textit{cis-} to \textit{trans-}
-golgi network.
+Also, there is a significant difference between control and infected chickens in line 7.
+This gene encodes traffiking membrane protein important for transporting proteins from \textit{cis-} to
+\textit{trans-}golgi network.
 
@@
 and \textit{CR3} receptors.
-Upon binding, \textit{LFA-1} and \textit{CR3}.
+
+(more on \textit{LFA-1} and \textit{CR3} to be added).
+
 As discussed above, \textit{RAC3, HCK, DYLNT1, DYNLL2, SEPT11, PFN2, ZDHHC7, PODXL} and \textit{PPP1R12A}
-could be activated by the integrins because of their roles in cell rearrangement.
+could be activated by the integrins because of their roles in cell rearrangement or molecular interaction.
 Interestingly, we found that in human, \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A} are in the regulation
-of actin cytoskeleton pathway (figure~\ref{actin_pathway}).
+of actin cytoskeleton pathway (figure~\ref{kegg_actin}).
 Moreover, at least two of these genes are co-present in other three pathways (table~\ref{tab:integrin}) that
 involved in immune responses.
-This strongly suggests that these four genes could contribute to inherent difference in immune response between
+This strongly suggests that these four genes could contribute to difference in immune responses between
 line 6 and 7 even though their expressions were not significantly different between control and infected groups.
 
-\subsubsection*{Splicing factors}
-
-The second group includes genes encode protein compoments of spliceosome or are involved in pre-mRNA splicing.
-Those genes are \textit{TRA2A, SRSF6, GEMIN6, SRSF3, HNRNPDL, SFSWAP, THOC1, THOC2, RNPC3} and \textit{SRSF5}.
-Even though alternative splicing is regulated by splicesome, it is difficult to elucidate the effect of alternative
-splice forms of spliceosome subunits on alternative splicing regulation.
-This is in part due to the fact that the spliceosome is the most complex protein structures composed of many different
-subunits each of which could have isoforms differing in function or activity~\cite{}.
+% \subsubsection*{Splicing factors}
+% 
+% (Not sure if I should write about splicing factors, I'm afraid it will make the paper too long and unfocused.)
+% 
+% The second group includes genes encode protein compoments of spliceosome or are involved in pre-mRNA splicing.
+% Those genes are \textit{TRA2A, SRSF6, GEMIN6, SRSF3, HNRNPDL, SFSWAP, THOC1, THOC2, RNPC3} and \textit{SRSF5}.
+% Even though alternative splicing is regulated by splicesome, it is difficult to elucidate the effect of alternative
+% splice forms of spliceosome subunits on alternative splicing regulation.
+% This is in part due to the fact that the spliceosome is the most complex protein structures composed of many different
+% subunits each of which could have isoforms differing in function or activity~\cite{}.
 
@@
 Rac3 is highly homologous to Rac1 and has been reported to possess the ability of promoting membrane ruffling, transformation, activation of
-c-Jun transcriptional activity and a co-activator of NF$\kappa$B (verbatim)~\cite{,werbajh2000rac}.
+c-Jun transcriptional activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.
 Activated Rac also regulates production of superoxide in neutrophils and macrophages.
 
-Interestingly, the alternative exon of \textit{PFN2} seems to disrupt the coding sequence that encode profilin domain (Pfam:PF00235).
+The alternative exon of \textit{PFN2} seems to disrupt the coding sequence that encode profilin domain (Pfam:PF00235).
 The profilin domain is essential for almost all orgamisms and its functions include regulating actin polymerization, controlling
@@
 
-Mutations that disrupt or create exonic splicing enhancers could alter splicing patterns leading to aberrant alternative splicing.
-In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing.
-A number of disease-associated single-nucleotide polymorphims in coding regions (cSNPs) that affect exon splicing enhancers (ESEs) and
-silencers (ESSs) have been characterized~\cite{blencowe2000exonic, wang2007splicing}.
+Although alternative splicing of genes in all groups could be regulated by \textit{cis-} and \textit{trans-}splicing factors,
+genes in group II are more likely to be regulated by \textit{cis-}regulatory factors.
+The ratios of isoform expression in this group were relatively stable within line, but were significantly different between lines.
+Investigation of nucleotide differences within exons of line 6 and 7 could reveal a possible role of cSNPs in regulating
+alternative splicing in this groups.
+We obtained a sequence of an alternative exon from line 6 and used Human Splicing Finder (HSF) to determine whether SNPs from
+line 7 could alter predicted ESEs or ESSs.
+SNPs in exons of \textit{ITGB2} and \textit{PFN2} between line 6 and 7 are listed in table~\ref{tab:deu_snps}.
 
-To investigate the possibility that differential isoforms expression is caused by SNPs in coding exons, we obtained exon sequences from line 6
-and used Human Splicing Enhancers (HSE)~\cite{} to determine whether SNPs from line 7 could alter predicted ESEs or ESSs.
-SNPs in exons of \textit{ITGB2, RAC3, PFN2} and \textit{PPP1R12A} between line 6 and 7 are listed in table~\ref{}.
+For \textit{ITGB2}, a putative ESE motif of SC35 is predicted to be broken when replace a T with a C at position 26 of the cDNA, which
+corresponds to a position 7,183,696 on chromosome 7.
+The disruption of this exon enhancer binding site support the fact that the level of exon inclusion in line 7 is
+practically zero (figure~\ref{itgb2}).
+On the other hand, there is also a putative binding site for an exon suppressor that could promote exclusion of the exon in line 7.
+The SNP is also predicted to affect putative ESE motifs of other splicing factors, however, those motifs do not seem to agree with
+exon expression data.
+
+Exon sequences of \textit{PFN2} from line 6 and 7 differ at a position 2,072 on cDNA or 23,221,934 on chromsome 9.
+A small insertion of two AA nucleotides is predicted to create a new ESE motif for Tra2-$\beta$ splicing regulator,
+whose function as a stabilizer of an enhancer complex~\cite{lopez1998alternative}.
+From exon expreesion data from, $\Psi$ of an exon with alternative splicing increases from 0.20-0.30 to about 0.50 in line 7.
+Possibly, Tra2 increases inclusion of the exon with alternative 3$\prime$ splice site via
+ESE-dependent 3$\prime$ splice site activation (figure~ref{pfn2}).
+Even though the linear distance between Tra2-$\beta$ binding site and the alternative 3$\prime$ splice site is
+greater than 1kb, Tra2 could bind close to the 3$\prime$ splice site if the secondary structure of the mRNA moves the
+ESE motif closer to the splice site.
 
@@
 \subsection{Gene Models Construction}
+
 Due to lack of complete gene models for chickens, we employed two methods to
 construct gene models from RNA-Seq reads.
-First, short reads were quality trimmed with conditri/2.1~\cite{} \texttt{(-cutfirst 10)}
+First, short reads were quality trimmed with conditri/2.1~\cite{}
 and assembled using Velvet/1.21~\cite{} and Oases/0.2.06~\cite{} to obtain long transcripts.
-Assembly was done with k value ranges from 21 to 31 for both local and global assembly
-(described in Gimme paper~\cite{}).
+Assembly was done with hash lengths range from 21 to 31 for both local and global assembly
+(described in Gimme paper).
 Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean~\cite{}.
 All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
-chromosomes removed) with BLAT~\cite{} \texttt{(-t=dna -q=dna -noHead -out=psl -mask=lower -extendThroughN)}.
-Filtered alignments from BLAT were then used to produced gene models using Gimme~\cite{}.
-
-Second, short reads were aligned to reference genome using Tophat/2.0.5~\cite{} and 
-gene models were built by Cufflinks2~\cite{} with default parameters.
-A combination of gene models from both set of gene models were used in this study.
+chromosomes removed) with BLAT~\cite{}.
+Second, reads were aligned to reference genome using Tophat/2.0.5~\cite{} and \textit{Ensemble} gene
+models release 73 was used to guide reference-based assembly by Cufflinks2~\cite{}.
+Alignments from BLAT and models from Cufflinks were then combined to construct gene models by Gimme
+(manuscript in preparation).
 
@@
 
-To identify DE genes, we obtained read counts using multiBamCov command from BEDTools/12.13.1~\cite{}
-for all datasets.
-Then DE gene were identified by DESeq/1.10.1~\cite{S:2010fu} from Bioconductor.
+To identify DE genes, reads were mapped to transcripts by RSEM v.1.10.1~\cite{li2011rsem}, which is
+also used to estimate genes expression and identify DE genes.
 Data from single- and paired-end datasets from the same line were treated as biological replicates.
-To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq/1.10.0~\cite{}.
+To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq v.1.10.0 based on
+chicken KEGG annotations.
 P-values were corrected by Benjamini-Hochberg multiple testing correction.
-Genes, patways and GO terms with corrected P-value $<0.05$ were considered significant.
+Genes, patways and GO terms with corrected P-value $<0.1$ were considered significant.
+Pathview~\cite{luo2013pathview} was used to create a KEGG pathway diagram with colors representing relative level of gene expressions.
 
@@
 Only variants with quality score $\ge20$ were used for mutation analyses.
-Exon enhancers and suppressors were predicted using Human Splicing Enhancer web portal~\cite{}.
+Exon enhancers and suppressors were predicted using Human Splicing Finder web portal~\cite{}.
 Human default parameter settings were used in all analyses.
@@
 motifs, and Fas exonic splicing silencers.
+
+\subsection{Protein domains search}
 
@@
 \section*{Figure Legends}
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=7in]{line67_KEGG_cleveland.pdf}
-\end{center}
-\caption{
-{\bf Enriched KEGG pathways.} Rest of figure 2  caption.  Caption 
-should be left justified, as specified by the options to the caption 
-package.
-}
-\label{line67_kegg}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=6in]{gga04145_degenes_multi.png}
-\end{center}
-\caption{
-{\bf Enriched KEGG pathways in line 7.}  Rest of figure 2  caption.  Caption 
-should be left justified, as specified by the options to the caption 
-package.
-}
-\label{kegg_phagosome}
-\end{figure}
 
@@
 
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics[width=7in]{line67_KEGG_cleveland.pdf}
+    \end{center}
+    \caption{
+        {\bf Enriched KEGG pathways.}
+    }
+    \label{line67_kegg}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics[width=6in]{gga04145_degenes_multi.png}
+    \end{center}
+    \caption{
+        {\bf Phagosome pathway.}
+    }
+    \label{kegg_phagosome}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics[width=6in]{hsa04810_deu_genes.png}
+    \end{center}
+    \caption{
+        {\bf Human regulation of actin cytosekeleton pathway.}
+    }
+    \label{kegg_actin}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics[width=6in]{itgb2_miso.pdf}
+    \end{center}
+    \caption{
+        {\bf ITGB2 exon expression.}
+        A single nuclotide substition (T$>$C) at position 7,183,696 is predicted to disrupt a putative ESE
+        motif which binds SC35 on the skipped exon of IRGB2 gene shown in this figure.
+        The prediction correlate with exclusion of the exon in line 7 ($\Psi \leq 0.02$).
+        A new binding site for a silencer is also predicted and could also promote the exclusion of the exon. 
+    }
+    \label{itgb2}
+\end{figure}
+
+\begin{figure}[!ht]
+    \begin{center}
+        \includegraphics[width=6in]{pfn2_miso.pdf}
+    \end{center}
+    \caption{
+        {\bf PFN2 exon expression.}
+        A small insertion of AA nucleotides at position 23,221,934 is predicted to create a putative ESE motif
+        which binds to Tra2 protein.
+        Tra2 could promote inclusion of exon with alternative 3$\prime$ splice site via ESE-dependent
+        3$\prime$ splice site activation resulting in higher expression of the exon in line 7.
+        Even though the predicted ESE motif is located at a substantial distance from the alternative 3$\prime$ splice site,
+        Tra2 could still regulate the splicing if a secondary structure of the mRNA moves the ESE motif closer to the
+        alternative 3$\prime$ splice site.
+    }
+    \label{pfn2}
+\end{figure}
+
 \section*{Tables}
+
+% \begin{table}[!ht]
+%     \caption{
+%     \bf{Gene models summary}}
+%     \begin{tabular}{ccc}
+%         \hline
+%         Method & Gene & Isoform \\
+%         \hline
+%         Assembly & 25,290 & 54,044 \\
+%         Cufflinks & 21,345 & 36,218 \\
+%         Combined & 24,980 & 46,613 \\
+%         \hline
+%     \end{tabular}
+%     \begin{flushleft}
+%         Number of genes and isoforms from gene models
+%         built from \textit{de novo} assembly and Cufflinks.
+%         Combination of both methods decreased the number of genes
+%         and isoforms by merging fragmented transcripts to
+%         form more complete gene models.
+%     \end{flushleft}
+%     \label{tab:gene_models}
+% \end{table}
 
@@
 \caption{
-\bf{Gene models summary}}
-\begin{tabular}{ccc}
+\bf{Genes regulated in opposite directions}}
+    \begin{tabular}{cccc}
         \hline
-Method & Gene & Isoform \\
+        Gene & Description & $log_{2}$FC & \\
+         & & Line 6 & Line 7 \\
         \hline
-Assembly & 25,290 & 54,044 \\
-Cufflinks & 21,345 & 36,218 \\
-Combined & 24,980 & 46,613 \\
+        LL & Lung lectin & -3.36 & 8.70 \\
+        GIF & Gastric intrinsic factor & -2.15 & 3.11 \\
+        SFTPA1 & Surfactant protein A1 & -4.84 & 3.73 \\
+        SCAF8 & SR-related CTD-associated factor 8 & -8.70 & 8.24 \\
+        LIMS1 & LIM and senescent cell antigen-like domains 1 & -2.26 & 1.33 \\
+        PPARG & Peroxisome proliferator-activated receptor gamma & -6.99 & 2.06 \\
+        C14ORF1 & Chromosome 14 open reading frame 1 & -3.11 & 2.71 \\
+        Unknown & Unknown & -3.08 & 1.62 \\
         \hline
-\end{tabular}
-\begin{flushleft}Number of genes and isoforms from gene models
-built from \textit{de novo} assembly and Cufflinks.
-Combination of both methods decreased the number of genes
-and isoforms by merging fragmented transcripts to
-form more complete gene models.
-\end{flushleft}
-\label{tab:gene_models}
-\end{table}
-
-\begin{table}[!ht]
-\caption{
-\bf{Genes regulated in opposite direction}}
-\begin{tabular}{cccccc}
-\hline
-Gene & Entrez ID & Fold Change & \\
- & & Line 6 & line 7 \\
-\hline
-TAF1D & 419002 & +1.08 & -1.13 \\
-LYG2 & 395708 & -1.70 & +2.57 \\
-SFTPA1 & 395308 & -5.12 & +3.51 \\
-LOC424145 & 424145 & +1.57 & -1.68 \\
-LL & 423630 & -3.30 & +8.31 \\
-SERPINB10 & 395715 & -1.12 & +0.87 \\
+        ATP8A2 & ATPase, aminophospholipid transporter, class I, type 8A, member 2 & 7.66 & -7.80 \\
+        S1PR1 & Sphingosine-1-phosphate receptor 1 & 1.31 & -6.95 \\
+        MED9 & Mediator complex subunit 9 & 8.22 & -3.15 \\
+        DNAJA2 & DnaJ (Hsp40) homolog, subfamily A, member 1 & 2.34 & -6.02 \\
+        RAD17 & RAD17 homolog (S. pombe) & 1.45 & -1.80 \\
+        PSMG3 & Proteosome assembly chaperone 3 & 1.52 & -1.26 \\
+        PNISR & PNN-interacting serine/argining-rich protein & 5.58 & -1.81 \\
+        THOC7 & THO complex 7 homolog (Drosophila) & 7.00 & -6.90 \\
+        YTHDC2 & YTH domain containing 2 & 1.00 & -1.74 \\
+        RPL39 & Ribosomal protein L39 & 3.31 & -1.59 \\
+        CD7 & CD7 molecule & 2.88 & -1.38 \\
+        NDUFB3 & NADH dehydrogenase (uqiquinone) 1 beta subcomplex 3 & 1.03 & -1.34 \\
+        LOC100858785 & Unknown & 1.26 & -1.75 \\
+        Unknown & Unknown & 6.30 & -5.22 \\
         \hline
@@
 \caption{
-\bf{DEU between Line 6 and Line 7 in infected birds}}
+\bf{Cytokine-related gene expression in response to MDV infection}}
+    \begin{tabular}{cccccc}
+        \hline
+        & & $log_{2}$FC & \\
+        Symbol & Description & Line 6 & Line 7 \\
+        \hline
+        IL2RG & Interleukin 2 receptor, $\gamma$ & -- & 0.55 \\
+        IL6 & Interleukin 6 (interferon, $\beta$ 2) & -- & 5.15 \\
+        IL6ST & Interleukin 6 signal transducer (gp130, oncostatin M receptor) & -- & 1.36 \\
+        IL8L1 & Interleukin 8-like 1 & 1.90 & -- \\
+        IL15 & Interleukin 15 & -- & 1.06 \\
+        IL18 & Interleukin 18 (interferon-$\gamma$ inducing factor) & 1.92 & 4.06 \\
+        IL18R1 & Interleukin 18 receptor 1 & 1.94 & 1.64 \\
+        IFNG & Interferon-$\gamma$ & 5.14 & 4.90 \\
+        IFNB & Interferon-$\beta$ & 4.83 & 5.64 \\
+        IFNA3 & Interferon-$\alpha$ 3 & 4.09 & 5.48 \\
+        IFNGR1 & Interferon-$\gamma$ receptor 1 & -- & 2.05 \\
+        IFNGR2 & Interferon-$\gamma$ receptor 2& -- & 0.50 \\
+        IFNAR1 & Interferon-$\alpha$,$\beta$ receptor 1 & -- & 1.46 \\
+        IFNAR2 & Interferon-$\alpha$,$\beta$ receptor 2 & -- & 0.58 \\
+        \hline
+    \end{tabular}
+    \begin{flushleft}
+    \end{flushleft}
+    \label{tab:cytokines}
+\end{table}
+
+\begin{table}[!ht]
+\caption{
+\bf{DEU between Line 6 and Line 7 in infected birds group I}}
 \begin{tabular}{cccccccc}
@@
 SE & chr2:13729.ev2 & ENSGALG00000010973 & \textbf{TRA2A} & 0.45 & \textbf{0.75} & 0.58 & 0.50 \\
-SE & chr5:634.ev3 & ENSG00000127152* & BCL11B & 0.07 & \textbf{0.31} & 0.06 & 0.04 \\
 SE & chr5:22858.ev1 & ENSGALG00000011127 & BCL11B & 0.07 & \textbf{0.30} & 0.06 & 0.04 \\
@@
 A3SS & chr18:11800.ev1 & ENSGALG00000002859 & \textbf{RAC3} & 0.31 & \textbf{0.15} & 0.33 & 0.39 \\
+\hline
+\end{tabular}
+\begin{flushleft}
+    *human homologs, Un=uninfected, Inf=infected.
+    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
+\end{flushleft}
+\label{tab:line67i_diff_line67u_one}
+\end{table}
+
+\begin{table}[!ht]
+\caption{
+\bf{DEU between Line 6 and Line 7 in infected birds group II}}
+\begin{tabular}{cccccccc}
+\hline
+& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
+Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
 \hline
@@
 A3SS & chr4:21136.ev1 & ENSGALG00000027908 & \textbf{LOC42228} & \textbf{0.72} & \textbf{0.61} & 0.87 & 0.91 \\
+\hline
+\end{tabular}
+\begin{flushleft}
+    *human homologs, Un=uninfected, Inf=infected.
+    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
+\end{flushleft}
+\label{tab:line67i_diff_line67u_two}
+\end{table}
+
+\begin{table}[!ht]
+\caption{
+\bf{DEU between Line 6 and Line 7 in infected birds group III and IV}}
+\begin{tabular}{cccccccc}
+\hline
+& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
+Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
 \hline
@@
 \begin{flushleft}
-    *human homologs, Un=uninfected, Inf=infected
+    *human homologs, Un=uninfected, Inf=infected.
+    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
 \end{flushleft}
-\label{tab:line67i_diff_line67u}
+\label{tab:line67i_diff_line67u_three}
 \end{table}
 
-\begin{table}[!ht]
-\caption{
-\bf{DEU Genes in Spliceosome pathway}}
-\begin{tabular}{cccccc}
-\hline
-Gene ID &  Ensembl & Symbol  & $-log2$FC* & \\
-        & & & Line 6 & Line 7 \\
-\hline
-chr26:16840 & ENSGALG00000000533 & SRSF3 & 0.43 & 0.18 \\
-chr2:13147 & ENSGALG00000014915 & THOC1 & -- & 0.21 \\
-chr5:21970 & ENSGALG00000009421 & SRSF5 & -- & 0.18 \\
-chr2:13729 & ENSGALG00000010973 & TRA2A & -- & -- \\
-chr4:20185 & ENSGALG00000008507 & THOC2 & -- & -- \\
-chr7:24350 & ENSGALG00000008038 & SF3B1 & -- & -- \\
-\hline
-\end{tabular}
-\begin{flushleft}
-    * FDR $<$ 0.05
-\end{flushleft}
-\label{tab:spliceosome}
-\end{table}
+% \begin{table}[!ht]
+% \caption{
+% \bf{DEU Genes in Spliceosome pathway}}
+% \begin{tabular}{cccccc}
+% \hline
+% Gene ID &  Ensembl & Symbol  & $-log2$FC* & \\
+%         & & & Line 6 & Line 7 \\
+% \hline
+% chr26:16840 & ENSGALG00000000533 & SRSF3 & 0.43 & 0.18 \\
+% chr2:13147 & ENSGALG00000014915 & THOC1 & -- & 0.21 \\
+% chr5:21970 & ENSGALG00000009421 & SRSF5 & -- & 0.18 \\
+% chr2:13729 & ENSGALG00000010973 & TRA2A & -- & -- \\
+% chr4:20185 & ENSGALG00000008507 & THOC2 & -- & -- \\
+% chr7:24350 & ENSGALG00000008038 & SF3B1 & -- & -- \\
+% \hline
+% \end{tabular}
+% \begin{flushleft}
+%     * FDR $<$ 0.05
+% \end{flushleft}
+% \label{tab:spliceosome}
+% \end{table}
 
@@
 \hline
-Gene &  Chromosome & Position & Reference & Alternative & & Strand \\
-&  & & & Line 6 & Line 7 \\
+Gene &  Chromosome & Position & Reference & Line 6 & Line 7 & Strand \\
 \hline
 ITGB2 & 7 & 7183696 & C & \textit{T} & C & - \\
-PFN2 & 9 & 23221934 & -  - & & \textit{AA} & + \\
+PFN2 & 9 & 23221934 & -  - & - - & \textit{AA} & + \\
 \hline
@@
 \end{flushleft}
-\label{tab:spliceosome}
+\label{tab:deu_snps}
 \end{table}
@@
 Gene &  cDNA Position & Linked SR protein & Type & Reference Motif & Mutant Motif & Variation \\
-\hline
-PFN2 & 2068 & Tra2-$\beta$ & ESE & AAAAT (81.02) & AAAAa & +16.19\% \\
- & 2069 & Tra2-$\beta$ & ESE & & AAAaa (94.14) & New site \\
- & 2070 & Tra2-$\beta$ & ESE & & AAAaaT (81.02) & New site \\
- & 2066 & & ESS$^{2}$ & & ACAAAAaa (38.13) & New site \\
- & 2067 & & ESS$^{2}$ & & CAAAAaaT (28.85) & New site \\
 \hline
@@
 \hline
+PFN2 & 2068 & Tra2-$\beta$ & ESE$^{1}$ & AAAAT (81.02) & AAAAa & +16.19\% \\
+& 2069 & Tra2-$\beta$ & ESE$^{1}$ & & AAAaa (94.14) & New site \\
+& 2070 & Tra2-$\beta$ & ESE$^{1}$ & & AAAaaT (81.02) & New site \\
+ & 2066 & & ESS$^{2}$ & & ACAAAAaa (38.13) & New site \\
+ & 2067 & & ESS$^{2}$ & & CAAAAaaT (28.85) & New site \\
+\hline
 

### 2014-06-20 paper.tex
added a major updates

@@
 Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
-Different MHC alleles have been associated with susceptibility and resistance to the disease.
-However, chickens line 6 and 7 show significant phenotypic differences when chanllenged with
-MDV even with the same MHC allele. Therefore, the major unanswered question is what genetic factors
-underlie the susceptibility and resistance of line 6 and 7.
-In this study, we identified differentially expressed genes and isoforms in response to Marek's disease infection
-from chickens line 6 and 7 and performed pathway and functional analysis.
-Results show that changes in expression of genes and isoforms involved in innate immune responses in line 6 may contribute
-to resistance of the disease. We also identified many single nucleotide polymorphisms (SNPs) between
-line 6 and 7 that may cause aberrant alternative splicing, which results in isoform with distictive functions.
+Understanding the genetic basis for non-MHC resistance would be of both fundamental and applied importance.
+In this study, using computational approaches of RNA-Seq data, we identified
+differentially expressed genes and isoforms in response to MDV infection
+from highly inbred chickens lines that are MD resistant and susceptible,
+and performed pathway and functional analysis.
+Results show that expression of genes and isoforms involved in cell cytoskeleton and
+adhesion molecules in line 6 may contribute to resistance of the disease by limiting
+T cell activation, which in turn prevents the spread of MDV to activated T cells.
+In addition, we identified exonic single nucleotide polymorphisms (SNPs) between
+lines 6 and 7 predicted their role in regulation of alternative splicing.
+Based on our results, we speculate that alternative splice forms of cell adhesion molecules
+and cytoskeleton play and important role in controlling T cell activation in line 6.
 
@@
 
-Marek's disease is an economically significant chicken disease that affects a poultry industry worldwide.
-Total of \$2 billion loss has been reported from outbreaks~\cite{}.
-The disease is caused by highly oncogenic Marek's disease virus (MDV) which causes T-cell lymphoma.
-Vaccination is effective in reducing incidence of tumor formation; however, current MD vaccines do not
-prevent infection or horizontal spread of the virus.
-Improper use of vaccines is a key factor that drives evolution of highly virulent strains in vaccinated
-flocks.
+Marek's disease (MD) is an economically significant chicken disease that affects the
+poultry industry worldwide with estimated annual loss of
+\$2 billion loss has been reported from outbreaks~\cite{morrow2004marek}.
+The disease is caused by the highly oncogenic Marek's disease virus (MDV)
+an alphaherpesvirus that induces T-cell lymphomas in susceptible birds.
+Vaccination is the primary control measure, which is effective in reducing
+incidence of tumor formation.
+However, since MD vaccines are not sterilizing, they do not prevent infection or
+horizontal spread of the virus.
+As a consequence, MDV field strains that overcome vaccinal protection have arisen repeatedly over time.
+Therefore, there is a need for sustainable alternative controls measures, such as improving genetic
+resistance.
 
-Many studies have reported strong associations between MHC alleles and resistance and susceptibility of
-the disease. For example, chickens with MHC allele B$^{21}$ is highly resistant in contrast to chickens with
-B$^{16}$ allele.
-However, chickens line 6 and 7 , even with common allele (B$^6$), exhibit different phenotypic response to
-MD.
-The major unanswered question is what genetic factors contribute to susceptibility and resistance of
-the disease.
+Many studies have reported strong associations between MHC alleles and resistance or susceptibility to
+MD. For example, chickens with MHC allele B$^{21}$ are highly resistant in contrast to chickens with
+B$^{16}$ allele which are highly susceptible.
+ADOL lines 6 and 7, both share (B$^6$) allele, yet exhibit different phenotypic responses to
+MD; e.g., challenge with JM/102W strain with result in 0 and 100\% MD incidence for lines 6 and 7,
+respectively.
+Thus, the major unanswered questions are what genetic factors, especially those that are non-MHC,
+contribute to susceptibility and resistance of the disease and what is the main contributing
+mechanism.
 
-Although gene expression has been shown to have a high correlation with phenotypes in many studies~\cite{}.
-Studies have shown that alternative splice forms play a significant role in many biological events and
-isoform expression levels can provide better signatures for some diseases~\cite{zhang2013isoform}.
-Regulation of alternative splicing by enhancers (ESE) and silencers (ESS) and their corresponding
-binding motifs have been identified in some organisms including human.
-Mutations that disrupt or create exonic splicing enhancers could alter splicing patterns leading to aberrant alternative splicing.
-In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing.
-A number of disease-associated single-nucleotide polymorphims in coding regions (cSNPs) that affect exon splicing enhancers (ESEs) and
-silencers (ESSs) have been characterized~\cite{blencowe2000exonic, wang2007splicing}.
-We hypothesized that SNPs between line 6 and 7, which are located in exons, are responsible for differential
-expression of some isoforms between chicken line 6 and line 7.
-We reported some genes that could be regulated by \textit{cis-}regulatory elements that may also play a role in
-immune reponses.
+In the past decades, significant efforts have been spent to study variations in a global gene
+expression between resistant and susceptible birds using
+microarray and RNA-Seq methods in order to identify non-MHC genes that contribute
+to resistance to MD~\cite{}.
+However, none of the studies have investigated differential expression of alternative isoforms,
+which are known to play a significant role in many biological events including
+immune responses.
+In addition, studies have shown that isoform expression levels can provide better signatures for some
+diseases~\cite{zhang2013isoform}.
+Changes of isoform expression levels are governed partly by two types of \textit{cis-}regulatory
+elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer (ESS),
+which are located within an exon sequence.
+A number of sequence motifs of ESE and ESS have been identified in human and some other organisms
+and could be predicted \textit{in silico}.
+Mutations that disrupt or create those motifs could alter splicing
+patterns leading to aberrant alternative splicing.
+A number of disease-associated single-nucleotide polymorphsims in coding regions
+(cSNPs) that affect ESEs and ESSs have been well
+characterized~\cite{blencowe2000exonic, wang2007splicing}.
+%In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing via this mechanism.
+Therefore, variations in isoform expression could lead to identification of SNPs
+that underlie genetic resistance to MD.
+In this article, we reported differential-expressed genes and isoforms that may contribute to
+resistance to MD as well as cSNPs that can potentially affect isoform expression levels.
 
 % Results and Discussion can be combined.
-\section*{Results and discussion}
+\section*{Results}
 
@@
 Many genes were found to be differentially expressed (DE) between control and infected
-chickens in line 6 and line 7.
-While the number of unique downregulated genes in both lines are approximately equal,
-the number of unique upregulated genes in line 7 is much greater than in line 6 as shown
-in figure~\ref{degenes_venn}.
+chickens in lines 6 and line 7.
+While the number of unique downregulated genes in both lines was approximately equal,
+the number of unique upregulated genes in line 7 was much greater compared
+to line 6 (Figure~\ref{degenes_venn})~\cite{lee1983ontogeny}.
 Interestingly, some genes that were differentially expressed in both lines were
-regulated in the opposite direction (table~\ref{tab:opposite}).
-Among genes downregulated in line 6 but upregulated in line 7,
-\textit{LL} (lung lectin) and \textit{SFTPA1}, which encode lectin proteins, are part of
-lectin-complement pathway.
+regulated in the opposite direction (Table~\ref{tab:opposite}).
+Among genes downregulated in line 6 but upregulated in line 7 were
+\textit{LL} (lung lectin) and \textit{SFTPA1}, which encode a calcium-dependent C-type
+lectin and a lung surfactant protein respectively.
+Both molecules are important in innate immunity~\cite{hogenkamp2008chicken,kingma2006defense}.
 \textit{LIMS1} is involved in cell differentiation and proliferation and
@@
 On the other hand, nearly all genes upregulated in line 6 but downregulated in line 7
-are involved in cell survival such as mRNAs splicing, cell growth, and protein systhesis,
+are involved in cell survival such as mRNAs splicing, cell growth, and protein synthesis,
 except CD7 whose function is involved in T cell-B cell interaction.
 This difference suggests that even at this stage of infection in line 6, the lytic phase 
-could be controlled or subsided because the viruses are driven to undergo latent infection early.
-Therefore, genes important for recovery of cell damages are upregulated.
+could be subsided, therefore, only genes important for survival of cell are upregulated in line 6.
 In comparison, the lytic phase in line 7 may still continue and as a result, genes involved in
@@
 In addition, type I interferon (\textit{IFN-$\gamma$} and \textit{IFN-$\beta$}) as well as \textit{INF-$\alpha$3} were found to be highly
-upregulated in infected chickens in both line 6 and 7 (table~\ref{tab:cytokines}).
+upregulated in infected chickens in both lines 6 and 7 (Table~\ref{tab:cytokines}).
 However, expressions of genes encoding their corresponding receptors
@@
 
-Some DE genes have been reported from previous Microarray studies were also found to be
+Some DE genes have been reported from previous microarray studies were also found to be
 differentially expressed in this study.
 For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
-chicken with MHC allele B\textsuperscript{19} at 4 d.p.i~\cite{sarson2008transcriptional}.
+chicken with the MHC allele B\textsuperscript{19} at 4 d.p.i~\cite{sarson2008transcriptional}.
 It was also found to be down-regulated about 3-fold in line 7 (susceptible).
-Similary, \textit{GMZA} gene reported to be upregulated across genetically different chickens
+Similarly, \textit{GMZA} gene reported to be upregulated across genetically different chickens
 (B$^{19}$, B$^{21}$ alleles), was also found to be highly upregulated.
@@
 were also found to be highly upregulated in both line 6 and 7.
-However, our results show the similar expression pattern of \textit{IL6} and \textit{IL18}, which were
+However, our results show the similar expression patterns for \textit{IL6} and \textit{IL18}, which were
 only upregulated in line 7 in early stage of infection (3-5 d.p.i).
 
-In contrary, \textit{IL15} has been reported non differential-expressed between control and infected chickens in
+In contrast, \textit{IL15} has been reported non differential-expressed between control
+and infected chickens in
 line 6$_{1}$ and 7$_{2}$~\cite{kaiser2003differential}; however,
@@
 
-\subsection*{Functional analysis of differntial-expressed genes}
+\subsection*{Functional analysis of differential-expressed genes}
 
 To determine pathways that were perturbed during the infection, data were
-analysed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{young2010method}.
+analyzed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{young2010method}.
 Significantly perturbed pathways (FDR $< 0.1$) from both lines that involved in
-immune response include Toll-like receptor signaling pathway, cytokine-cytokine receptor interaction,
-intestinal immune network for IgA production and cell adhesion molecules (CAMs).
+immune response include TLR signaling pathway, cytokine-cytokine receptor interaction,
+intestinal immune network for IgA production, and cell adhesion molecules (CAMs)~\ref{line67_kegg}.
 Some other pathways important in response to viral infection and only significantly
 enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway,
-NOD-like receptor signaling pathway and lysosome.
-Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed genes in phagosome pathway.
+NOD-like receptor signaling pathway, and lysosome.
+Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed
+genes in the phagosome pathway.
 Although MHC I gene (\textit{BF1}) was differentially expressed in both lines, other genes involved in
@@
 suggesting that new MHC I molecules were actively produced and expressed on cell surface.
+% (the minor class I is upregulated but the major class I is downregulated)
 Furthermore, gene ontology analysis of biological processes (GO:BP) shows that categories involved in
@@
 At this stage of infection, our results suggest that lytic infection of MDV stimulates
-both innate and the adaptive immune response, which leads to activation of T cells in line 7.
+both innate and adaptive immune responses, which leads to activation of T cells in line 7.
 Only activated T cells are infected by MDV, therefore, the lytic phase could facilitate the spread of
 the viruses by enhancing expansion of activated T cells.
-In contrast, the immune response in line 6 could be able to drive the viruses to latency phase early
-and only activate the adaptive immune response in a very short duration.
-Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell contact between B cells and
+Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell
+contact between B cells and
 T cells during antigen presentation or B cells activation by T helper cells.
@@
 
-Immune system is isoform-rich and many genes express different isoforms with distinctive functions in response to
-stimuli such as stress, chemicals and infection.
-Changes in expression of splice forms of immune related genes have been reported to be associated with increase
-susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
+The immune system is isoform-rich and many genes express different isoforms with
+distinctive functions in response to stimuli such as stress, chemicals and infection.
+Changes in expression of splice forms of immune related genes have been reported to
+be associated with increase susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
 Studying differential isoform expression could therefore shed lights into inherent differences between 
-line 6 and 7 that confer resistance or susceptibility to MD.
+lines 6 and 7 that confer resistance or susceptibility to MD.
 
-In the past decades, microarray technology has been used to study gene and isoform expressions in many studies, but
-its sensitiviy for detection of structurally similar isoforms is low and known or predicted annotations
-are required to design probes~\cite{kane2000assessment}.
-Although RNA-Seq method can provide a reliable estimate of an exon expression compared to microarray~\cite{pan2008deep}
-and is not constrained to those limitations, studying isoform
-expressions using RNA-Seq is still not straightforward because of short read lengths.
+In the past decades, microarray technology has been used to study gene and isoform
+expressions in many studies, but its sensitivity for detection of structurally similar
+isoforms is low, and known or predicted annotations are required to design probes~\cite{kane2000assessment}.
+Although RNA-Seq method can provide a reliable estimate of an exon expression compared
+to microarray~\cite{pan2008deep} and is not constrained to those limitations, studying isoform
+expressions using RNA-Seq is still not straightforward because of the short read lengths.
 Reads from current Illumina technology are generally not long enough to span across all exons in an isoform.
@@
 
-Due to those issues, it is not feasible to accurately estimate exression of isoforms, especially when
+Due to those issues, it is not feasible to accurately estimate expression of isoforms, especially when
 gene annotation is constructed from \textit{de novo} assembly~\cite{trapnell2013differential}.
@@
 involved in a splicing event are examined.
-The expression of exon inclusion is calculated as percent spliced in (Psi or $\Psi$), which can be used to
+The expression of exon inclusion is calculated as Percent Spliced In (Psi or $\Psi$), which can be used to
 infer the portion of transcripts that include the exon in each sample~\cite{Katz:2010iv}.
-In this study, we investigated the three most common alternative splicing events in vertebrates, which are skipped
-exons (SE), an alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
+In this study, we investigated the three most common alternative splicing events in
+vertebrates, which are skipped exons (SE), an alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
 
@@
 
-List of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
-in infected chickens is shown in table~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
+Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
+in infected chickens are shown in
+Tables~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
 and~\ref{tab:line67i_diff_line67u_three}.
@@
 
-Group I includes genes that $\Psi$s were up or down regulated in infected chickens in line 6 only.
-This group include \textit{BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
+Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes with $\Psi$s that were
+up- or down-regulated in infected chickens in line 6 only.
+This group includes \textit{BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
 associated C2H2-type zinc finger protein encoding gene, which functions as a tumour-suppressor
@@
 two adjacent C2H2-type finger protein domains.
-Another gene related to B and T cell lymphoma is \textit{SIK2} (salt-inducible kinase 2).
-This gene has been reported to have a negative effect on T cell lymphoma by limiting the transcription of
+Another gene related to B and T cell lymphomas is \textit{SIK2} (salt-inducible kinase 2).
+This gene has been reported to have a negative effect on T cell lymphomas by limiting the transcription of
 HTLV-1 virus~\cite{tang2013lkb1}.
-Some other genes are important in pre-mRNA splicing including \textit{TRA2A}, \textit{SRSF6} and \textit{GEMIN6}.
+Some other genes are important in pre-mRNA splicing including \textit{TRA2A},
+\textit{SRSF6} and \textit{GEMIN6}.
 \textit{TRA2A} (transformer-2 protein homolog alpha) encodes RNA recognition motif (RRM),
 the skipped exon is not found to encode any known protein domain.
-\textit{SRSF6} (SR splicing factor 6) encodes a nuclear protein that belongs to the splicing factor protein family.
+\textit{SRSF6} (SR splicing factor 6) encodes a nuclear protein that belongs to the splicing
+factor protein family.
 \textit{GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
@@
 This gene encodes small GTPases, belonging to Ras family, that regulates a wide variety of cellular events
-including cell growth, cytoskeletal reorganization and the activation of protein kinases.
+including cell growth, cytoskeletal reorganization, and the activation of protein kinases.
 The role of small GTPases in immune responses is discussed further below.
 
-In group II, $\Psi$ values were relatively stable in control and infected chickens in line 6 and line 7.
+In group II (Table~\ref{tab:line67i_diff_line67u_two}),
+ $\Psi$ values were relatively stable in control and infected chickens within line, but not between lines.
 Genes that could play an important role in immune responses are \textit{ITGB2} and \textit{HCK}.
+\textit{HCK} transmits signal from cell surface receptors such as \textit{FCGR1A, FCGR2A, IL2, IL6,
+IL18}, and integrins (\textit{ITGB1, ITGB2}).
 \textit{ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of \textit{LFA-1} and \textit{CR3} receptors.
@@
 \textit{CR3} binds to a vast array of ligands and molecules including complement C3bi, microbial proteins,
-ICAM-1 and -2, ECM proteins and coagulation proteins~\cite{}.
+ICAM-1 and -2, ECM proteins, and coagulation proteins~\cite{}.
 It plays a significant role in neutrophils and monocytes activation including phagocytosis, adhesion and
 migration~\cite{}.
-Mutations in CD18 gene has been associated with type 1 leukocyte adhesion deficiency (LAD-1),
-an autosomal-recessive inherited disease found in few families.
-The disease is characterized by impair of lymphocytes in adherent-dependent functions, lack of accumulation
-to the site of infection and recurrent bacterial and fungal infection~\cite{}.
-\textit{HCK} transmits signal from cell surface receptors such as \textit{FCGR1A, FCGR2A, IL2, IL6,
-IL18} and integrins (\textit{ITGB1, ITGB2}).
-Based on our gene model, the skipped exon of \textit{ITGB2} is located in the 5$\prime$ UTR;
-whereas, the skipped exon of \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).
-Some other genes in this group are involved in cell rearrangement and cytokinesis that could be
-regulated by signals from integrins including \textit{DYLNT1,DYNLL2, SEPT11, PFN2} and \textit{ZDHHC7}.
-The role of \textit{ITGB2} in innate immune responses will be discuss further in the next section.
+The role of \textit{ITGB2} in immune responses is discussed further in the next section.
+\textit{DYNLT1,DYNLL2, SEPT11, PFN2}, and \textit{ZDHHC7} are also involved in cell
+rearrangement and cytokinesis.
+In particular, \textit{DYNLT1} and \textit{DYNLL2} are dynein proteins that have been demonstrated to
+regulate T cell activation by driving T cell receptor microclusters (TCR-MCs)
+toward the center of immune synapse~\cite{hashimoto2011dynein}.
 
-Group III includes genes that exhibit differential isoform expression only in infected chickens in line 7.
-A number of genes in this group encode proteins that are parts of spliceosome: \textit{SRSF3}, \textit{HNRNPDL},
+Group III (Table~\ref{tab:line67i_diff_line67u_three})
+includes genes that exhibit differential isoform expression only in infected line 7.
+A number of genes in this group encode proteins that are parts of
+spliceosome: \textit{SRSF3}, \textit{HNRNPDL},
 \textit{SFSWAP}, \textit{THOC1}, \textit{RNPC3} and \textit{SRSF5}.
 Two genes involved in cell-cell contact regulated by integrins are also in this group.
-\textit{PPP1R12A} is a myosin phosphatase that regulates the interaction of actin and myosin downstream of GTPase
-Rho proteins.
-\textit{PODXL} encodes protein that functions in integrin-dependent manner to increase cell-cell contact.
-Connections of these two genes and integrins will be discussed in the next section.
+\textit{PPP1R12A} is a myosin phosphatase that regulates the interaction of actin
+and myosin downstream of GTPase Rho proteins.
+\textit{PODXL} encodes PODX-like protein that functions in integrin-dependent
+manner as both pro-adhesive and anti-adhesive molecules.
+This protein is involved in cell-to-cell contact, cell trafficking, and cancer progression~\cite{}.
 
-The last group only has one gene, \textit{GOSR1}.
+The last group (Group IV, Table~\ref{tab:line67i_diff_line67u_three})
+only has one gene, \textit{GOSR1}.
 The $\Psi$ value of this gene was less than 0.20 cutoff in control and infected chickens in line 6, but it is
@@
 Also, there is a significant difference between control and infected chickens in line 7.
-This gene encodes traffiking membrane protein important for transporting proteins from \textit{cis-} to
-\textit{trans-}golgi network.
+This gene encodes a trafficking membrane protein important for transporting proteins
+from \textit{cis-} to \textit{trans-}golgi network.
 
-\subsubsection*{Integrins and actin cytoskeleton pathway}
+\subsubsection*{Roles of \textit{LFA-1} and actin cytoskeleton pathway in T cells activation}
 
-Integrins are a special kind of receptors that transmit signals bidirectionally across cell membrane.
-They are heterodimeric composed of an $\alpha$ (large) and a $\beta$ (small) subunits.
-$\beta_{2}$ (CD18) subunit is exclusively expressed on lymphocytes and APCs as a component of \textit{LFA-1}
-and \textit{CR3} receptors.
+Integrins are a special kind of receptors that transmit signals bidirectionally across the cell membrane.
+They are heterodimeric composed of an $\alpha$ (large) and a $\beta$ (small)
+subunits~\cite{wang2010immunopathologies}.
+$\beta_{2}$ (CD18) subunit encoded by \textit{ITGB2} is expressed on lymphocytes and antigen presenting cells
+(APCs) as a component of \textit{LFA-1} and \textit{CR3} receptors.
+LFA-1 binds to its ligand ICAM-1 to help form a synapse that bring APC and T cell together to
+initiate antigen presentation leading to T cell activation~\cite{dustin2000immunological}.
+Absence of LFA-1 leads to impaired functions of lymphocytes in proliferation and
+tumor rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.
+Mutations in \textit{ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency (LAD-1),
+an autosomal-recessive inherited disease found in few families.
+The disease is characterized by impair of lymphocytes in adherent-dependent functions, lack of accumulation
+to the site of infection and recurrent bacterial and fungal infection~\cite{springer1987lymphocyte}.
 
-(more on \textit{LFA-1} and \textit{CR3} to be added).
-
-As discussed above, \textit{RAC3, HCK, DYLNT1, DYNLL2, SEPT11, PFN2, ZDHHC7, PODXL} and \textit{PPP1R12A}
-could be activated by the integrins because of their roles in cell rearrangement or molecular interaction.
-Interestingly, we found that in human, \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A} are in the regulation
-of actin cytoskeleton pathway (figure~\ref{kegg_actin}).
-Moreover, at least two of these genes are co-present in other three pathways (table~\ref{tab:integrin}) that
-involved in immune responses.
-This strongly suggests that these four genes could contribute to difference in immune responses between
-line 6 and 7 even though their expressions were not significantly different between control and infected groups.
-
-% \subsubsection*{Splicing factors}
-% 
-% (Not sure if I should write about splicing factors, I'm afraid it will make the paper too long and unfocused.)
-% 
-% The second group includes genes encode protein compoments of spliceosome or are involved in pre-mRNA splicing.
-% Those genes are \textit{TRA2A, SRSF6, GEMIN6, SRSF3, HNRNPDL, SFSWAP, THOC1, THOC2, RNPC3} and \textit{SRSF5}.
-% Even though alternative splicing is regulated by splicesome, it is difficult to elucidate the effect of alternative
-% splice forms of spliceosome subunits on alternative splicing regulation.
-% This is in part due to the fact that the spliceosome is the most complex protein structures composed of many different
-% subunits each of which could have isoforms differing in function or activity~\cite{}.
+Actin cytoskeleton is very important in T cell activation because it enhances activity of
+LFA-1 by increasing its avidity and recruiting signaling molecules necessary for
+downstream signaling~\cite{dustin2000immunological, van2000avidity}.
+Cytoskeleton proteins binding to cytoplasmic domain of LFA-1 are thought to play an important role in
+driving LFA-1 to aggregation on the cell surface, resulting in increased avidity.
+Aggregation of LFA-1 has been demonstrated to be essential for lymphocytes to bind to the
+ligand~\cite{van1994extracellular}.
+Interestingly, \textit{RAC3}, \textit{PFN2}, and \textit{PPP1R12A}, which are involved in
+actin cytoskeleton pathway (Figure~\ref{kegg_actin}), also expressed different ratios of alternative splice
+forms between lines 6 and 7.
+Some of these genes are also co-present in three other pathways
+that involved in immune responses (Table~\ref{tab:integrin}).
+This could be speculated that pre-mRNA splicing of these genes is co-regulated by splicing
+regulators or some genetic factors.
 
 \subsubsection*{Protein domains}
-To understand the role of the alternative splice forms on the function of these genes, transcripts were
-translated to protein sequences by ESTscan~\cite{}.
-Protein sequences were then searched for annotated protein domains using InterProSearch~\cite{}.
-Besides \textit{ITGB2}, other genes have alternative exons located in coding regions and could potentially affect
-functional protein domains in some ways.
-The exon with alternative 3$\prime$ splice site of \textit{RAC3} encode part of a protein domain identified as small GTPase of Ras subfamily
-(ProSiteProfiles:PS5142 and SMART:SM00173).
-Rac3 is highly homologous to Rac1 and has been reported to possess the ability of promoting membrane ruffling, transformation, activation of
+To understand the role of the alternative splice forms on the function of these genes,
+transcripts were translated to protein sequences by ESTscan~\cite{iseli1999estscan}.
+Protein sequences were then searched for annotated protein domains using
+the standalone version of InterPro Scan~\cite{quevillon2005interproscan}.
+Besides \textit{ITGB2}, other genes have alternative exons located in coding regions and
+could potentially affect functional protein domains in some ways.
+The exon with alternative 3$\prime$ splice site of \textit{RAC3} encode part of a protein
+domain identified as small GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).
+Rac3 is highly homologous to Rac1 and has been reported to possess the ability of
+promoting membrane ruffling, transformation, activation of
 c-Jun transcriptional activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.
@@
 
-The alternative exon of \textit{PFN2} seems to disrupt the coding sequence that encode profilin domain (Pfam:PF00235).
-The profilin domain is essential for almost all orgamisms and its functions include regulating actin polymerization, controlling
-complex network of molecular interaction and transmitting signals from small-GTPase pathway.
+The alternative exon of \textit{PFN2} seems to disrupt the coding sequence that
+encode profilin domain (Pfam:PF00235).
+The profilin domain is essential for almost all organisms and its functions include
+regulating actin polymerization, controlling complex network of molecular interaction
+and transmitting signals from small-GTPase pathway.
 It also binds to Rac effector molecules and a number of other ligands~\cite{witke2004role}.
-The skipped exon of \textit{PPP1R12A} encode part of a protein domain annotated as protein phosphatase 1 (PIRSF:PIRSF038141).
+The skipped exon of \textit{PPP1R12A} encodes part of a protein domain annotated as protein
+phosphatase 1 (PIRSF:PIRSF038141).
+The functional effects of alternative splicing on these functional domains have to be
+further investigated by simulation or an experiment.
+
+% Based on our gene model, the skipped exon of \textit{ITGB2} is located in the 5$\prime$ UTR;
+% whereas, the skipped exon of \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).
 
@@
 
-Although alternative splicing of genes in all groups could be regulated by \textit{cis-} and \textit{trans-}splicing factors,
+Although alternative splicing of genes in all groups could be regulated by
+\textit{cis-} and \textit{trans-}splicing factors,
 genes in group II are more likely to be regulated by \textit{cis-}regulatory factors.
-The ratios of isoform expression in this group were relatively stable within line, but were significantly different between lines.
-Investigation of nucleotide differences within exons of line 6 and 7 could reveal a possible role of cSNPs in regulating
-alternative splicing in this groups.
-We obtained a sequence of an alternative exon from line 6 and used Human Splicing Finder (HSF) to determine whether SNPs from
-line 7 could alter predicted ESEs or ESSs.
-SNPs in exons of \textit{ITGB2} and \textit{PFN2} between line 6 and 7 are listed in table~\ref{tab:deu_snps}.
+The ratios of isoform expression in this group were relatively stable within line,
+but were significantly different between lines.
+Investigation of nucleotide differences within exons of line 6 and 7 could reveal
+a possible role of cSNPs in regulating alternative splicing in this group.
+We obtained a sequence of an alternative exon from line 6 and used Human Splicing Finder (HSF)
+to determine whether SNPs from line 7 could alter predicted ESEs or ESSs.
+Results from some genes involved in cytokinesis are discussed in this section.
+Exonic SNPs from line 6 and 7 are listed in Table~\ref{tab:deu_snps}.
 
-For \textit{ITGB2}, a putative ESE motif of SC35 is predicted to be broken when replace a T with a C at position 26 of the cDNA, which
-corresponds to a position 7,183,696 on chromosome 7.
-The disruption of this exon enhancer binding site support the fact that the level of exon inclusion in line 7 is
-practically zero (figure~\ref{itgb2}).
-On the other hand, there is also a putative binding site for an exon suppressor that could promote exclusion of the exon in line 7.
-The SNP is also predicted to affect putative ESE motifs of other splicing factors, however, those motifs do not seem to agree with
-exon expression data.
+For \textit{ITGB2}, a SNP (T) at position 26 of the cDNA from line 6, which corresponds to
+position 7,183,696 on chromosome 7 is located in a predicted binding site for
+SC35, which is an exon enhancer.
+Although the exon is not expressed in line 7, we found that there is no polymorphisms between
+lines 6 and 7 according to SNPs data from the genome resequencing.
+Therefore, this SNP may not be accounted for exclusion of the exon in
+line 7 (Figure~\ref{itgb2}).
 
-Exon sequences of \textit{PFN2} from line 6 and 7 differ at a position 2,072 on cDNA or 23,221,934 on chromsome 9.
-A small insertion of two AA nucleotides is predicted to create a new ESE motif for Tra2-$\beta$ splicing regulator,
-whose function as a stabilizer of an enhancer complex~\cite{lopez1998alternative}.
-From exon expreesion data from, $\Psi$ of an exon with alternative splicing increases from 0.20-0.30 to about 0.50 in line 7.
-Possibly, Tra2 increases inclusion of the exon with alternative 3$\prime$ splice site via
-ESE-dependent 3$\prime$ splice site activation (figure~ref{pfn2}).
-Even though the linear distance between Tra2-$\beta$ binding site and the alternative 3$\prime$ splice site is
-greater than 1kb, Tra2 could bind close to the 3$\prime$ splice site if the secondary structure of the mRNA moves the
-ESE motif closer to the splice site.
+Exon sequences of \textit{PFN2} from line 6 and 7 differ at position 2,072 on
+cDNA or 23,221,934 on chromosome 9.
+A small insertion of two AA nucleotides is predicted to create a new binding site
+for Tra2-$\beta$ splicing regulator, which serves as a stabilizer of an enhancer
+complex~\cite{lopez1998alternative}.
+From exon expression data, $\Psi$ of an exon with alternative splicing increases
+from 0.20-0.30 to about 0.50 in line 7.
+Tra2 could possibly increase inclusion of the exon with alternative 3$\prime$ splice site via
+ESE-dependent 3$\prime$ splice site activation (Figure~\ref{pfn2}).
+Even though the linear distance between Tra2-$\beta$ binding site and the alternative
+3$\prime$ splice site is greater than 1kb, Tra2 could possibly get close to
+the 3$\prime$ splice site in the secondary structure of the mRNA.
+
+Replacement of A with G nucleotide in the skipped exon of \textit{DYNLL2} from line 6
+is predicted to slightly alter the binding site of several ESEs as well as to create
+a new binding site for 9G8.
+This exon is upregulated in line 7 compared to line 6, therefore, the present
+of the new binding site for 9G8 exon enhancer helps support the expression results.
+In addition, A nucleotide is this position matches the reference
+nucleotide, therefore, we could expect this exon to be expressed in other datasets.
+According to EST tags on the UCSC genome browser, the exon has been found and sequenced
+from chicken eyes (15d post-hatched, EST sequence:DR424100).
+
+For \textit{DYNLT1}, the skipped exon was absent in line 7, therefore, we could not
+determine the SNPs from RNA-Seq data.
+However, data from DNA resequencing showed that there was a polymorphism at position
+51,357,865 on chromosome 3.
+Replacing G with C from line 7 is predicted to slightly alter predicted binding sites
+for 9G8 as well as SRp55 exon enhancer.
+However, whether the alteration down regulates the exon in line 7 is unclear.
+
+There are too many SNPs in the exon of \textit{SEPT11} and \textit{ZDHHC7}, making it
+unfeasible to predict which SNP regulates the exon expression.
+Therefore, we do not discuss these two genes in this section.
+However, results from HSF analysis of these two genes and
+other genes in this group are provided in the supplementary materials.
 
 \section*{Discussion}
+
+Custom gene models built from combination of gene models from \textit{de novo} assembly,
+reference-based assembly, and Ensembl allowed us to identify genes and isoforms
+that might play an important role in resistance to MD.
+Results from gene expression analysis indicate that adaptive immune responses were active
+during lytic infection in line 7.
+Because only activated T cells are infected, elicitation of adaptive immune responses could
+help spread the viruses by recruiting and activating more T cells.
+Therefore, active immune responses are favorable for the virus.
+On the other hand, the delay of adaptive immune responses in line 6, in this case,
+could benefit the host by limiting infection of activated T cells.
+
+Using exon-centric approach, we identified several genes involved in
+actin cytoskeleton, especially \textit{ITGB2},
+that expressed different ratios of alternative isoforms between lines 6 and 7.
+As described above, \textit{ITGB2} encodes $\beta$-subunit of integrins including LFA-1, which exclusively
+expressed in lymphocytes and plays a major role in lymphoproliferation,
+antigen presentation, T cells activation, and cytotoxicity.
+In patients with LFA-1 deficiency (LAD), the response of lymphocytes to mitogens is
+decreased~\cite{springer1987lymphocyte}.
+The decrease in responsiveness to mitogens have been shown to correlate with resistance to MD
+by Lee and Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds
+(line 6 and N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line 7 and P).
+Even though the exact mechanism is not known, incompetency of T cells in response to
+stimuli appears to benefit resistant birds because MD virus could not induce T cells
+to proliferate and cause them to undergo neoplastic transformation.
+It has also been suggested that the mechanism that controls
+both lymphocyte proliferation induced by MD virus and lymphocyte proliferation induced by
+immune response is the same~\cite{pazderka1975histocompatibility}.
+Therefore, it may be useful to consider a link between deficiency of lymphocytes in line 6 to
+the alternative splice form of \textit{ITGB2} that is only expressed in line 6.
+Although the exon included in the alternative splice form is non-coding, it could serve important functions in
+translation or posttranscriptional regulation.
+
+Interestingly, by grouping genes based on patterns of $\Psi$s, we found that many genes in group
+II: \textit{ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and \textit{ZDHHC7}, are involved in cytokinesis or
+cell synapse, which are important for aggregation of LFA-1 and T cell activation as discussed above.
+The $\Psi$ values in group II are more similar within line than between lines in both control and
+infected groups, suggesting the possibility that the alternative splicing is
+controlled by a genetic mechanism.
+Based on the independent criteria, enrichment of cytoskeletal genes in this group could not be
+just a mere coincidence, but an evidence suggesting that inefficiency of T cell activation
+in line 6 is linked to a genetic factor that regulate alternative splicing of \textit{ITGB2}
+and other genes in the cytoskeleton system.
+Experimental validation of exonic SNPs provided by this study could shed some lights on
+underlying polymorphisms that contribute to MD resistance.
+
+To conclude, we attempted to employ custom gene models built from RNA-Seq data to
+identify underlying genetic factors contributing to MD resistance.
+We identified a number of genes that expressed different ratios of alternative splice forms
+between lines 6 and 7.
+Interestingly, several genes are involved in actin cytoskeleton and cytokinesis, which have
+not been of great interest in the filed of MD research.
+Even though we mainly discussed the possible role of \textit{ITGB2} in MD resistance, other
+genes cannot be precluded and should be a candidate for further investigation and experimental studies.
+Moreover, a full mechanism of MD resistance may be highly complex and require more data from different
+stages of infection as well as a greater sequencing depth in order to identify all
+genes and isoforms involved.
+To enable the study of unannotated gene and isoform expression,
+our approach of constructing gene models from RNA-Seq should be iteratively used to extend
+the Ensembl gene models.
+
+% However, to mechanism that contribute to resistance to MD could be very complex and integration
+% of global gene expression data as well as experimental data from different stages of infection will be
+% required to fully elucidate.
+% Moreover, better gene models and a reference genome will be necessary.
 
@@
 A5SS & chr2:12775.ev1 & ENSGALG00000011488 & \textbf{CMTM7} & \textbf{0.55} & \textbf{0.67} & 0.37 & 0.41 \\
-A3SS & chr8:25257.ev1 & ENSGALG00000008939 & \textbf{FUBP1} & \textbf{0.58} & \textbf{0.74} & 0.41 & 0.46 \\
+A3SS & chr8:25257.ev1 & ENSGALG00000008939 & FUBP1 & \textbf{0.58} & \textbf{0.74} & 0.41 & 0.46 \\
 A3SS & chr9:25750.ev1 & ENSGALG00000010410 & \textbf{PFN2} & \textbf{0.71} & \textbf{0.78} & 0.53 & 0.50 \\
@@
 A3SS & chr11:8703.ev1 & ENSGALG00000020987 & \textbf{ZDHHC7} & \textbf{0.42} & \textbf{0.23} & 0.57 & 0.55 \\
-A3SS & chr4:21136.ev1 & ENSGALG00000027908 & \textbf{LOC42228} & \textbf{0.72} & \textbf{0.61} & 0.87 & 0.91 \\
+A3SS & chr4:21136.ev1 & ENSGALG00000027908 & \textbf{LOC422528} & \textbf{0.72} & \textbf{0.61} & 0.87 & 0.91 \\
 \hline
@@
 
-% \begin{table}[!ht]
-% \caption{
-% \bf{DEU Genes in Spliceosome pathway}}
-% \begin{tabular}{cccccc}
-% \hline
-% Gene ID &  Ensembl & Symbol  & $-log2$FC* & \\
-%         & & & Line 6 & Line 7 \\
-% \hline
-% chr26:16840 & ENSGALG00000000533 & SRSF3 & 0.43 & 0.18 \\
-% chr2:13147 & ENSGALG00000014915 & THOC1 & -- & 0.21 \\
-% chr5:21970 & ENSGALG00000009421 & SRSF5 & -- & 0.18 \\
-% chr2:13729 & ENSGALG00000010973 & TRA2A & -- & -- \\
-% chr4:20185 & ENSGALG00000008507 & THOC2 & -- & -- \\
-% chr7:24350 & ENSGALG00000008038 & SF3B1 & -- & -- \\
-% \hline
-% \end{tabular}
-% \begin{flushleft}
-%     * FDR $<$ 0.05
-% \end{flushleft}
-% \label{tab:spliceosome}
-% \end{table}
-
 \begin{table}[!ht]
@@
 \hline
-ITGB2 & 7 & 7183696 & C & \textit{T} & C & - \\
-PFN2 & 9 & 23221934 & -  - & - - & \textit{AA} & + \\
+ITGB2 & 7 & 7183696 & C & \textit{T} & . & - \\
+PFN2 & 9 & 23221934 & -  - & . & \textit{AA} & + \\
+DYNLT1 & 3 & 51357865 & C & . & \textit{G} & - \\
+DYNLL2 & 19 & 8694148 & G & . & \textit{A} & - \\
 \hline
@@
 \hline
-ITGB2 & 20 & SC35 & ESE$^{1}$ & TGCTCATG (78.19) & & Site broken \\
- & 22 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (91.15) & +18.03\% \\
- & 22 & SF2/ASF(IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (89.23) & +15.53\% \\
- & 22 & SF2/ASF, SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (91.15) & +22.27\% \\
- & 22 & SF2/ASF, SF2/ASF & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (89.23) & +19.69\% \\
- & 24 & SRp55 & ESE$^{1}$ & & CACGGA (79.30) & New site \\
- & 26 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & & CGGAGAT (80.00) & New site \\
- & 26 & SF2/ASF & ESE$^{1}$ & & CGGAGAT (75.36) & New site \\
- & 21 & & ESS$^{3}$ & & GCTCACGG (63.35) & New site (-5.59) \\
- & 23 & & ESS$^{3}$ & TCATGGAG (61.41) & & Site broken (3.16) \\
- & 23 & & ESS$^{3}$ & ATGGAGAT (64.83) & ACGGAGAT & -5.16\% \\
- & 24 & & ESS$^{4}$ & CATGGA (65.48) & CACGGA (65.48) & 0\% \\
+% ITGB2 & 20 & SC35 & ESE$^{1}$ & TGCTCATG (78.19) & & Site broken \\
+%  & 22 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (91.15) & +18.03\% \\
+%  & 22 & SF2/ASF(IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CTCATGG (77.23) & CTCACGG (89.23) & +15.53\% \\
+%  & 22 & SF2/ASF, SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (91.15) & +22.27\% \\
+%  & 22 & SF2/ASF, SF2/ASF & ESE$^{1}$ & CTCATGG (74.55) & CTCACGG (89.23) & +19.69\% \\
+%  & 24 & SRp55 & ESE$^{1}$ & & CACGGA (79.30) & New site \\
+%  & 26 & SF2/ASF(IgM-BRCA1) & ESE$^{1}$ & & CGGAGAT (80.00) & New site \\
+%  & 26 & SF2/ASF & ESE$^{1}$ & & CGGAGAT (75.36) & New site \\
+%  & 21 & & ESS$^{3}$ & & GCTCACGG (63.35) & New site (-5.59) \\
+%  & 23 & & ESS$^{3}$ & TCATGGAG (61.41) & & Site broken (3.16) \\
+%  & 23 & & ESS$^{3}$ & ATGGAGAT (64.83) & ACGGAGAT & -5.16\% \\
+%  & 24 & & ESS$^{4}$ & CATGGA (65.48) & CACGGA (65.48) & 0\% \\
+DYNLT1 & 59 & SRp55, SRp40 & ESE$^{1}$& TGAATC (78.47) & TGAATGC (79.10) & +0.81\% \\
+& 60 & 9G8 & ESE$^{2}$ & GAATCC (60.07) & GAATGC (63.49) & +5.70\% \\
+\hline
+DYNLL2 & 2 & SF2/ASF (IgM-BRCA1) & ESE$^{1}$ & CTCCGGG (86.38) & CTCCGAG (72.69) & -15.85\% \\
+& 2 & SF2/ASF, SF2/ASF (IgM-BRCA1) & ESE$^{1}$ & CTCCGGG (79.91) & CTCCGAG (72.69) & -9.03\% \\
+& 4 & SF2/ASF (IgM-BRCA1) & ESE$^{1}$& CCGGGGT (73.00) & CCGAGGT (86.23) & 18.12\% \\
+& 4 & SF2/ASF (IgM-BRCA1), SF2/ASF & ESE$^{1}$ & CCGGGGT (73.00) & CCGAGGT (82.94) & 13.61\% \\
+& 6 & 9G8 & ESE$^{2}$ & & GAGGTG (60.67) & New site \\
+& 6 & hnRNP A1 & ESS$^{4}$& & GAGGTG (74.05) & New site \\
 \hline
@@
 & 2070 & Tra2-$\beta$ & ESE$^{1}$ & & AAAaaT (81.02) & New site \\
- & 2066 & & ESS$^{2}$ & & ACAAAAaa (38.13) & New site \\
- & 2067 & & ESS$^{2}$ & & CAAAAaaT (28.85) & New site \\
+ & 2066 & & ESS$^{3}$ & & ACAAAAaa (38.13) & New site \\
+ & 2067 & & ESS$^{3}$ & & CAAAAaaT (28.85) & New site \\
 \hline
@@
     $^{1}$ESE Finder matrices for SRp40, SC35, SF2/ASF and SRp55 proteins.
-    $^{2}$Predicted PESS Octamers from Zhang \& Chasin.
-    $^{3}$Silencer motif from Sironi et al.
+    $^{2}$ESE motifs from HSF.
+    $^{3}$Predicted PESS Octamers from Zhang \& Chasin.
     $^{4}$hnRNP motif.

### 2014-07-15 paper.tex
updated format

@@
 Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
-Understanding the genetic basis for non-MHC resistance would be of both fundamental and applied importance.
-In this study, using computational approaches of RNA-Seq data, we identified
-differentially expressed genes and isoforms in response to MDV infection
-from highly inbred chickens lines that are MD resistant and susceptible,
-and performed pathway and functional analysis.
-Results show that expression of genes and isoforms involved in cell cytoskeleton and
-adhesion molecules in line 6 may contribute to resistance of the disease by limiting
-T cell activation, which in turn prevents the spread of MDV to activated T cells.
-In addition, we identified exonic single nucleotide polymorphisms (SNPs) between
-lines 6 and 7 predicted their role in regulation of alternative splicing.
-Based on our results, we speculate that alternative splice forms of cell adhesion molecules
-and cytoskeleton play and important role in controlling T cell activation in line 6.
+Understanding the genetic basis for non-MHC resistance would be of both
+fundamental and applied importance.  In this study, using computational
+approaches of RNA-Seq data, we identified differentially expressed genes and
+isoforms in response to MDV infection from highly inbred chickens lines that are
+MD resistant and susceptible, and performed pathway and functional analysis.
+Results show that expression of genes and isoforms involved in cell cytoskeleton
+and adhesion molecules in line 6 may contribute to resistance of the disease by
+limiting T cell activation, which in turn prevents the spread of MDV to
+activated T cells.  In addition, we identified exonic single nucleotide
+polymorphisms (SNPs) between lines 6 and 7 predicted their role in regulation of
+alternative splicing.  Based on our results, we speculate that alternative
+splice forms of cell adhesion molecules and cytoskeleton play and important role
+in controlling T cell activation in line 6.
 
@@
 
-Marek's disease (MD) is an economically significant chicken disease that affects the
-poultry industry worldwide with estimated annual loss of
-\$2 billion loss has been reported from outbreaks~\cite{morrow2004marek}.
-The disease is caused by the highly oncogenic Marek's disease virus (MDV)
-an alphaherpesvirus that induces T-cell lymphomas in susceptible birds.
-Vaccination is the primary control measure, which is effective in reducing
-incidence of tumor formation.
+Marek's disease (MD) is an economically significant chicken disease that affects
+the poultry industry worldwide with estimated annual loss of \$2 billion loss
+has been reported from outbreaks~\cite{morrow2004marek}.  The disease is caused
+by the highly oncogenic Marek's disease virus (MDV), an alphaherpesvirus that
+induces T-cell lymphomas in susceptible birds.  Vaccination is the primary
+control measure, which is effective in reducing incidence of tumor formation.
 However, since MD vaccines are not sterilizing, they do not prevent infection or
-horizontal spread of the virus.
-As a consequence, MDV field strains that overcome vaccinal protection have arisen repeatedly over time.
-Therefore, there is a need for sustainable alternative controls measures, such as improving genetic
-resistance.
+horizontal spread of the virus.  As a consequence, MDV field strains that
+overcome vaccinal protection have arisen repeatedly over time.  Therefore, there
+is a need for sustainable alternative controls measures, such as improving
+genetic resistance.
 
-Many studies have reported strong associations between MHC alleles and resistance or susceptibility to
-MD. For example, chickens with MHC allele B$^{21}$ are highly resistant in contrast to chickens with
-B$^{16}$ allele which are highly susceptible.
-ADOL lines 6 and 7, both share (B$^6$) allele, yet exhibit different phenotypic responses to
-MD; e.g., challenge with JM/102W strain with result in 0 and 100\% MD incidence for lines 6 and 7,
-respectively.
-Thus, the major unanswered questions are what genetic factors, especially those that are non-MHC,
-contribute to susceptibility and resistance of the disease and what is the main contributing
-mechanism.
+Many studies have reported strong associations between MHC alleles and
+resistance or susceptibility to MD.  For example, chickens with MHC allele
+B$^{21}$ are highly resistant in contrast to chickens with B$^{16}$ allele which
+are highly susceptible.  ADOL lines 6 and 7, both share (B$^{2}$) allele, yet
+exhibit different phenotypic responses to MD; e.g., challenge with JM/102W
+strain with result in 0 and 100\% MD incidence for lines 6 and 7, respectively.
+Thus, the major unanswered questions are what genetic factors, especially those
+that are non-MHC, contribute to susceptibility and resistance of the disease and
+what is the main contributing mechanism.
 
-In the past decades, significant efforts have been spent to study variations in a global gene
-expression between resistant and susceptible birds using
-microarray and RNA-Seq methods in order to identify non-MHC genes that contribute
-to resistance to MD~\cite{}.
-However, none of the studies have investigated differential expression of alternative isoforms,
-which are known to play a significant role in many biological events including
-immune responses.
-In addition, studies have shown that isoform expression levels can provide better signatures for some
-diseases~\cite{zhang2013isoform}.
-Changes of isoform expression levels are governed partly by two types of \textit{cis-}regulatory
-elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer (ESS),
-which are located within an exon sequence.
-A number of sequence motifs of ESE and ESS have been identified in human and some other organisms
-and could be predicted \textit{in silico}.
-Mutations that disrupt or create those motifs could alter splicing
-patterns leading to aberrant alternative splicing.
-A number of disease-associated single-nucleotide polymorphsims in coding regions
-(cSNPs) that affect ESEs and ESSs have been well
-characterized~\cite{blencowe2000exonic, wang2007splicing}.
-%In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing via this mechanism.
+In the past decades, significant efforts have been spent to study variations in
+a global gene expression between resistant and susceptible birds using
+microarray and RNA-Seq methods in order to identify non-MHC genes that
+contribute to resistance to MD~\cite{}.  However, none of the studies have
+investigated differential expression of alternative isoforms, which are known to
+play a significant role in many biological events including immune responses.
+In addition, studies have shown that isoform expression levels can provide
+better signatures for some diseases~\cite{zhang2013isoform}.  Changes of isoform
+expression levels are governed partly by two types of {\em cis-}regulatory
+elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer (ESS), which
+are located within an exon sequence.  A number of sequence motifs of ESE and ESS
+have been identified in human and some other organisms and could be predicted
+{\em in silico}.  Mutations that disrupt or create those motifs could alter
+splicing patterns leading to aberrant alternative splicing.  A number of
+disease-associated single-nucleotide polymorphsims in coding regions (cSNPs)
+that affect ESEs and ESSs have been well characterized~\cite{blencowe2000exonic,
+wang2007splicing}.
+%In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing
+%via this mechanism.
 Therefore, variations in isoform expression could lead to identification of SNPs
-that underlie genetic resistance to MD.
-In this article, we reported differential-expressed genes and isoforms that may contribute to
-resistance to MD as well as cSNPs that can potentially affect isoform expression levels.
+that underlie genetic resistance to MD.  In this article, we reported
+differential-expressed genes and isoforms that may contribute to resistance to
+MD as well as cSNPs that can potentially affect isoform expression levels.
 
@@
 
-To extensively study gene and isoform expression, we incorporated \textit{Ensembl} gene models
-release 73 with \textit{de novo} and a reference-guided transcriptome assembly to build custom
-gene models.
-The models, therefore, include both \textit{Ensembl} annotated transcripts
-and putative transcripts and genes.
-The advantage of using custom gene models is it allows an investigation of unannotated genes and
-isoforms, which is necessary for in-depth transcriptome study of an immune system.
+To extensively study gene and isoform expression, we incorporated ENSEMBL gene
+models release 73 with {\em de novo} and a reference-guided transcriptome
+assembly to build custom gene models.  The models, therefore, include both
+ENSEMBL annotated transcripts and putative transcripts and genes.  The advantage
+of using custom gene models is it allows an investigation of unannotated genes
+and isoforms, which is necessary for in-depth transcriptome study of an immune
+system.
 
-Many genes were found to be differentially expressed (DE) between control and infected
-chickens in lines 6 and line 7.
-While the number of unique downregulated genes in both lines was approximately equal,
-the number of unique upregulated genes in line 7 was much greater compared
-to line 6 (Figure~\ref{degenes_venn})~\cite{lee1983ontogeny}.
-Interestingly, some genes that were differentially expressed in both lines were
-regulated in the opposite direction (Table~\ref{tab:opposite}).
-Among genes downregulated in line 6 but upregulated in line 7 were
-\textit{LL} (lung lectin) and \textit{SFTPA1}, which encode a calcium-dependent C-type
-lectin and a lung surfactant protein respectively.
-Both molecules are important in innate immunity~\cite{hogenkamp2008chicken,kingma2006defense}.
-\textit{LIMS1} is involved in cell differentiation and proliferation and
-\textit{PPARG} is a suppressor of \textit{NF$\kappa$B}-mediated proinflamatory response.
-On the other hand, nearly all genes upregulated in line 6 but downregulated in line 7
-are involved in cell survival such as mRNAs splicing, cell growth, and protein synthesis,
-except CD7 whose function is involved in T cell-B cell interaction.
-This difference suggests that even at this stage of infection in line 6, the lytic phase 
-could be subsided, therefore, only genes important for survival of cell are upregulated in line 6.
-In comparison, the lytic phase in line 7 may still continue and as a result, genes involved in
-immune responses are still upregulated.
-In addition, type I interferon (\textit{IFN-$\gamma$} and \textit{IFN-$\beta$}) as well as \textit{INF-$\alpha$3} were found to be highly
-upregulated in infected chickens in both lines 6 and 7 (Table~\ref{tab:cytokines}).
-However, expressions of genes encoding their corresponding receptors
-were not different in line 6, but upregulated in line 7.
-This could also reflect the ongoing immune responses in line 7.
+Many genes were found to be differentially expressed (DE) between control and
+infected chickens in lines 6 and line 7.  While the number of unique
+downregulated genes in both lines was approximately equal, the number of unique
+upregulated genes in line 7 was much greater compared to line 6
+(Figure~\ref{degenes_venn})~\cite{lee1983ontogeny}.  Interestingly, some genes
+that were differentially expressed in both lines were regulated in the opposite
+direction (Table~\ref{tab:opposite}).  Among genes downregulated in line 6 but
+upregulated in line 7 were {\em LL} (lung lectin) and {\em SFTPA1}, which encode
+a calcium-dependent C-type lectin and a lung surfactant protein respectively.
+Both molecules are important in innate
+immunity~\cite{hogenkamp2008chicken,kingma2006defense}.  {\em LIMS1} is involved
+in cell differentiation and proliferation and {\em PPARG} is a suppressor of
+{\em NF$\kappa$B}-mediated proinflamatory response.  On the other hand, nearly
+all genes upregulated in line 6 but downregulated in line 7 are involved in cell
+survival such as mRNAs splicing, cell growth, and protein synthesis, except CD7
+whose function is involved in T cell-B cell interaction.  This difference
+suggests that even at this stage of infection in line 6, the lytic phase could
+be subsided, therefore, only genes important for survival of cell are
+upregulated in line 6.  In comparison, the lytic phase in line 7 may still
+continue and as a result, genes involved in immune responses are still
+upregulated.  In addition, type I interferon ({\em IFN-$\gamma$} and {\em
+IFN-$\beta$}) as well as {\em INF-$\alpha$3} were found to be highly upregulated
+in infected chickens in both lines 6 and 7 (Table~\ref{tab:cytokines}).
+However, expressions of genes encoding their corresponding receptors were not
+different in line 6, but upregulated in line 7.  This could also reflect the
+ongoing immune responses in line 7.
 
-Some DE genes have been reported from previous microarray studies were also found to be
-differentially expressed in this study.
-For example, B6.1 (Bu-1) gene is known to be down-regulated approximately 2.27 fold in susceptible
-chicken with the MHC allele B\textsuperscript{19} at 4 d.p.i~\cite{sarson2008transcriptional}.
-It was also found to be down-regulated about 3-fold in line 7 (susceptible).
-Similarly, \textit{GMZA} gene reported to be upregulated across genetically different chickens
-(B$^{19}$, B$^{21}$ alleles), was also found to be highly upregulated.
-In contrast, some genes that have been reported to be highly expressed in resistant chickens were
-downregulated in both lines. Those genes are \textit{AMIGO2, MMP13} and \textit{CLEC3B}, which were
-found to be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
-% Several genes involved in innate immune response were differentially expressed in resistant chickens:
-% \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5 was upregulated.
-Other immune genes reported to be highly expressed in susceptible chickens (line 7$_{2}$) including
-\textit{AVD, ART1, NOS2, CXCL13L2, MX1} and \textit{SOCS1}~\cite{smith2011systems}
-were also found to be highly upregulated in both line 6 and 7.
-However, our results show the similar expression patterns for \textit{IL6} and \textit{IL18}, which were
-only upregulated in line 7 in early stage of infection (3-5 d.p.i).
+Some DE genes have been reported from previous microarray studies were also
+found to be differentially expressed in this study.  For example, B6.1 (Bu-1)
+gene is known to be down-regulated approximately 2.27 fold in susceptible
+chicken with the MHC allele B\textsuperscript{19} at 4
+d.p.i~\cite{sarson2008transcriptional}.  It was also found to be down-regulated
+about 3-fold in line 7 (susceptible).  Similarly, {\em GMZA} gene reported to be
+upregulated across genetically different chickens (B$^{19}$, B$^{21}$ alleles),
+was also found to be highly upregulated.  In contrast, some genes that have been
+reported to be highly expressed in resistant chickens were downregulated in both
+lines. Those genes are {\em AMIGO2, MMP13} and {\em CLEC3B}, which were found to
+be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
+% Several genes involved in innate immune response were differentially expressed
+% in resistant chickens: \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all
+% those genes, only HSP5 was upregulated.
+Other immune genes reported to be highly expressed in susceptible chickens (line
+7$_{2}$) including {\em AVD, ART1, NOS2, CXCL13L2, MX1} and {\em
+SOCS1}~\cite{smith2011systems} were also found to be highly upregulated in both
+line 6 and 7.  However, our results show the similar expression patterns for
+{\em IL6} and {\em IL18}, which were only upregulated in line 7 in early stage
+of infection (3-5 d.p.i).
 
-In contrast, \textit{IL15} has been reported non differential-expressed between control
-and infected chickens in
-line 6$_{1}$ and 7$_{2}$~\cite{kaiser2003differential}; however,
-it was only upregulated in line 7 in our study.
-Expression of \textit{IL15} is induced by \textit{TLR9}, which binds to non-methylated CpG
-residues present in genomes of many DNA viruses, including herpes simplex virus.
-This cytokine auto-regulates the expression of \textit{CD40}, which is
-a transmembrane receptor required for activation of macrophages by CD4 T cells.
-Consequently, \textit{CD40} was only upregulated in line 7 (data not shown).
+In contrast, {\em IL15} has been reported non differential-expressed between
+control and infected chickens in line 6$_{1}$ and
+7$_{2}$~\cite{kaiser2003differential}; however, it was only upregulated in line
+7 in our study.  Expression of {\em IL15} is induced by {\em TLR9}, which binds
+to non-methylated CpG residues present in genomes of many DNA viruses, including
+herpes simplex virus.  This cytokine auto-regulates the expression of {\em
+CD40}, which is a transmembrane receptor required for activation of macrophages
+by CD4 T cells.  Consequently, {\em CD40} was only upregulated in line 7 (data
+not shown).
 
@@
 To determine pathways that were perturbed during the infection, data were
-analyzed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq data~\cite{young2010method}.
-Significantly perturbed pathways (FDR $< 0.1$) from both lines that involved in
-immune response include TLR signaling pathway, cytokine-cytokine receptor interaction,
-intestinal immune network for IgA production, and cell adhesion molecules (CAMs)~\ref{line67_kegg}.
-Some other pathways important in response to viral infection and only significantly
-enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling pathway,
-NOD-like receptor signaling pathway, and lysosome.
+analyzed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq
+data~\cite{young2010method}.  Significantly perturbed pathways (FDR $< 0.1$)
+from both lines that involved in immune response include TLR signaling pathway,
+cytokine-cytokine receptor interaction, intestinal immune network for IgA
+production, and cell adhesion molecules (CAMs)~\ref{line67_kegg}.  Some other
+pathways important in response to viral infection and only significantly
+enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling
+pathway, NOD-like receptor signaling pathway, and lysosome.
 Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed
-genes in the phagosome pathway.
-Although MHC I gene (\textit{BF1}) was differentially expressed in both lines, other genes involved in
-expressing newly synthesized MHC I were only upregulated in line 7
-suggesting that new MHC I molecules were actively produced and expressed on cell surface.
+genes in the phagosome pathway.  Although MHC I gene ({\em BF1}) was
+differentially expressed in both lines, other genes involved in expressing newly
+synthesized MHC I were only upregulated in line 7 suggesting that new MHC I
+molecules were actively produced and expressed on cell surface.
 % (the minor class I is upregulated but the major class I is downregulated)
-Furthermore, gene ontology analysis of biological processes (GO:BP) shows that categories involved in
-both adaptive and innate immune responses were enriched in line 7 (supplementary material).
-On the other hand, only categories involved in innate immune responses were enriched in line 6.
-In addition, enrichment of apoptosis pathway in line 7 indicates that the programmed
-cell death could be induced by CTL response to eliminate the viruses.
+Furthermore, gene ontology analysis of biological processes (GO:BP) shows that
+categories involved in both adaptive and innate immune responses were enriched
+in line 7 (supplementary material).  On the other hand, only categories involved
+in innate immune responses were enriched in line 6.  In addition, enrichment of
+apoptosis pathway in line 7 indicates that the programmed cell death could be
+induced by CTL response to eliminate the viruses.
 
-At this stage of infection, our results suggest that lytic infection of MDV stimulates
-both innate and adaptive immune responses, which leads to activation of T cells in line 7.
-Only activated T cells are infected by MDV, therefore, the lytic phase could facilitate the spread of
-the viruses by enhancing expansion of activated T cells.
-Due to cell-associate nature of MDV, the viruses transfer to T cells via cell-to-cell
-contact between B cells and
-T cells during antigen presentation or B cells activation by T helper cells.
-Therefore, it is beneficial for the host to restrain such contact.
-However, it is not clear how chickens in line 6 control the lytic infection of MDV.
+At this stage of infection, our results suggest that lytic infection of MDV
+stimulates both innate and adaptive immune responses, which leads to activation
+of T cells in line 7.  Only activated T cells are infected by MDV, therefore,
+the lytic phase could facilitate the spread of the viruses by enhancing
+expansion of activated T cells.  Due to cell-associate nature of MDV, the
+viruses transfer to T cells via cell-to-cell contact between B cells and T cells
+during antigen presentation or B cells activation by T helper cells.  Therefore,
+it is beneficial for the host to restrain such contact.  However, it is not
+clear how chickens in line 6 control the lytic infection of MDV.
 % Two mechanisms have been speculated to contribute to MD resistance.
@@
 The immune system is isoform-rich and many genes express different isoforms with
-distinctive functions in response to stimuli such as stress, chemicals and infection.
-Changes in expression of splice forms of immune related genes have been reported to
-be associated with increase susceptibility and poor prognosis of diseases~\cite{lynch2004consequences}.
-Studying differential isoform expression could therefore shed lights into inherent differences between 
-lines 6 and 7 that confer resistance or susceptibility to MD.
+distinctive functions in response to stimuli such as stress, chemicals and
+infection.  Changes in expression of splice forms of immune related genes have
+been reported to be associated with increase susceptibility and poor prognosis
+of diseases~\cite{lynch2004consequences}.  Studying differential isoform
+expression could therefore shed lights into inherent differences between lines 6
+and 7 that confer resistance or susceptibility to MD.
 
-In the past decades, microarray technology has been used to study gene and isoform
-expressions in many studies, but its sensitivity for detection of structurally similar
-isoforms is low, and known or predicted annotations are required to design probes~\cite{kane2000assessment}.
-Although RNA-Seq method can provide a reliable estimate of an exon expression compared
-to microarray~\cite{pan2008deep} and is not constrained to those limitations, studying isoform
-expressions using RNA-Seq is still not straightforward because of the short read lengths.
-Reads from current Illumina technology are generally not long enough to span across all exons in an isoform.
-In most cases, only exons in close proximity are covered by the same read, which makes it
-difficult to accurately predict a full structure of the isoform.
-In addition, some genes are fused due to overlapping untranslated regions (UTRs),
-which can also result in erroneous predicted isoform structures.
+In the past decades, microarray technology has been used to study gene and
+isoform expressions in many studies, but its sensitivity for detection of
+structurally similar isoforms is low, and known or predicted annotations are
+required to design probes~\cite{kane2000assessment}.  Although RNA-Seq method
+can provide a reliable estimate of an exon expression compared to
+microarray~\cite{pan2008deep} and is not constrained to those limitations,
+studying isoform expressions using RNA-Seq is still not straightforward because
+of the short read lengths.  Reads from current Illumina technology are generally
+not long enough to span across all exons in an isoform.  In most cases, only
+exons in close proximity are covered by the same read, which makes it difficult
+to accurately predict a full structure of the isoform.  In addition, some genes
+are fused due to overlapping untranslated regions (UTRs), which can also result
+in erroneous predicted isoform structures.
 
-Due to those issues, it is not feasible to accurately estimate expression of isoforms, especially when
-gene annotation is constructed from \textit{de novo} assembly~\cite{trapnell2013differential}.
-To avoid these issues, we chose to study exon expression instead of isoform expression.
-Using MISO with exon-centric method, only reads spanning across a few exons are used and only exons
-involved in a splicing event are examined.
-The expression of exon inclusion is calculated as Percent Spliced In (Psi or $\Psi$), which can be used to
-infer the portion of transcripts that include the exon in each sample~\cite{Katz:2010iv}.
-In this study, we investigated the three most common alternative splicing events in
-vertebrates, which are skipped exons (SE), an alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
+Due to those issues, it is not feasible to accurately estimate expression of
+isoforms, especially when gene annotation is constructed from {\em de novo}
+assembly~\cite{trapnell2013differential}.  To avoid these issues, we chose to
+study exon expression instead of isoform expression.  Using MISO with
+exon-centric method, only reads spanning across a few exons are used and only
+exons involved in a splicing event are examined.  The expression of exon
+inclusion is calculated as Percent Spliced In (Psi or $\Psi$), which can be used
+to infer the portion of transcripts that include the exon in each
+sample~\cite{Katz:2010iv}.  In this study, we investigated the three most common
+alternative splicing events in vertebrates, which are skipped exons (SE), an
+alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
 
@@
 
-Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20 when compared to line 7
-in infected chickens are shown in
+Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20
+when compared to line 7 in infected chickens are shown in
 Tables~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
-and~\ref{tab:line67i_diff_line67u_three}.
-Genes can be categorized roughly into four groups based on the pattern of $\Psi$ across control and
-infected birds in both lines.
+and~\ref{tab:line67i_diff_line67u_three}.  Genes can be categorized roughly into
+four groups based on the pattern of $\Psi$ across control and infected birds in
+both lines.
 
-Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes with $\Psi$s that were
-up- or down-regulated in infected chickens in line 6 only.
-This group includes \textit{BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
-associated C2H2-type zinc finger protein encoding gene, which functions as a tumour-suppressor
-in T-cell lymphoma in human.
-According to homologous alignments on UCSC genome browser, a splice form with the skipped exon
-is similar to mouse \textit{BCL11B isoform b}.
-The skipped exon was expressed 30\% in infected line 6 chickens; whereas it was rarely expressed
-(4-7\%) in the control line 6 and both groups in line 7.
-The skipped exon was not found to encode any known protein domain, however, it is in the middle of
-two adjacent C2H2-type finger protein domains.
-Another gene related to B and T cell lymphomas is \textit{SIK2} (salt-inducible kinase 2).
-This gene has been reported to have a negative effect on T cell lymphomas by limiting the transcription of
-HTLV-1 virus~\cite{tang2013lkb1}.
-Some other genes are important in pre-mRNA splicing including \textit{TRA2A},
-\textit{SRSF6} and \textit{GEMIN6}.
-\textit{TRA2A} (transformer-2 protein homolog alpha) encodes RNA recognition motif (RRM),
-the skipped exon is not found to encode any known protein domain.
-\textit{SRSF6} (SR splicing factor 6) encodes a nuclear protein that belongs to the splicing
-factor protein family.
-\textit{GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
-These genes could play a significant role in regulating inclusion of alternative exons.
-A gene in this group that might be important for innate immune responses is \textit{RAC3} (Ras-related C3
-botulinum toxin substrate 3).
-This gene encodes small GTPases, belonging to Ras family, that regulates a wide variety of cellular events
-including cell growth, cytoskeletal reorganization, and the activation of protein kinases.
-The role of small GTPases in immune responses is discussed further below.
+Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes with $\Psi$s
+that were up- or down-regulated in infected chickens in line 6 only.  This group
+includes {\em BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell
+lymphoma associated C2H2-type zinc finger protein encoding gene, which functions
+as a tumour-suppressor in T-cell lymphoma in human.  According to homologous
+alignments on UCSC genome browser, a splice form with the skipped exon is
+similar to mouse {\em BCL11B isoform b}.  The skipped exon was expressed 30\% in
+infected line 6 chickens; whereas it was rarely expressed (4-7\%) in the control
+line 6 and both groups in line 7.  The skipped exon was not found to encode any
+known protein domain, however, it is in the middle of two adjacent C2H2-type
+finger protein domains.  Another gene related to B and T cell lymphomas is {\em
+SIK2} (salt-inducible kinase 2).  This gene has been reported to have a negative
+effect on T cell lymphomas by limiting the transcription of HTLV-1
+virus~\cite{tang2013lkb1}.  Some other genes are important in pre-mRNA splicing
+including {\em TRA2A}, {\em SRSF6} and {\em GEMIN6}.  {\em TRA2A} (transformer-2
+protein homolog alpha) encodes RNA recognition motif (RRM), the skipped exon is
+not found to encode any known protein domain.  {\em SRSF6} (SR splicing factor
+6) encodes a nuclear protein that belongs to the splicing factor protein family.
+{\em GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
+These genes could play a significant role in regulating inclusion of alternative
+exons.  A gene in this group that might be important for innate immune responses
+is {\em RAC3} (Ras-related C3 botulinum toxin substrate 3).  This gene encodes
+small GTPases, belonging to Ras family, that regulates a wide variety of
+cellular events including cell growth, cytoskeletal reorganization, and the
+activation of protein kinases.  The role of small GTPases in immune responses is
+discussed further below.
 
-In group II (Table~\ref{tab:line67i_diff_line67u_two}),
- $\Psi$ values were relatively stable in control and infected chickens within line, but not between lines.
-Genes that could play an important role in immune responses are \textit{ITGB2} and \textit{HCK}.
-\textit{HCK} transmits signal from cell surface receptors such as \textit{FCGR1A, FCGR2A, IL2, IL6,
-IL18}, and integrins (\textit{ITGB1, ITGB2}).
-\textit{ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of \textit{LFA-1} and \textit{CR3} receptors.
-\textit{LFA-1} plays an important role in adhesion of lymphocytes with other cells.
-\textit{CR3} binds to a vast array of ligands and molecules including complement C3bi, microbial proteins,
-ICAM-1 and -2, ECM proteins, and coagulation proteins~\cite{}.
-It plays a significant role in neutrophils and monocytes activation including phagocytosis, adhesion and
-migration~\cite{}.
-The role of \textit{ITGB2} in immune responses is discussed further in the next section.
-\textit{DYNLT1,DYNLL2, SEPT11, PFN2}, and \textit{ZDHHC7} are also involved in cell
-rearrangement and cytokinesis.
-In particular, \textit{DYNLT1} and \textit{DYNLL2} are dynein proteins that have been demonstrated to
+In group II (Table~\ref{tab:line67i_diff_line67u_two}), $\Psi$ values were
+relatively stable in control and infected chickens within line, but not between
+lines.  Genes that could play an important role in immune responses are {\em
+ITGB2} and {\em HCK}.  {\em HCK} transmits signal from cell surface receptors
+such as {\em FCGR1A, FCGR2A, IL2, IL6, IL18}, and integrins ({\em ITGB1,
+ITGB2}).  {\em ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of {\em LFA-1}
+and {\em CR3} receptors.  {\em LFA-1} plays an important role in adhesion of
+lymphocytes with other cells.  {\em CR3} binds to a vast array of ligands and
+molecules including complement C3bi, microbial proteins, ICAM-1 and -2, ECM
+proteins, and coagulation proteins~\cite{}.  It plays a significant role in
+neutrophils and monocytes activation including phagocytosis, adhesion and
+migration~\cite{}.  The role of {\em ITGB2} in immune responses is discussed
+further in the next section.  {\em DYNLT1,DYNLL2, SEPT11, PFN2}, and {\em
+ZDHHC7} are also involved in cell rearrangement and cytokinesis.  In particular,
+{\em DYNLT1} and {\em DYNLL2} are dynein proteins that have been demonstrated to
 regulate T cell activation by driving T cell receptor microclusters (TCR-MCs)
@@
 
-Group III (Table~\ref{tab:line67i_diff_line67u_three})
-includes genes that exhibit differential isoform expression only in infected line 7.
-A number of genes in this group encode proteins that are parts of
-spliceosome: \textit{SRSF3}, \textit{HNRNPDL},
-\textit{SFSWAP}, \textit{THOC1}, \textit{RNPC3} and \textit{SRSF5}.
-Two genes involved in cell-cell contact regulated by integrins are also in this group.
-\textit{PPP1R12A} is a myosin phosphatase that regulates the interaction of actin
-and myosin downstream of GTPase Rho proteins.
-\textit{PODXL} encodes PODX-like protein that functions in integrin-dependent
-manner as both pro-adhesive and anti-adhesive molecules.
-This protein is involved in cell-to-cell contact, cell trafficking, and cancer progression~\cite{}.
+Group III (Table~\ref{tab:line67i_diff_line67u_three}) includes genes that
+exhibit differential isoform expression only in infected line 7.  A number of
+genes in this group encode proteins that are parts of spliceosome: {\em SRSF3},
+{\em HNRNPDL}, {\em SFSWAP}, {\em THOC1}, {\em RNPC3} and {\em SRSF5}.  Two
+genes involved in cell-cell contact regulated by integrins are also in this
+group.  {\em PPP1R12A} is a myosin phosphatase that regulates the interaction of
+actin and myosin downstream of GTPase Rho proteins.  {\em PODXL} encodes
+PODX-like protein that functions in integrin-dependent manner as both
+pro-adhesive and anti-adhesive molecules.  This protein is involved in
+cell-to-cell contact, cell trafficking, and cancer progression~\cite{}.
 
-The last group (Group IV, Table~\ref{tab:line67i_diff_line67u_three})
-only has one gene, \textit{GOSR1}.
-The $\Psi$ value of this gene was less than 0.20 cutoff in control and infected chickens in line 6, but it is
-significantly different between infected chickens in line 6 and 7.
-Also, there is a significant difference between control and infected chickens in line 7.
-This gene encodes a trafficking membrane protein important for transporting proteins
-from \textit{cis-} to \textit{trans-}golgi network.
+The last group (Group IV, Table~\ref{tab:line67i_diff_line67u_three}) only has
+one gene, {\em GOSR1}.  The $\Psi$ value of this gene was less than 0.20 cutoff
+in control and infected chickens in line 6, but it is significantly different
+between infected chickens in line 6 and 7.  Also, there is a significant
+difference between control and infected chickens in line 7.  This gene encodes a
+trafficking membrane protein important for transporting proteins from {\em cis-}
+to {\em trans-}golgi network.
 
-\subsubsection*{Roles of \textit{LFA-1} and actin cytoskeleton pathway in T cells activation}
+\subsubsection*{Roles of {\em LFA-1} and actin cytoskeleton pathway in T cells activation}
 
-Integrins are a special kind of receptors that transmit signals bidirectionally across the cell membrane.
-They are heterodimeric composed of an $\alpha$ (large) and a $\beta$ (small)
-subunits~\cite{wang2010immunopathologies}.
-$\beta_{2}$ (CD18) subunit encoded by \textit{ITGB2} is expressed on lymphocytes and antigen presenting cells
-(APCs) as a component of \textit{LFA-1} and \textit{CR3} receptors.
-LFA-1 binds to its ligand ICAM-1 to help form a synapse that bring APC and T cell together to
-initiate antigen presentation leading to T cell activation~\cite{dustin2000immunological}.
-Absence of LFA-1 leads to impaired functions of lymphocytes in proliferation and
-tumor rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.
-Mutations in \textit{ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency (LAD-1),
-an autosomal-recessive inherited disease found in few families.
-The disease is characterized by impair of lymphocytes in adherent-dependent functions, lack of accumulation
-to the site of infection and recurrent bacterial and fungal infection~\cite{springer1987lymphocyte}.
+Integrins are a special kind of receptors that transmit signals bidirectionally
+across the cell membrane.  They are heterodimeric composed of an $\alpha$
+(large) and a $\beta$ (small) subunits~\cite{wang2010immunopathologies}.
+$\beta_{2}$ (CD18) subunit encoded by {\em ITGB2} is expressed on lymphocytes
+and antigen presenting cells (APCs) as a component of {\em LFA-1} and {\em CR3}
+receptors.  LFA-1 binds to its ligand ICAM-1 to help form a synapse that bring
+APC and T cell together to initiate antigen presentation leading to T cell
+activation~\cite{dustin2000immunological}.  Absence of LFA-1 leads to impaired
+functions of lymphocytes in proliferation and tumor
+rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.  Mutations in {\em
+ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency
+(LAD-1), an autosomal-recessive inherited disease found in few families.  The
+disease is characterized by impair of lymphocytes in adherent-dependent
+functions, lack of accumulation to the site of infection and recurrent bacterial
+and fungal infection~\cite{springer1987lymphocyte}.
 
-Actin cytoskeleton is very important in T cell activation because it enhances activity of
-LFA-1 by increasing its avidity and recruiting signaling molecules necessary for
-downstream signaling~\cite{dustin2000immunological, van2000avidity}.
-Cytoskeleton proteins binding to cytoplasmic domain of LFA-1 are thought to play an important role in
-driving LFA-1 to aggregation on the cell surface, resulting in increased avidity.
-Aggregation of LFA-1 has been demonstrated to be essential for lymphocytes to bind to the
-ligand~\cite{van1994extracellular}.
-Interestingly, \textit{RAC3}, \textit{PFN2}, and \textit{PPP1R12A}, which are involved in
-actin cytoskeleton pathway (Figure~\ref{kegg_actin}), also expressed different ratios of alternative splice
-forms between lines 6 and 7.
-Some of these genes are also co-present in three other pathways
-that involved in immune responses (Table~\ref{tab:integrin}).
-This could be speculated that pre-mRNA splicing of these genes is co-regulated by splicing
-regulators or some genetic factors.
+Actin cytoskeleton is very important in T cell activation because it enhances
+activity of LFA-1 by increasing its avidity and recruiting signaling molecules
+necessary for downstream signaling~\cite{dustin2000immunological,
+van2000avidity}.  Cytoskeleton proteins binding to cytoplasmic domain of LFA-1
+are thought to play an important role in driving LFA-1 to aggregation on the
+cell surface, resulting in increased avidity.  Aggregation of LFA-1 has been
+demonstrated to be essential for lymphocytes to bind to the
+ligand~\cite{van1994extracellular}.  Interestingly, {\em RAC3, PFN2,} and {\em
+PPP1R12A}, which are involved in actin cytoskeleton pathway
+(Figure~\ref{kegg_actin}), also expressed different ratios of alternative splice
+forms between lines 6 and 7.  Some of these genes are also co-present in three
+other pathways that involved in immune responses (Table~\ref{tab:integrin}).
+This could be speculated that pre-mRNA splicing of these genes is co-regulated
+by splicing regulators or some genetic factors.
 
 \subsubsection*{Protein domains}
-To understand the role of the alternative splice forms on the function of these genes,
-transcripts were translated to protein sequences by ESTscan~\cite{iseli1999estscan}.
-Protein sequences were then searched for annotated protein domains using
-the standalone version of InterPro Scan~\cite{quevillon2005interproscan}.
-Besides \textit{ITGB2}, other genes have alternative exons located in coding regions and
-could potentially affect functional protein domains in some ways.
-The exon with alternative 3$\prime$ splice site of \textit{RAC3} encode part of a protein
-domain identified as small GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).
-Rac3 is highly homologous to Rac1 and has been reported to possess the ability of
-promoting membrane ruffling, transformation, activation of
-c-Jun transcriptional activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.
-Activated Rac also regulates production of superoxide in neutrophils and macrophages.
 
-The alternative exon of \textit{PFN2} seems to disrupt the coding sequence that
-encode profilin domain (Pfam:PF00235).
-The profilin domain is essential for almost all organisms and its functions include
-regulating actin polymerization, controlling complex network of molecular interaction
-and transmitting signals from small-GTPase pathway.
-It also binds to Rac effector molecules and a number of other ligands~\cite{witke2004role}.
-The skipped exon of \textit{PPP1R12A} encodes part of a protein domain annotated as protein
-phosphatase 1 (PIRSF:PIRSF038141).
-The functional effects of alternative splicing on these functional domains have to be
-further investigated by simulation or an experiment.
+To understand the role of the alternative splice forms on the function of these
+genes, transcripts were translated to protein sequences by
+ESTscan~\cite{iseli1999estscan}.  Protein sequences were then searched for
+annotated protein domains using the standalone version of InterPro
+Scan~\cite{quevillon2005interproscan}.  Besides {\em ITGB2}, other genes have
+alternative exons located in coding regions and could potentially affect
+functional protein domains in some ways.  The exon with alternative 3$\prime$
+splice site of {\em RAC3} encode part of a protein domain identified as small
+GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).  Rac3 is
+highly homologous to Rac1 and has been reported to possess the ability of
+promoting membrane ruffling, transformation, activation of c-Jun transcriptional
+activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.  Activated
+Rac also regulates production of superoxide in neutrophils and macrophages.
 
-% Based on our gene model, the skipped exon of \textit{ITGB2} is located in the 5$\prime$ UTR;
-% whereas, the skipped exon of \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).
+The alternative exon of {\em PFN2} seems to disrupt the coding sequence that
+encode profilin domain (Pfam:PF00235).  The profilin domain is essential for
+almost all organisms and its functions include regulating actin polymerization,
+controlling complex network of molecular interaction and transmitting signals
+from small-GTPase pathway.  It also binds to Rac effector molecules and a number
+of other ligands~\cite{witke2004role}.  The skipped exon of {\em PPP1R12A}
+encodes part of a protein domain annotated as protein phosphatase 1
+(PIRSF:PIRSF038141).  The functional effects of alternative splicing on these
+functional domains have to be further investigated by simulation or an
+experiment.
 
-\subsubsection*{Prediction of \textit{cis}-regulatory elements in alternative splicing exons}
+% Based on our gene model, the skipped exon of \textit{ITGB2} is located in the
+% 5$\prime$ UTR; whereas, the skipped exon of \textit{HCK}, encodes protein
+% tyrosine kinase (Pfam:PF07714).
 
-Although alternative splicing of genes in all groups could be regulated by
-\textit{cis-} and \textit{trans-}splicing factors,
-genes in group II are more likely to be regulated by \textit{cis-}regulatory factors.
-The ratios of isoform expression in this group were relatively stable within line,
-but were significantly different between lines.
-Investigation of nucleotide differences within exons of line 6 and 7 could reveal
-a possible role of cSNPs in regulating alternative splicing in this group.
-We obtained a sequence of an alternative exon from line 6 and used Human Splicing Finder (HSF)
-to determine whether SNPs from line 7 could alter predicted ESEs or ESSs.
-Results from some genes involved in cytokinesis are discussed in this section.
-Exonic SNPs from line 6 and 7 are listed in Table~\ref{tab:deu_snps}.
+\subsubsection*{Prediction of {\em cis}-regulatory elements in alternative splicing exons}
 
-For \textit{ITGB2}, a SNP (T) at position 26 of the cDNA from line 6, which corresponds to
-position 7,183,696 on chromosome 7 is located in a predicted binding site for
-SC35, which is an exon enhancer.
-Although the exon is not expressed in line 7, we found that there is no polymorphisms between
-lines 6 and 7 according to SNPs data from the genome resequencing.
-Therefore, this SNP may not be accounted for exclusion of the exon in
-line 7 (Figure~\ref{itgb2}).
+Although alternative splicing of genes in all groups could be regulated by {\em
+cis-} and {\em trans-}splicing factors, genes in group II are more likely to be
+regulated by {\em cis-}regulatory factors.  The ratios of isoform expression in
+this group were relatively stable within line, but were significantly different
+between lines.  Investigation of nucleotide differences within exons of line 6
+and 7 could reveal a possible role of cSNPs in regulating alternative splicing
+in this group.  We obtained a sequence of an alternative exon from line 6 and
+used Human Splicing Finder (HSF) to determine whether SNPs from line 7 could
+alter predicted ESEs or ESSs.  Results from some genes involved in cytokinesis
+are discussed in this section.  Exonic SNPs from line 6 and 7 are listed in
+Table~\ref{tab:deu_snps}.
 
-Exon sequences of \textit{PFN2} from line 6 and 7 differ at position 2,072 on
-cDNA or 23,221,934 on chromosome 9.
-A small insertion of two AA nucleotides is predicted to create a new binding site
-for Tra2-$\beta$ splicing regulator, which serves as a stabilizer of an enhancer
-complex~\cite{lopez1998alternative}.
+For {\em ITGB2}, a SNP (T) at position 26 of the cDNA from line 6, which
+corresponds to position 7,183,696 on chromosome 7 is located in a predicted
+binding site for SC35, which is an exon enhancer.  Although the exon is not
+expressed in line 7, we found that there is no polymorphisms between lines 6 and
+7 according to SNPs data from the genome resequencing.  Therefore, this SNP may
+not be accounted for exclusion of the exon in line 7 (Figure~\ref{itgb2}).
+
+Exon sequences of {\em PFN2} from line 6 and 7 differ at position 2,072 on cDNA
+or 23,221,934 on chromosome 9.  A small insertion of two AA nucleotides is
+predicted to create a new binding site for Tra2-$\beta$ splicing regulator,
+which serves as a stabilizer of an enhancer complex~\cite{lopez1998alternative}.
 From exon expression data, $\Psi$ of an exon with alternative splicing increases
-from 0.20-0.30 to about 0.50 in line 7.
-Tra2 could possibly increase inclusion of the exon with alternative 3$\prime$ splice site via
-ESE-dependent 3$\prime$ splice site activation (Figure~\ref{pfn2}).
-Even though the linear distance between Tra2-$\beta$ binding site and the alternative
-3$\prime$ splice site is greater than 1kb, Tra2 could possibly get close to
-the 3$\prime$ splice site in the secondary structure of the mRNA.
+from 0.20-0.30 to about 0.50 in line 7.  Tra2 could possibly increase inclusion
+of the exon with alternative 3$\prime$ splice site via ESE-dependent 3$\prime$
+splice site activation (Figure~\ref{pfn2}).  Even though the linear distance
+between Tra2-$\beta$ binding site and the alternative 3$\prime$ splice site is
+greater than 1kb, Tra2 could possibly get close to the 3$\prime$ splice site in
+the secondary structure of the mRNA.
 
-Replacement of A with G nucleotide in the skipped exon of \textit{DYNLL2} from line 6
-is predicted to slightly alter the binding site of several ESEs as well as to create
-a new binding site for 9G8.
-This exon is upregulated in line 7 compared to line 6, therefore, the present
-of the new binding site for 9G8 exon enhancer helps support the expression results.
-In addition, A nucleotide is this position matches the reference
-nucleotide, therefore, we could expect this exon to be expressed in other datasets.
-According to EST tags on the UCSC genome browser, the exon has been found and sequenced
-from chicken eyes (15d post-hatched, EST sequence:DR424100).
+Replacement of A with G nucleotide in the skipped exon of {\em DYNLL2} from line
+6 is predicted to slightly alter the binding site of several ESEs as well as to
+create a new binding site for 9G8.  This exon is upregulated in line 7 compared
+to line 6, therefore, the present of the new binding site for 9G8 exon enhancer
+helps support the expression results.  In addition, A nucleotide is this
+position matches the reference nucleotide, therefore, we could expect this exon
+to be expressed in other datasets.  According to EST tags on the UCSC genome
+browser, the exon has been found and sequenced from chicken eyes (15d
+post-hatched, EST sequence:DR424100).
 
-For \textit{DYNLT1}, the skipped exon was absent in line 7, therefore, we could not
-determine the SNPs from RNA-Seq data.
-However, data from DNA resequencing showed that there was a polymorphism at position
-51,357,865 on chromosome 3.
-Replacing G with C from line 7 is predicted to slightly alter predicted binding sites
-for 9G8 as well as SRp55 exon enhancer.
-However, whether the alteration down regulates the exon in line 7 is unclear.
+For {\em DYNLT1}, the skipped exon was absent in line 7, therefore, we could not
+determine the SNPs from RNA-Seq data.  However, data from DNA resequencing
+showed that there was a polymorphism at position 51,357,865 on chromosome 3.
+Replacing G with C from line 7 is predicted to slightly alter predicted binding
+sites for 9G8 as well as SRp55 exon enhancer.  However, whether the alteration
+down regulates the exon in line 7 is unclear.
 
-There are too many SNPs in the exon of \textit{SEPT11} and \textit{ZDHHC7}, making it
-unfeasible to predict which SNP regulates the exon expression.
-Therefore, we do not discuss these two genes in this section.
-However, results from HSF analysis of these two genes and
-other genes in this group are provided in the supplementary materials.
+There are too many SNPs in the exon of {\em SEPT11} and {\em ZDHHC7}, making it
+unfeasible to predict which SNP regulates the exon expression.  Therefore, we do
+not discuss these two genes in this section.  However, results from HSF analysis
+of these two genes and other genes in this group are provided in the
+supplementary materials.
 
@@
 
-Custom gene models built from combination of gene models from \textit{de novo} assembly,
-reference-based assembly, and Ensembl allowed us to identify genes and isoforms
-that might play an important role in resistance to MD.
-Results from gene expression analysis indicate that adaptive immune responses were active
-during lytic infection in line 7.
-Because only activated T cells are infected, elicitation of adaptive immune responses could
-help spread the viruses by recruiting and activating more T cells.
-Therefore, active immune responses are favorable for the virus.
-On the other hand, the delay of adaptive immune responses in line 6, in this case,
-could benefit the host by limiting infection of activated T cells.
+Custom gene models built from combination of gene models from {\em de novo}
+assembly, reference-based assembly, and Ensembl allowed us to identify genes and
+isoforms that might play an important role in resistance to MD.  Results from
+gene expression analysis indicate that adaptive immune responses were active
+during lytic infection in line 7.  Because only activated T cells are infected,
+elicitation of adaptive immune responses could help spread the viruses by
+recruiting and activating more T cells.  Therefore, active immune responses are
+favorable for the virus.  On the other hand, the delay of adaptive immune
+responses in line 6, in this case, could benefit the host by limiting infection
+of activated T cells.
 
-Using exon-centric approach, we identified several genes involved in
-actin cytoskeleton, especially \textit{ITGB2},
-that expressed different ratios of alternative isoforms between lines 6 and 7.
-As described above, \textit{ITGB2} encodes $\beta$-subunit of integrins including LFA-1, which exclusively
-expressed in lymphocytes and plays a major role in lymphoproliferation,
-antigen presentation, T cells activation, and cytotoxicity.
-In patients with LFA-1 deficiency (LAD), the response of lymphocytes to mitogens is
-decreased~\cite{springer1987lymphocyte}.
-The decrease in responsiveness to mitogens have been shown to correlate with resistance to MD
-by Lee and Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds
-(line 6 and N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line 7 and P).
-Even though the exact mechanism is not known, incompetency of T cells in response to
-stimuli appears to benefit resistant birds because MD virus could not induce T cells
-to proliferate and cause them to undergo neoplastic transformation.
-It has also been suggested that the mechanism that controls
-both lymphocyte proliferation induced by MD virus and lymphocyte proliferation induced by
-immune response is the same~\cite{pazderka1975histocompatibility}.
-Therefore, it may be useful to consider a link between deficiency of lymphocytes in line 6 to
-the alternative splice form of \textit{ITGB2} that is only expressed in line 6.
-Although the exon included in the alternative splice form is non-coding, it could serve important functions in
-translation or posttranscriptional regulation.
+Using exon-centric approach, we identified several genes involved in actin
+cytoskeleton, especially {\em ITGB2}, that expressed different ratios of
+alternative isoforms between lines 6 and 7.  As described above, {\em ITGB2}
+encodes $\beta$-subunit of integrins including LFA-1, which exclusively
+expressed in lymphocytes and plays a major role in lymphoproliferation, antigen
+presentation, T cells activation, and cytotoxicity.  In patients with LFA-1
+deficiency (LAD), the response of lymphocytes to mitogens is
+decreased~\cite{springer1987lymphocyte}.  The decrease in responsiveness to
+mitogens have been shown to correlate with resistance to MD by Lee and
+Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
+N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
+7 and P).
 
-Interestingly, by grouping genes based on patterns of $\Psi$s, we found that many genes in group
-II: \textit{ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and \textit{ZDHHC7}, are involved in cytokinesis or
-cell synapse, which are important for aggregation of LFA-1 and T cell activation as discussed above.
-The $\Psi$ values in group II are more similar within line than between lines in both control and
-infected groups, suggesting the possibility that the alternative splicing is
-controlled by a genetic mechanism.
-Based on the independent criteria, enrichment of cytoskeletal genes in this group could not be
-just a mere coincidence, but an evidence suggesting that inefficiency of T cell activation
-in line 6 is linked to a genetic factor that regulate alternative splicing of \textit{ITGB2}
-and other genes in the cytoskeleton system.
-Experimental validation of exonic SNPs provided by this study could shed some lights on
-underlying polymorphisms that contribute to MD resistance.
+Even though the exact mechanism is not known, incompetency of T cells in
+response to stimuli appears to benefit resistant birds because MD virus could
+not induce T cells to proliferate and cause them to undergo neoplastic
+transformation.  It has also been suggested that the mechanism that controls
+both lymphocyte proliferation induced by MD virus and lymphocyte proliferation
+induced by immune response is the same~\cite{pazderka1975histocompatibility}.
+Therefore, it may be useful to consider a link between deficiency of lymphocytes
+in line 6 to the alternative splice form of {\em ITGB2} that is only expressed
+in line 6.  Although the exon included in the alternative splice form is
+non-coding, it could serve important functions in translation or
+posttranscriptional regulation.
 
-To conclude, we attempted to employ custom gene models built from RNA-Seq data to
-identify underlying genetic factors contributing to MD resistance.
-We identified a number of genes that expressed different ratios of alternative splice forms
-between lines 6 and 7.
-Interestingly, several genes are involved in actin cytoskeleton and cytokinesis, which have
-not been of great interest in the filed of MD research.
-Even though we mainly discussed the possible role of \textit{ITGB2} in MD resistance, other
-genes cannot be precluded and should be a candidate for further investigation and experimental studies.
-Moreover, a full mechanism of MD resistance may be highly complex and require more data from different
-stages of infection as well as a greater sequencing depth in order to identify all
-genes and isoforms involved.
-To enable the study of unannotated gene and isoform expression,
-our approach of constructing gene models from RNA-Seq should be iteratively used to extend
-the Ensembl gene models.
+Interestingly, by grouping genes based on patterns of $\Psi$s, we found that
+many genes in group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em
+ZDHHC7}, are involved in cytokinesis or cell synapse, which are important for
+aggregation of LFA-1 and T cell activation as discussed above.  The $\Psi$
+values in group II are more similar within line than between lines in both
+control and infected groups, suggesting the possibility that the alternative
+splicing is controlled by a genetic mechanism.  Based on the independent
+criteria, enrichment of cytoskeletal genes in this group could not be just a
+mere coincidence, but an evidence suggesting that inefficiency of T cell
+activation in line 6 is linked to a genetic factor that regulate alternative
+splicing of {\em ITGB2} and other genes in the cytoskeleton system.
+Experimental validation of exonic SNPs provided by this study could shed some
+lights on underlying polymorphisms that contribute to MD resistance.
 
-% However, to mechanism that contribute to resistance to MD could be very complex and integration
-% of global gene expression data as well as experimental data from different stages of infection will be
-% required to fully elucidate.
+To conclude, we attempted to employ custom gene models built from RNA-Seq data
+to identify underlying genetic factors contributing to MD resistance.  We
+identified a number of genes that expressed different ratios of alternative
+splice forms between lines 6 and 7.  Interestingly, several genes are involved
+in actin cytoskeleton and cytokinesis, which have not been of great interest in
+the filed of MD research.  Even though we mainly discussed the possible role of
+{\em ITGB2} in MD resistance, other genes cannot be precluded and should be a
+candidate for further investigation and experimental studies.  Moreover, a full
+mechanism of MD resistance may be highly complex and require more data from
+different stages of infection as well as a greater sequencing depth in order to
+identify all genes and isoforms involved.  To enable the study of unannotated
+gene and isoform expression, our approach of constructing gene models from
+RNA-Seq should be iteratively used to extend the Ensembl gene models.
+
+% However, to mechanism that contribute to resistance to MD could be very
+% complex and integration of global gene expression data as well as experimental
+% data from different stages of infection will be required to fully elucidate.
 % Moreover, better gene models and a reference genome will be necessary.
@@
 Due to lack of complete gene models for chickens, we employed two methods to
-construct gene models from RNA-Seq reads.
-First, short reads were quality trimmed with conditri/2.1~\cite{}
-and assembled using Velvet/1.21~\cite{} and Oases/0.2.06~\cite{} to obtain long transcripts.
-Assembly was done with hash lengths range from 21 to 31 for both local and global assembly
-(described in Gimme paper).
-Poly-A tails were trimmed and low complexity transcripts were removed by Seqclean~\cite{}.
-All transcripts were then aligned to chicken reference genome (galGal4, with unplaced and random 
-chromosomes removed) with BLAT~\cite{}.
-Second, reads were aligned to reference genome using Tophat/2.0.5~\cite{} and \textit{Ensemble} gene
-models release 73 was used to guide reference-based assembly by Cufflinks2~\cite{}.
-Alignments from BLAT and models from Cufflinks were then combined to construct gene models by Gimme
-(manuscript in preparation).
+construct gene models from RNA-Seq reads.  First, short reads were quality
+trimmed with conditri/2.1~\cite{} and assembled using Velvet/1.21~\cite{} and
+Oases/0.2.06~\cite{} to obtain long transcripts.  Assembly was done with hash
+lengths range from 21 to 31 for both local and global assembly (described in
+Gimme paper).  Poly-A tails were trimmed and low complexity transcripts were
+removed by Seqclean~\cite{}.  All transcripts were then aligned to chicken
+reference genome (galGal4, with unplaced and random chromosomes removed) with
+BLAT~\cite{}.  Second, reads were aligned to reference genome using
+Tophat/2.0.5~\cite{} and ENSEMBLE gene models release 73 was used to guide
+reference-based assembly by Cufflinks2~\cite{}.  Alignments from BLAT and models
+from Cufflinks were then combined to construct gene models by Gimme (manuscript
+in preparation).
 
@@
 
-To identify DE genes, reads were mapped to transcripts by RSEM v.1.10.1~\cite{li2011rsem}, which is
-also used to estimate genes expression and identify DE genes.
-Data from single- and paired-end datasets from the same line were treated as biological replicates.
-To identify enriched pathways and ontology terms, a list of DE genes was analysed by GOSeq v.1.10.0 based on
-chicken KEGG annotations.
-P-values were corrected by Benjamini-Hochberg multiple testing correction.
-Genes, patways and GO terms with corrected P-value $<0.1$ were considered significant.
-Pathview~\cite{luo2013pathview} was used to create a KEGG pathway diagram with colors representing relative level of gene expressions.
+To identify DE genes, reads were mapped to transcripts by RSEM
+v.1.10.1~\cite{li2011rsem}, which is also used to estimate genes expression and
+identify DE genes.  Data from single- and paired-end datasets from the same line
+were treated as biological replicates.  To identify enriched pathways and
+ontology terms, a list of DE genes was analysed by GOSeq v.1.10.0 based on
+chicken KEGG annotations.  P-values were corrected by Benjamini-Hochberg
+multiple testing correction.  Genes, patways and GO terms with corrected P-value
+$<0.1$ were considered significant.  Pathview~\cite{luo2013pathview} was used to
+create a KEGG pathway diagram with colors representing relative level of gene
+expressions.
 
 \subsection{Differential Splicing Event}
+
 Gene models were converted to alternative splicing models using a Python script.
-In order to increase sensitivity, read counts from single- and paired-end samples were combined and treated
-as single-end reads for splicing event analysis with MISO/0.4.9~\cite{Katz:2010iv}.
-Splicing events with Bayes factor $>10$ and $\Delta\Psi>0.20$ were considered significant.
-Read coverages and $\Psi$ distributions were plotted using Sashimi plot~\cite{Katz:2013vx}.
+In order to increase sensitivity, read counts from single- and paired-end
+samples were combined and treated as single-end reads for splicing event
+analysis with MISO/0.4.9~\cite{Katz:2010iv}.  Splicing events with Bayes factor
+$>10$ and $\Delta\Psi>0.20$ were considered significant.  Read coverages and
+$\Psi$ distributions were plotted using Sashimi plot~\cite{Katz:2013vx}.
 
-\subsection{Variant Calling and \textit{In Silico} Splicing Analysis}
-Variants were called using mpileup command from SAMTools/0.1.18~\cite{} and BCFTools~\cite{}.
-Only variants with quality score $\ge20$ were used for mutation analyses.
-Exon enhancers and suppressors were predicted using Human Splicing Finder web portal~\cite{}.
-Human default parameter settings were used in all analyses.
-The following regulatory sequences were used to determined the effect of variants:
-HSF integrated matrices for serine/arginine-rich proteins (SRp40, SC35, SF2/ASF, SF2/ASF,
-IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE), RESCUE-ESE hexamer (RESCUE-ESE),
-putative 8-mer ESEs (PESEs) and putative 8-mers exonic splicing silencers (PESSs),
-exon-identity and intron-identiry elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding
-motifs, and Fas exonic splicing silencers.
+\subsection{Variant Calling and {\em In Silico} Splicing Analysis} Variants were
+called using mpileup command from SAMTools/0.1.18~\cite{} and BCFTools~\cite{}.
+Only variants with quality score $\ge20$ were used for mutation analyses.  Exon
+enhancers and suppressors were predicted using Human Splicing Finder web
+portal~\cite{}.  Human default parameter settings were used in all analyses.
+The following regulatory sequences were used to determined the effect of
+variants: HSF integrated matrices for serine/arginine-rich proteins (SRp40,
+SC35, SF2/ASF, SF2/ASF, IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE),
+RESCUE-ESE hexamer (RESCUE-ESE), putative 8-mer ESEs (PESEs) and putative 8-mers
+exonic splicing silencers (PESSs), exon-identity and intron-identiry elements
+(EIEs and IIEs), hetronuclear ribonucleoprotein-binding motifs, and Fas exonic
+splicing silencers.
 
@@
 %         Number of genes and isoforms from gene models
-%         built from \textit{de novo} assembly and Cufflinks.
+%         built from {\em de novo} assembly and Cufflinks.
 %         Combination of both methods decreased the number of genes
@@
 \caption{
-\bf{Pathways containing \textit{RAC3, ITGB2, PFN2} and \textit{PPP1R12A}}}
+\bf{Pathways containing {\em RAC3, ITGB2, PFN2} and {\em PPP1R12A}}}
 \begin{tabular}{ccc}
@@
 \hline
-hsa04810 & Regulation of actin cytoskeleton & \textit{RAC3, ITGB2, PPP1R12A, PFN2} \\
-hsa04015 & RAP1 signaling pathway & \textit{RAC3, ITGB2, PFN2} \\
-hsa04650 & Natural killer cells cytotoxicity & \textit{RAC3, ITGB2} \\
-hsa05416 & Viral myocarditis & \textit{RAC3, ITGB2} \\
-hsa04510 & Focal adhesion & \textit{RAC3, PPP1R12A} \\
+hsa04810 & Regulation of actin cytoskeleton & {\em RAC3, ITGB2, PPP1R12A, PFN2} \\
+hsa04015 & RAP1 signaling pathway & {\em RAC3, ITGB2, PFN2} \\
+hsa04650 & Natural killer cells cytotoxicity & {\em RAC3, ITGB2} \\
+hsa05416 & Viral myocarditis & {\em RAC3, ITGB2} \\
+hsa04510 & Focal adhesion & {\em RAC3, PPP1R12A} \\
 \hline
@@
 \hline
-ITGB2 & 7 & 7183696 & C & \textit{T} & . & - \\
-PFN2 & 9 & 23221934 & -  - & . & \textit{AA} & + \\
-DYNLT1 & 3 & 51357865 & C & . & \textit{G} & - \\
-DYNLL2 & 19 & 8694148 & G & . & \textit{A} & - \\
+ITGB2 & 7 & 7183696 & C & {\em T} & . & - \\
+PFN2 & 9 & 23221934 & -  - & . & {\em AA} & + \\
+DYNLT1 & 3 & 51357865 & C & . & {\em G} & - \\
+DYNLL2 & 19 & 8694148 & G & . & {\em A} & - \\
 \hline

### 2014-07-16 paper.tex
polished the conclusion part

@@
 % Results and Discussion can be combined.
-\section*{Results}
+\section*{Results and Discussion}
 
-\subsection*{Differential gene expression in response to MDV infection}
+\subsection*{Differential expression results from our method are comparable to
+previous studies}
 
@@
 assembly to build custom gene models.  The models, therefore, include both
-ENSEMBL annotated transcripts and putative transcripts and genes.  The advantage
-of using custom gene models is it allows an investigation of unannotated genes
-and isoforms, which is necessary for in-depth transcriptome study of an immune
-system.
+ENSEMBL annotated transcripts and putative genes and isoforms.  The advantage of
+using custom gene models is it allows an investigation of unannotated genes and
+isoforms, which is necessary for in-depth study of an immune system.
+
+Some DE genes have been reported from previous microarray studies were also
+found to be differentially expressed in this study.  For example, B6.1 (Bu-1)
+gene is known to be down-regulated approximately 2.27 fold in susceptible
+chicken with the MHC allele B\textsuperscript{19} at 4
+d.p.i~\cite{sarson2008transcriptional}.  It was also found to be down-regulated
+about 3-fold in line 7 (susceptible).  Similarly, {\em GMZA} gene reported to be
+upregulated across genetically different chickens (B\textsuperscript{19},
+B\textsuperscript{21} alleles),
+was also found to be highly upregulated.  In contrast, some genes that have been
+reported to be highly expressed in resistant chickens were downregulated in both
+lines. Those genes are {\em AMIGO2, MMP13} and {\em CLEC3B}, which were found to
+be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
+% Several genes involved in innate immune response were differentially expressed
+% in resistant chickens: \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all
+% those genes, only HSP5 was upregulated.
+Other immune genes reported to be highly expressed in susceptible chickens (line
+7) including {\em AVD, ART1, NOS2, CXCL13L2, MX1} and {\em
+SOCS1}~\cite{smith2011systems} were also found to be highly upregulated in both
+line 6 and 7.  However, our results show the similar expression patterns for
+{\em IL6} and {\em IL18}, which were only upregulated in line 7 in early stage
+of infection ($3-5$ d.p.i).
+
+In contrast, {\em IL15} has been reported non differential-expressed between
+control and infected chickens in line 6 and 7~\cite{kaiser2003differential};
+however, it was only upregulated in line 7 in our study.  Expression of {\em
+IL15} is induced by {\em TLR9}, which binds to non-methylated CpG residues
+present in genomes of many DNA viruses, including herpes simplex virus.  This
+cytokine auto-regulates the expression of {\em CD40}, which is a transmembrane
+receptor required for activation of macrophages by CD4 T cells.  Consequently,
+{\em CD40} was only upregulated in line 7 (data not shown).
+
+\subsection*{Differential expression suggests inactive adaptive immune responses
+in line 6}
 
@@
 upregulated genes in line 7 was much greater compared to line 6
-(Figure~\ref{degenes_venn})~\cite{lee1983ontogeny}.  Interestingly, some genes
+(Figure~\ref{degenes_venn}).  Interestingly, some genes
 that were differentially expressed in both lines were regulated in the opposite
@@
 
-Some DE genes have been reported from previous microarray studies were also
-found to be differentially expressed in this study.  For example, B6.1 (Bu-1)
-gene is known to be down-regulated approximately 2.27 fold in susceptible
-chicken with the MHC allele B\textsuperscript{19} at 4
-d.p.i~\cite{sarson2008transcriptional}.  It was also found to be down-regulated
-about 3-fold in line 7 (susceptible).  Similarly, {\em GMZA} gene reported to be
-upregulated across genetically different chickens (B$^{19}$, B$^{21}$ alleles),
-was also found to be highly upregulated.  In contrast, some genes that have been
-reported to be highly expressed in resistant chickens were downregulated in both
-lines. Those genes are {\em AMIGO2, MMP13} and {\em CLEC3B}, which were found to
-be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
-% Several genes involved in innate immune response were differentially expressed
-% in resistant chickens: \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all
-% those genes, only HSP5 was upregulated.
-Other immune genes reported to be highly expressed in susceptible chickens (line
-7$_{2}$) including {\em AVD, ART1, NOS2, CXCL13L2, MX1} and {\em
-SOCS1}~\cite{smith2011systems} were also found to be highly upregulated in both
-line 6 and 7.  However, our results show the similar expression patterns for
-{\em IL6} and {\em IL18}, which were only upregulated in line 7 in early stage
-of infection (3-5 d.p.i).
-
-In contrast, {\em IL15} has been reported non differential-expressed between
-control and infected chickens in line 6$_{1}$ and
-7$_{2}$~\cite{kaiser2003differential}; however, it was only upregulated in line
-7 in our study.  Expression of {\em IL15} is induced by {\em TLR9}, which binds
-to non-methylated CpG residues present in genomes of many DNA viruses, including
-herpes simplex virus.  This cytokine auto-regulates the expression of {\em
-CD40}, which is a transmembrane receptor required for activation of macrophages
-by CD4 T cells.  Consequently, {\em CD40} was only upregulated in line 7 (data
-not shown).
-
 \subsection*{Functional analysis of differential-expressed genes}
@@
 cytokine-cytokine receptor interaction, intestinal immune network for IgA
-production, and cell adhesion molecules (CAMs)~\ref{line67_kegg}.  Some other
-pathways important in response to viral infection and only significantly
+production, and cell adhesion molecules (CAMs)(figure~\ref{line67_kegg}).  Some
+other pathways important in response to viral infection and only significantly
 enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling
@@
 synthesized MHC I were only upregulated in line 7 suggesting that new MHC I
-molecules were actively produced and expressed on cell surface.
-% (the minor class I is upregulated but the major class I is downregulated)
-Furthermore, gene ontology analysis of biological processes (GO:BP) shows that
-categories involved in both adaptive and innate immune responses were enriched
-in line 7 (supplementary material).  On the other hand, only categories involved
-in innate immune responses were enriched in line 6.  In addition, enrichment of
-apoptosis pathway in line 7 indicates that the programmed cell death could be
-induced by CTL response to eliminate the viruses.
+molecules were actively produced.  Furthermore, gene ontology analysis of
+biological processes (GO:BP) shows that categories involved in both adaptive and
+innate immune responses were enriched in line 7 (supplementary material).  On
+the other hand, only categories involved in innate immune responses were
+enriched in line 6.  In addition, enrichment of apoptosis pathway in line 7
+indicates that the programmed cell death could be induced by CTL response to
+eliminate the viruses.
 
@@
 
-\subsection*{Differential exon usage (DEU)}
+\subsection*{Genes with differential exon usage (DEU) in response to MDV
+infection can be divided into four groups based on their patterns of expression}
 
@@
 alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
-
-\subsubsection*{DEU genes in response to MDV}
-
 Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20
@@
 
-\subsubsection*{Roles of {\em LFA-1} and actin cytoskeleton pathway in T cells activation}
+\subsubsection*{Roles of {\em LFA-1} and actin cytoskeleton pathway in T cells
+activation}
 
@@
 
+Using exon-centric approach, we identified several genes involved in actin
+cytoskeleton, especially {\em ITGB2}, that expressed different ratios of
+alternative isoforms between lines 6 and 7.  As described above, {\em ITGB2}
+encodes $\beta$-subunit of integrins including LFA-1, which exclusively
+expressed in lymphocytes and plays a major role in lymphoproliferation, antigen
+presentation, T cells activation, and cytotoxicity.  In patients with LFA-1
+deficiency (LAD), the response of lymphocytes to mitogens is
+decreased~\cite{springer1987lymphocyte}.  The decrease in responsiveness to
+mitogens have been shown to correlate with resistance to MD by Lee and
+Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
+N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
+7 and P).
+
+Even though the exact mechanism is not known, incompetency of T cells in
+response to stimuli appears to benefit resistant birds because MD virus could
+not induce T cells to proliferate and cause them to undergo neoplastic
+transformation.  It has also been suggested that the mechanism that controls
+both lymphocyte proliferation induced by MD virus and lymphocyte proliferation
+induced by immune response is the same~\cite{pazderka1975histocompatibility}.
+Therefore, it may be useful to consider a link between deficiency of lymphocytes
+in line 6 to the alternative splice form of {\em ITGB2} that is only expressed
+in line 6.  Although the exon included in the alternative splice form is
+non-coding, it could serve important functions in translation or
+posttranscriptional regulation.
+
+Interestingly, by grouping genes based on patterns of $\Psi$s, we found that
+many genes in group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em
+ZDHHC7}, are involved in cytokinesis or cell synapse, which are important for
+aggregation of LFA-1 and T cell activation as discussed above.  The $\Psi$
+values in group II are more similar within line than between lines in both
+control and infected groups, suggesting the possibility that the alternative
+splicing is controlled by a genetic mechanism.  Based on the independent
+criteria, enrichment of cytoskeletal genes in this group could not be just a
+mere coincidence, but an evidence suggesting that inefficiency of T cell
+activation in line 6 is linked to a genetic factor that regulate alternative
+splicing of {\em ITGB2} and other genes in the cytoskeleton system.
+Experimental validation of exonic SNPs provided by this study could shed some
+lights on underlying polymorphisms that contribute to MD resistance.
+
 \subsubsection*{Protein domains}
@@
 
-\subsubsection*{Prediction of {\em cis}-regulatory elements in alternative splicing exons}
+\subsubsection*{Prediction of {\em cis}-regulatory elements in alternative
+splicing exons}
 
@@
 
-\section*{Discussion}
+\section*{Conclusion}
 
 Custom gene models built from combination of gene models from {\em de novo}
-assembly, reference-based assembly, and Ensembl allowed us to identify genes and
-isoforms that might play an important role in resistance to MD.  Results from
-gene expression analysis indicate that adaptive immune responses were active
-during lytic infection in line 7.  Because only activated T cells are infected,
-elicitation of adaptive immune responses could help spread the viruses by
-recruiting and activating more T cells.  Therefore, active immune responses are
-favorable for the virus.  On the other hand, the delay of adaptive immune
-responses in line 6, in this case, could benefit the host by limiting infection
-of activated T cells.
+assembly, reference-based assembly, and ENSEMBL had allowed us to identify genes
+and isoforms that might play an important role in resistance to MD.  Results
+from gene expression analysis indicated that adaptive immune responses were
+active during lytic infection in line 7, but not in line 6.  Because only
+activated T cells are infected by MDV, we speculated that elicitation of
+adaptive immune responses could help spread the viruses by recruiting and
+activating more T cells and.  In contrast, the delay of adaptive immune
+responses could benefit the host by limiting infection of activated T cells.
 
-Using exon-centric approach, we identified several genes involved in actin
-cytoskeleton, especially {\em ITGB2}, that expressed different ratios of
-alternative isoforms between lines 6 and 7.  As described above, {\em ITGB2}
-encodes $\beta$-subunit of integrins including LFA-1, which exclusively
-expressed in lymphocytes and plays a major role in lymphoproliferation, antigen
-presentation, T cells activation, and cytotoxicity.  In patients with LFA-1
-deficiency (LAD), the response of lymphocytes to mitogens is
-decreased~\cite{springer1987lymphocyte}.  The decrease in responsiveness to
-mitogens have been shown to correlate with resistance to MD by Lee and
-Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
-N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
-7 and P).
-
-Even though the exact mechanism is not known, incompetency of T cells in
-response to stimuli appears to benefit resistant birds because MD virus could
-not induce T cells to proliferate and cause them to undergo neoplastic
-transformation.  It has also been suggested that the mechanism that controls
-both lymphocyte proliferation induced by MD virus and lymphocyte proliferation
-induced by immune response is the same~\cite{pazderka1975histocompatibility}.
-Therefore, it may be useful to consider a link between deficiency of lymphocytes
-in line 6 to the alternative splice form of {\em ITGB2} that is only expressed
-in line 6.  Although the exon included in the alternative splice form is
-non-coding, it could serve important functions in translation or
-posttranscriptional regulation.
-
-Interestingly, by grouping genes based on patterns of $\Psi$s, we found that
-many genes in group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em
-ZDHHC7}, are involved in cytokinesis or cell synapse, which are important for
-aggregation of LFA-1 and T cell activation as discussed above.  The $\Psi$
-values in group II are more similar within line than between lines in both
-control and infected groups, suggesting the possibility that the alternative
-splicing is controlled by a genetic mechanism.  Based on the independent
-criteria, enrichment of cytoskeletal genes in this group could not be just a
-mere coincidence, but an evidence suggesting that inefficiency of T cell
-activation in line 6 is linked to a genetic factor that regulate alternative
-splicing of {\em ITGB2} and other genes in the cytoskeleton system.
-Experimental validation of exonic SNPs provided by this study could shed some
-lights on underlying polymorphisms that contribute to MD resistance.
-
-To conclude, we attempted to employ custom gene models built from RNA-Seq data
-to identify underlying genetic factors contributing to MD resistance.  We
-identified a number of genes that expressed different ratios of alternative
-splice forms between lines 6 and 7.  Interestingly, several genes are involved
-in actin cytoskeleton and cytokinesis, which have not been of great interest in
-the filed of MD research.  Even though we mainly discussed the possible role of
-{\em ITGB2} in MD resistance, other genes cannot be precluded and should be a
-candidate for further investigation and experimental studies.  Moreover, a full
-mechanism of MD resistance may be highly complex and require more data from
-different stages of infection as well as a greater sequencing depth in order to
-identify all genes and isoforms involved.  To enable the study of unannotated
-gene and isoform expression, our approach of constructing gene models from
-RNA-Seq should be iteratively used to extend the Ensembl gene models.
+To elucidate the molecular mechanism of MD resistance, we investigated
+differential isoforms expression between lines 6 and 7 and identified a number
+of genes that could be responsible for difference in immune responses. Notably,
+several genes are involved in actin cytoskeleton and cytokinesis, which are
+important for the functions of lymphocytes and immune cells but have not been of
+great interest in the filed of MD research.  Even though we mainly discussed the
+possible role of {\em ITGB2} in MD resistance, other genes cannot be precluded
+and should be a candidate for further investigation and experimental studies.
+Moreover, a full mechanism of MD resistance is highly complex and more data from
+different stages of infection as well as a greater sequencing depth will be
+required to identify all genes and isoforms involved.  To enable the study of
+unannotated gene and isoform expression, our approach of constructing gene
+models from RNA-Seq should be iteratively used to extend the ENSEMBL gene models
+to construct more complete gene models.
 

### 2014-07-16 paper.tex
reformatted other sections

@@
 resistance or susceptibility to MD.  For example, chickens with MHC allele
-B$^{21}$ are highly resistant in contrast to chickens with B$^{16}$ allele which
-are highly susceptible.  ADOL lines 6 and 7, both share (B$^{2}$) allele, yet
-exhibit different phenotypic responses to MD; e.g., challenge with JM/102W
-strain with result in 0 and 100\% MD incidence for lines 6 and 7, respectively.
-Thus, the major unanswered questions are what genetic factors, especially those
-that are non-MHC, contribute to susceptibility and resistance of the disease and
-what is the main contributing mechanism.
+B\textsuperscript{21} are highly resistant in contrast to chickens with
+B\textsuperscript{19} allele which are highly susceptible.  ADOL lines 6 and 7,
+both share (B\textsuperscript{2}) allele, yet exhibit different phenotypic
+responses to MD; e.g., challenge with JM/102W strain with result in 0 and 100\%
+MD incidence for lines 6 and 7, respectively.  Thus, the major unanswered
+questions are what genetic factors, especially those that are non-MHC,
+contribute to susceptibility and resistance of the disease and what is the main
+contributing mechanism.
 
@@
 microarray and RNA-Seq methods in order to identify non-MHC genes that
-contribute to resistance to MD~\cite{}.  However, none of the studies have
-investigated differential expression of alternative isoforms, which are known to
-play a significant role in many biological events including immune responses.
-In addition, studies have shown that isoform expression levels can provide
-better signatures for some diseases~\cite{zhang2013isoform}.  Changes of isoform
-expression levels are governed partly by two types of {\em cis-}regulatory
-elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer (ESS), which
-are located within an exon sequence.  A number of sequence motifs of ESE and ESS
-have been identified in human and some other organisms and could be predicted
-{\em in silico}.  Mutations that disrupt or create those motifs could alter
-splicing patterns leading to aberrant alternative splicing.  A number of
-disease-associated single-nucleotide polymorphsims in coding regions (cSNPs)
-that affect ESEs and ESSs have been well characterized~\cite{blencowe2000exonic,
-wang2007splicing}.
+contribute to resistance to
+MD~\cite{sarson2008transcriptional,morgan2001induction,vallejo1998genetic,yonash1999high,bumstead1998genomic}.
+However, none of the studies have investigated differential expression of
+alternative isoforms, which are known to play a significant role in many
+biological events including immune responses.  In addition, studies have shown
+that isoform expression levels can provide better signatures for some
+diseases~\cite{zhang2013isoform}.  Changes of isoform expression levels are
+governed partly by two types of {\em cis-}regulatory elements: Exon Splicing
+Enhancer (ESE) and Exon Splicing Silencer (ESS), which are located within an
+exon sequence.  A number of sequence motifs of ESE and ESS have been identified
+in human and some other organisms and could be predicted {\em in silico}.
+Mutations that disrupt or create those motifs could alter splicing patterns
+leading to aberrant alternative splicing.  A number of disease-associated
+single-nucleotide polymorphsims in coding regions (SNPs) that affect ESEs and
+ESSs have been well characterized~\cite{blencowe2000exonic, wang2007splicing}.
 %In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing
@@
 differential-expressed genes and isoforms that may contribute to resistance to
-MD as well as cSNPs that can potentially affect isoform expression levels.
+MD as well as SNPs that can potentially affect isoform expression levels.
 
@@
 
-\subsection*{Differential expression suggests inactive adaptive immune responses
-in line 6}
+\subsection*{Differential gene expression indicates active immune responses to
+ongoing lytic infection in line 7}
 
@@
 
-\subsection*{Functional analysis of differential-expressed genes}
+\subsection*{Functional analysis of differential-expressed genes indicates
+inactive adaptive immune responses in line 6}
 
@@
 it is beneficial for the host to restrain such contact.  However, it is not
-clear how chickens in line 6 control the lytic infection of MDV.
-% Two mechanisms have been speculated to contribute to MD resistance.
-% First innate immune responses could be highly effective and could activate
-% strong adaptive immune responses that rapidly control viral replication and force the viruses to undergo latent phase.
-% Second, innate immune resposnes itself could be highly effective in limiting viral replication~\cite{smith2011systems}.
+clear how chickens in line 6 control the lytic infection of MDV.  Two mechanisms
+have been speculated to contribute to MD resistance.  First innate immune
+responses could be highly effective and could activate strong adaptive immune
+responses that rapidly control viral replication and force the viruses to
+undergo latent phase.  Second, innate immune resposnes itself could be highly
+effective in limiting viral replication~\cite{smith2011systems}.
 
@@
 pro-adhesive and anti-adhesive molecules.  This protein is involved in
-cell-to-cell contact, cell trafficking, and cancer progression~\cite{}.
+cell-to-cell contact, cell trafficking, and cancer
+progression~\cite{nielsen2009role,somasiri2004overexpression}.
 
@@
 
-\subsubsection*{Roles of {\em LFA-1} and actin cytoskeleton pathway in T cells
+\subsection*{Roles of {\em LFA-1} and actin cytoskeleton in T cells
 activation}
 
-Integrins are a special kind of receptors that transmit signals bidirectionally
-across the cell membrane.  They are heterodimeric composed of an $\alpha$
-(large) and a $\beta$ (small) subunits~\cite{wang2010immunopathologies}.
-$\beta_{2}$ (CD18) subunit encoded by {\em ITGB2} is expressed on lymphocytes
-and antigen presenting cells (APCs) as a component of {\em LFA-1} and {\em CR3}
-receptors.  LFA-1 binds to its ligand ICAM-1 to help form a synapse that bring
-APC and T cell together to initiate antigen presentation leading to T cell
-activation~\cite{dustin2000immunological}.  Absence of LFA-1 leads to impaired
-functions of lymphocytes in proliferation and tumor
-rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.  Mutations in {\em
-ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency
+By grouping genes based on patterns of $\Psi$s, we found that many genes in
+group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em ZDHHC7}, are
+involved in cytokinesis or cell synapse, which are important for T cell
+activation.  As described above, {\em ITGB2} encodes $\beta$-subunit of
+integrins including LFA-1, which exclusively expressed in lymphocytes and plays
+a major role in lymphoproliferation, antigen presentation, T cells activation,
+and cytotoxicity.  Integrins are a special kind of receptors that transmit
+signals bidirectionally across the cell membrane.  They are heterodimeric
+composed of an $\alpha$ (large) and a $\beta$ (small)
+subunits~\cite{wang2010immunopathologies}.  $\beta_{2}$ (CD18) subunit encoded
+by {\em ITGB2} is expressed on lymphocytes and antigen presenting cells (APCs)
+as a component of {\em LFA-1} and {\em CR3} receptors.  LFA-1 binds to its
+ligand ICAM-1 to help form a synapse that bring APC and T cell together to
+initiate antigen presentation leading to T cell
+activation~\cite{dustin2000immunological}.
+
+Absence of LFA-1 leads to impaired functions of lymphocytes in proliferation and
+tumor rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.  Mutations in
+{\em ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency
 (LAD-1), an autosomal-recessive inherited disease found in few families.  The
@@
 functions, lack of accumulation to the site of infection and recurrent bacterial
-and fungal infection~\cite{springer1987lymphocyte}.
+and fungal infection~\cite{springer1987lymphocyte}.  In addition, the response
+of lymphocytes to mitogens is decreased in patients with
+LAD~\cite{springer1987lymphocyte}.  The decrease in responsiveness to mitogens
+have been shown to correlate with resistance to MD by Lee and
+Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
+N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
+7 and P).
 
@@
 other pathways that involved in immune responses (Table~\ref{tab:integrin}).
-This could be speculated that pre-mRNA splicing of these genes is co-regulated
+It could be speculated that pre-mRNA splicing of these genes is co-regulated
 by splicing regulators or some genetic factors.
 
-Using exon-centric approach, we identified several genes involved in actin
-cytoskeleton, especially {\em ITGB2}, that expressed different ratios of
-alternative isoforms between lines 6 and 7.  As described above, {\em ITGB2}
-encodes $\beta$-subunit of integrins including LFA-1, which exclusively
-expressed in lymphocytes and plays a major role in lymphoproliferation, antigen
-presentation, T cells activation, and cytotoxicity.  In patients with LFA-1
-deficiency (LAD), the response of lymphocytes to mitogens is
-decreased~\cite{springer1987lymphocyte}.  The decrease in responsiveness to
-mitogens have been shown to correlate with resistance to MD by Lee and
-Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
-N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
-7 and P).
+\subsection*{Prediction of functional domains of splice forms of genes in the
+actin cytoskeleton pathway}
+
+To predict the function of the alternative splice forms of genes in actin
+cytoskeleton pathway, transcript sequences were translated to protein sequences
+by ESTscan~\cite{iseli1999estscan}.  Protein sequences were then searched for
+annotated protein domains using the standalone version of InterPro
+Scan~\cite{quevillon2005interproscan}.  Besides {\em ITGB2}, other genes have
+alternative exons located in coding regions and could potentially affect
+functional protein domains in some ways.  The exon with alternative 3$\prime$
+splice site of {\em RAC3} encode part of a protein domain identified as small
+GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).  Rac3 is
+highly homologous to Rac1 and has been reported to possess the ability of
+promoting membrane ruffling, transformation, activation of c-Jun transcriptional
+activity and a co-activator of NF$\kappa$B~\cite{werbajh2000rac}.  Activated Rac
+also regulates production of superoxide in neutrophils and macrophages.
+
+The alternative exon of {\em PFN2} seems to disrupt the coding sequence that
+encode profilin domain (Pfam:PF00235).  The profilin domain is essential for
+almost all organisms and its functions include regulating actin polymerization,
+controlling complex network of molecular interaction and transmitting signals
+from small-GTPase pathway.  It also binds to Rac effector molecules and a number
+of other ligands~\cite{witke2004role}.  The skipped exon of {\em PPP1R12A}
+encodes part of a protein domain annotated as protein phosphatase 1
+(PIRSF:PIRSF038141).  The functional effects of alternative splicing on these
+functional domains have to be further investigated by simulation or an
+experiment.
 
@@
 
-Interestingly, by grouping genes based on patterns of $\Psi$s, we found that
-many genes in group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em
-ZDHHC7}, are involved in cytokinesis or cell synapse, which are important for
-aggregation of LFA-1 and T cell activation as discussed above.  The $\Psi$
-values in group II are more similar within line than between lines in both
-control and infected groups, suggesting the possibility that the alternative
-splicing is controlled by a genetic mechanism.  Based on the independent
-criteria, enrichment of cytoskeletal genes in this group could not be just a
-mere coincidence, but an evidence suggesting that inefficiency of T cell
-activation in line 6 is linked to a genetic factor that regulate alternative
-splicing of {\em ITGB2} and other genes in the cytoskeleton system.
-Experimental validation of exonic SNPs provided by this study could shed some
-lights on underlying polymorphisms that contribute to MD resistance.
-
-\subsubsection*{Protein domains}
-
-To understand the role of the alternative splice forms on the function of these
-genes, transcripts were translated to protein sequences by
-ESTscan~\cite{iseli1999estscan}.  Protein sequences were then searched for
-annotated protein domains using the standalone version of InterPro
-Scan~\cite{quevillon2005interproscan}.  Besides {\em ITGB2}, other genes have
-alternative exons located in coding regions and could potentially affect
-functional protein domains in some ways.  The exon with alternative 3$\prime$
-splice site of {\em RAC3} encode part of a protein domain identified as small
-GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).  Rac3 is
-highly homologous to Rac1 and has been reported to possess the ability of
-promoting membrane ruffling, transformation, activation of c-Jun transcriptional
-activity and a co-activator of NF$\kappa$B~\cite{,werbajh2000rac}.  Activated
-Rac also regulates production of superoxide in neutrophils and macrophages.
-
-The alternative exon of {\em PFN2} seems to disrupt the coding sequence that
-encode profilin domain (Pfam:PF00235).  The profilin domain is essential for
-almost all organisms and its functions include regulating actin polymerization,
-controlling complex network of molecular interaction and transmitting signals
-from small-GTPase pathway.  It also binds to Rac effector molecules and a number
-of other ligands~\cite{witke2004role}.  The skipped exon of {\em PPP1R12A}
-encodes part of a protein domain annotated as protein phosphatase 1
-(PIRSF:PIRSF038141).  The functional effects of alternative splicing on these
-functional domains have to be further investigated by simulation or an
-experiment.
-
 % Based on our gene model, the skipped exon of \textit{ITGB2} is located in the
@@
 
-\subsubsection*{Prediction of {\em cis}-regulatory elements in alternative
-splicing exons}
+\subsection*{Prediction of {\em cis}-regulatory elements in alternative
+splicing exons of genes in group II}
 
@@
 this group were relatively stable within line, but were significantly different
-between lines.  Investigation of nucleotide differences within exons of line 6
-and 7 could reveal a possible role of cSNPs in regulating alternative splicing
-in this group.  We obtained a sequence of an alternative exon from line 6 and
-used Human Splicing Finder (HSF) to determine whether SNPs from line 7 could
-alter predicted ESEs or ESSs.  Results from some genes involved in cytokinesis
-are discussed in this section.  Exonic SNPs from line 6 and 7 are listed in
+between lines.  Investigation of nucleotide differences within exons of lines 6
+and 7 could reveal a possible role of SNPs in regulating alternative splicing in
+this group.  We obtained a sequence of an alternative exon from line 6 and used
+Human Splicing Finder (HSF) to determine whether SNPs from line 7 could alter
+predicted ESEs or ESSs.  Results from some genes involved in cytokinesis are
+discussed in this section.  Exonic SNPs from line 6 and 7 are listed in
 Table~\ref{tab:deu_snps}.
@@
 7 according to SNPs data from the genome resequencing.  Therefore, this SNP may
-not be accounted for exclusion of the exon in line 7 (Figure~\ref{itgb2}).
-
-Exon sequences of {\em PFN2} from line 6 and 7 differ at position 2,072 on cDNA
-or 23,221,934 on chromosome 9.  A small insertion of two AA nucleotides is
+not be accounted for exclusion of the exon in line 7 (Figure~\ref{itgb2}).  Exon
+sequences of {\em PFN2} from line 6 and 7 differ at position 2,072 on cDNA or
+23,221,934 on chromosome 9.  A small insertion of two AA nucleotides is
 predicted to create a new binding site for Tra2-$\beta$ splicing regulator,
@@
 browser, the exon has been found and sequenced from chicken eyes (15d
-post-hatched, EST sequence:DR424100).
-
-For {\em DYNLT1}, the skipped exon was absent in line 7, therefore, we could not
-determine the SNPs from RNA-Seq data.  However, data from DNA resequencing
-showed that there was a polymorphism at position 51,357,865 on chromosome 3.
-Replacing G with C from line 7 is predicted to slightly alter predicted binding
-sites for 9G8 as well as SRp55 exon enhancer.  However, whether the alteration
-down regulates the exon in line 7 is unclear.
+post-hatched, EST sequence:DR424100).  For {\em DYNLT1}, the skipped exon was
+absent in line 7, therefore, we could not determine the SNPs from RNA-Seq data.
+However, data from DNA resequencing showed that there was a polymorphism at
+position 51,357,865 on chromosome 3.  Replacing G with C from line 7 is
+predicted to slightly alter predicted binding sites for 9G8 as well as SRp55
+exon enhancer.  However, whether the alteration down regulates the exon in line
+7 is unclear.
 
@@
 of these two genes and other genes in this group are provided in the
-supplementary materials.
+supplementary materials.  Experimental validation of exonic SNPs provided by
+this study could shed some lights on underlying polymorphisms that contribute to
 
@@
 
-% However, to mechanism that contribute to resistance to MD could be very
-% complex and integration of global gene expression data as well as experimental
-% data from different stages of infection will be required to fully elucidate.
-% Moreover, better gene models and a reference genome will be necessary.
-
 % You may title this section "Methods" or "Models". 
@@
 % authors may use "Analysis"
-\section*{Materials and Methods}
-\subsection{Gene Models Construction}
+\section*{Materials and methods}
+
+\subsection{Sequences and quality trimming}
+
+mRNAs were extracted from spleens of control and infected chickens lines 6 and 7
+(4 d.p.i).  Sequence libraries were prepared by standard Illumina unstranded
+single- and paired-end protocols.  Library size of the paired-end datasets is
+approximately 175 bp.  Read lengths are 75 bp in both single- and paired-end
+libraries.  Reads were quality trimmed by Condetri 2.1~\cite{smeds2011condetri}
+with quality score cutoff of 30.  The first 10 bases were removed due to
+non-uniform distribution of nucleotides.
+
+\subsection*{Gene models construction}
 
 Due to lack of complete gene models for chickens, we employed two methods to
-construct gene models from RNA-Seq reads.  First, short reads were quality
-trimmed with conditri/2.1~\cite{} and assembled using Velvet/1.21~\cite{} and
-Oases/0.2.06~\cite{} to obtain long transcripts.  Assembly was done with hash
-lengths range from 21 to 31 for both local and global assembly (described in
-Gimme paper).  Poly-A tails were trimmed and low complexity transcripts were
-removed by Seqclean~\cite{}.  All transcripts were then aligned to chicken
+construct gene models from RNA-Seq reads.  First, short reads were assembled
+using Velvet/1.2.03~\cite{Zerbino:2008vu} and Oases/0.2.06~\cite{Schulz:2012je}
+to obtain long transcripts.  Assembly was done with hash lengths range from 21
+to 31 for both local and global assembly (described in Gimme paper).  Poly-A
+tails were trimmed and low complexity transcripts were removed by
+Seqclean~\cite{seqclean}.  All transcripts were then aligned to chicken
 reference genome (galGal4, with unplaced and random chromosomes removed) with
-BLAT~\cite{}.  Second, reads were aligned to reference genome using
-Tophat/2.0.5~\cite{} and ENSEMBLE gene models release 73 was used to guide
-reference-based assembly by Cufflinks2~\cite{}.  Alignments from BLAT and models
-from Cufflinks were then combined to construct gene models by Gimme (manuscript
-in preparation).
+BLAT~\cite{Kent:2002tv}.  Second, reads were aligned to reference genome using
+Tophat/2.0.9~\cite{Trapnell:2009dp} and ENSEMBLE gene models release 73 was used
+to guide reference-based assembly by Cufflinks2~\cite{Trapnell:2010kd}.
+Alignments from BLAT and models from Cufflinks were then combined to construct
+gene models by Gimme (manuscript in preparation).
 
-\subsection{Differential Gene Expression and Gene Ontology}
+\subsection*{Differential gene expression and gene ontology}
 
@@
 
-\subsection{Differential Splicing Event}
+\subsection*{Differential splicing event}
 
@@
 
-\subsection{Variant Calling and {\em In Silico} Splicing Analysis} Variants were
-called using mpileup command from SAMTools/0.1.18~\cite{} and BCFTools~\cite{}.
-Only variants with quality score $\ge20$ were used for mutation analyses.  Exon
+\subsection*{Variant calling and {\em in silico} splicing analysis}
+
+Variants were called using mpileup command from
+SAMTools/0.1.18~\cite{li2009sequence} and BCFTools~\cite{bcftools}.  Only
+variants with quality score $\ge20$ were used for mutation analyses.  Exon
 enhancers and suppressors were predicted using Human Splicing Finder web
-portal~\cite{}.  Human default parameter settings were used in all analyses.
-The following regulatory sequences were used to determined the effect of
-variants: HSF integrated matrices for serine/arginine-rich proteins (SRp40,
-SC35, SF2/ASF, SF2/ASF, IgM/BRCA1, and SRp55), exonic splicing enhancer (ESE),
-RESCUE-ESE hexamer (RESCUE-ESE), putative 8-mer ESEs (PESEs) and putative 8-mers
-exonic splicing silencers (PESSs), exon-identity and intron-identiry elements
-(EIEs and IIEs), hetronuclear ribonucleoprotein-binding motifs, and Fas exonic
-splicing silencers.
+portal~\cite{desmet2009human}.  Human default parameter settings were used in
+all analyses.  The following regulatory sequences were used to determined the
+effect of variants: HSF integrated matrices for serine/arginine-rich proteins
+(SRp40, SC35, SF2/ASF, SF2/ASF, IgM/BRCA1, and SRp55), exonic splicing enhancer
+(ESE), RESCUE-ESE hexamer (RESCUE-ESE), putative 8-mer ESEs (PESEs) and putative
+8-mers exonic splicing silencers (PESSs), exon-identity and intron-identiry
+elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding motifs, and Fas
+exonic splicing silencers.
 
-\subsection{Protein domains search}
+\subsection*{Protein domains search}
 

### 2014-07-25 paper.tex
updated text, rearranged figures and etc

@@
 \section*{Abstract}
-Marek's disease (MD) is caused by highly oncogenic Marek's disease virus (MDV).
-Understanding the genetic basis for non-MHC resistance would be of both
-fundamental and applied importance.  In this study, using computational
-approaches of RNA-Seq data, we identified differentially expressed genes and
-isoforms in response to MDV infection from highly inbred chickens lines that are
-MD resistant and susceptible, and performed pathway and functional analysis.
-Results show that expression of genes and isoforms involved in cell cytoskeleton
-and adhesion molecules in line 6 may contribute to resistance of the disease by
-limiting T cell activation, which in turn prevents the spread of MDV to
-activated T cells.  In addition, we identified exonic single nucleotide
-polymorphisms (SNPs) between lines 6 and 7 predicted their role in regulation of
-alternative splicing.  Based on our results, we speculate that alternative
-splice forms of cell adhesion molecules and cytoskeleton play and important role
-in controlling T cell activation in line 6.
+Marek's disease (MD) is caused by highly oncogenic Marek's
+disease virus (MDV).  Understanding the genetic basis for non-MHC
+resistance would be of both fundamental and applied importance.
+In this study, using computational approaches of RNA-Seq data, we
+identified differentially expressed genes and isoforms in
+response to MDV infection from highly inbred chickens lines that
+are MD resistant (line 6), susceptible (line 7), and performed
+pathway and functional analysis.  Results show that expression of
+genes and isoforms involved in cell cytoskeleton and adhesion
+molecules in the resistant line may contribute to resistance of
+the disease by limiting T cell activation, which in turn prevents
+the spread of MDV to activated T cells.  In addition, we
+identified exonic single nucleotide polymorphisms (SNPs) between
+lines predicted role in regulation of alternative splicing.
+Based on our results, we speculate that alternative splice forms
+of cell adhesion molecules and cytoskeleton play and important
+role in controlling T cell activation in the resistant line.
 
@@
 
-Marek's disease (MD) is an economically significant chicken disease that affects
-the poultry industry worldwide with estimated annual loss of \$2 billion loss
-has been reported from outbreaks~\cite{morrow2004marek}.  The disease is caused
-by the highly oncogenic Marek's disease virus (MDV), an alphaherpesvirus that
-induces T-cell lymphomas in susceptible birds.  Vaccination is the primary
-control measure, which is effective in reducing incidence of tumor formation.
-However, since MD vaccines are not sterilizing, they do not prevent infection or
-horizontal spread of the virus.  As a consequence, MDV field strains that
-overcome vaccinal protection have arisen repeatedly over time.  Therefore, there
-is a need for sustainable alternative controls measures, such as improving
+Marek's disease (MD) is an economically significant chicken
+disease that affects the poultry industry worldwide with
+estimated annual of \$2 billion~\cite{morrow2004marek}.  The
+disease is caused by the highly oncogenic Marek's disease virus
+(MDV), an alphaherpesvirus that induces T-cell lymphomas in
+susceptible birds.  Vaccination is the primary control measure,
+which is effective in reducing incidence of tumor formation.
+However, since MD vaccines are not sterilizing, they do not
+prevent infection or horizontal spread of the virus.  As a
+consequence, MDV field strains that overcome vaccinal protection
+have arisen repeatedly over time.  Therefore, there is a need for
+sustainable alternative controls measures, such as improving
 genetic resistance.
 
-Many studies have reported strong associations between MHC alleles and
-resistance or susceptibility to MD.  For example, chickens with MHC allele
-B\textsuperscript{21} are highly resistant in contrast to chickens with
-B\textsuperscript{19} allele which are highly susceptible.  ADOL lines 6 and 7,
-both share (B\textsuperscript{2}) allele, yet exhibit different phenotypic
-responses to MD; e.g., challenge with JM/102W strain with result in 0 and 100\%
-MD incidence for lines 6 and 7, respectively.  Thus, the major unanswered
-questions are what genetic factors, especially those that are non-MHC,
-contribute to susceptibility and resistance of the disease and what is the main
-contributing mechanism.
-
-In the past decades, significant efforts have been spent to study variations in
-a global gene expression between resistant and susceptible birds using
-microarray and RNA-Seq methods in order to identify non-MHC genes that
-contribute to resistance to
+Many studies have reported strong associations between MHC
+alleles and resistance or susceptibility to MD.  For example,
+chickens with MHC allele B\textsuperscript{21} are highly
+resistant in contrast to chickens with B\textsuperscript{19}
+allele which are highly susceptible.  ADOL lines 6 and 7, both
+share (B\textsuperscript{2}) allele, yet exhibit different
+phenotypic responses; e.g., challenge with JM/102W strain result
+in 0 and 100\% MD incidence for lines 6 and 7, respectively.
+Thus, the major unanswered questions are what genetic factors,
+especially those that are non-MHC, contribute to susceptibility
+and resistance of the disease and what is the main contributing
+mechanism.
+
+In the past decades, significant efforts have been to study
+variations in a global gene expression between resistant and
+susceptible birds using microarray and RNA-Seq methods in order
+to identify non-MHC genes that contribute to resistance to
 MD~\cite{sarson2008transcriptional,morgan2001induction,vallejo1998genetic,yonash1999high,bumstead1998genomic}.
-However, none of the studies have investigated differential expression of
-alternative isoforms, which are known to play a significant role in many
-biological events including immune responses.  In addition, studies have shown
-that isoform expression levels can provide better signatures for some
-diseases~\cite{zhang2013isoform}.  Changes of isoform expression levels are
-governed partly by two types of {\em cis-}regulatory elements: Exon Splicing
-Enhancer (ESE) and Exon Splicing Silencer (ESS), which are located within an
-exon sequence.  A number of sequence motifs of ESE and ESS have been identified
-in human and some other organisms and could be predicted {\em in silico}.
-Mutations that disrupt or create those motifs could alter splicing patterns
-leading to aberrant alternative splicing.  A number of disease-associated
-single-nucleotide polymorphsims in coding regions (SNPs) that affect ESEs and
-ESSs have been well characterized~\cite{blencowe2000exonic, wang2007splicing}.
-%In fact, 15\% of mutations that cause genetic diseases affect pre-mRNA splicing
-%via this mechanism.
-Therefore, variations in isoform expression could lead to identification of SNPs
-that underlie genetic resistance to MD.  In this article, we reported
-differential-expressed genes and isoforms that may contribute to resistance to
-MD as well as SNPs that can potentially affect isoform expression levels.
+However, none of the studies have investigated differential
+expression of alternative isoforms, which are known to play a
+significant role in many biological events including immune
+responses.  In addition, studies have shown that isoform
+expression levels can provide better signatures for some
+diseases~\cite{zhang2013isoform}.  Changes of isoform expression
+levels are governed partly by two types of {\em cis-}regulatory
+elements: Exon Splicing Enhancer (ESE) and Exon Splicing Silencer
+(ESS), which are located within an exon sequence.  A number of
+sequence motifs of ESE and ESS have been identified in human and
+some other organisms and could be predicted {\em in silico}.
+Mutations that disrupt or create those motifs could alter
+splicing patterns leading to aberrant alternative splicing.  A
+number of disease-associated single-nucleotide polymorphsims in
+coding regions (SNPs) that affect ESEs and ESSs have been well
+characterized~\cite{blencowe2000exonic, wang2007splicing}.
+%In fact, 15\% of mutations that cause genetic diseases affect
+%pre-mRNA splicing via this mechanism.
+Therefore, variations in isoform expression could lead to
+identification of SNPs that underlie genetic resistance to MD.
+In this article, we reported differential-expressed genes and
+isoforms that may contribute to resistance to MD as well as SNPs
+that can potentially affect isoform expression levels.
 
@@
 
-To extensively study gene and isoform expression, we incorporated ENSEMBL gene
-models release 73 with {\em de novo} and a reference-guided transcriptome
-assembly to build custom gene models.  The models, therefore, include both
-ENSEMBL annotated transcripts and putative genes and isoforms.  The advantage of
-using custom gene models is it allows an investigation of unannotated genes and
-isoforms, which is necessary for in-depth study of an immune system.
-
-Some DE genes have been reported from previous microarray studies were also
-found to be differentially expressed in this study.  For example, B6.1 (Bu-1)
-gene is known to be down-regulated approximately 2.27 fold in susceptible
-chicken with the MHC allele B\textsuperscript{19} at 4
-d.p.i~\cite{sarson2008transcriptional}.  It was also found to be down-regulated
-about 3-fold in line 7 (susceptible).  Similarly, {\em GMZA} gene reported to be
-upregulated across genetically different chickens (B\textsuperscript{19},
-B\textsuperscript{21} alleles),
-was also found to be highly upregulated.  In contrast, some genes that have been
-reported to be highly expressed in resistant chickens were downregulated in both
-lines. Those genes are {\em AMIGO2, MMP13} and {\em CLEC3B}, which were found to
-be downregulated more than 2-fold~\cite{sarson2008transcriptional}.
-% Several genes involved in innate immune response were differentially expressed
-% in resistant chickens: \textit{DDT, NMU, VIPR2} and \textit{HSP5}. Of all
-% those genes, only HSP5 was upregulated.
-Other immune genes reported to be highly expressed in susceptible chickens (line
-7) including {\em AVD, ART1, NOS2, CXCL13L2, MX1} and {\em
-SOCS1}~\cite{smith2011systems} were also found to be highly upregulated in both
-line 6 and 7.  However, our results show the similar expression patterns for
-{\em IL6} and {\em IL18}, which were only upregulated in line 7 in early stage
-of infection ($3-5$ d.p.i).
-
-In contrast, {\em IL15} has been reported non differential-expressed between
-control and infected chickens in line 6 and 7~\cite{kaiser2003differential};
-however, it was only upregulated in line 7 in our study.  Expression of {\em
-IL15} is induced by {\em TLR9}, which binds to non-methylated CpG residues
-present in genomes of many DNA viruses, including herpes simplex virus.  This
-cytokine auto-regulates the expression of {\em CD40}, which is a transmembrane
-receptor required for activation of macrophages by CD4 T cells.  Consequently,
-{\em CD40} was only upregulated in line 7 (data not shown).
-
-\subsection*{Differential gene expression indicates active immune responses to
-ongoing lytic infection in line 7}
-
-Many genes were found to be differentially expressed (DE) between control and
-infected chickens in lines 6 and line 7.  While the number of unique
-downregulated genes in both lines was approximately equal, the number of unique
-upregulated genes in line 7 was much greater compared to line 6
-(Figure~\ref{degenes_venn}).  Interestingly, some genes
-that were differentially expressed in both lines were regulated in the opposite
-direction (Table~\ref{tab:opposite}).  Among genes downregulated in line 6 but
-upregulated in line 7 were {\em LL} (lung lectin) and {\em SFTPA1}, which encode
-a calcium-dependent C-type lectin and a lung surfactant protein respectively.
-Both molecules are important in innate
-immunity~\cite{hogenkamp2008chicken,kingma2006defense}.  {\em LIMS1} is involved
-in cell differentiation and proliferation and {\em PPARG} is a suppressor of
-{\em NF$\kappa$B}-mediated proinflamatory response.  On the other hand, nearly
-all genes upregulated in line 6 but downregulated in line 7 are involved in cell
-survival such as mRNAs splicing, cell growth, and protein synthesis, except CD7
-whose function is involved in T cell-B cell interaction.  This difference
-suggests that even at this stage of infection in line 6, the lytic phase could
-be subsided, therefore, only genes important for survival of cell are
-upregulated in line 6.  In comparison, the lytic phase in line 7 may still
-continue and as a result, genes involved in immune responses are still
-upregulated.  In addition, type I interferon ({\em IFN-$\gamma$} and {\em
-IFN-$\beta$}) as well as {\em INF-$\alpha$3} were found to be highly upregulated
-in infected chickens in both lines 6 and 7 (Table~\ref{tab:cytokines}).
-However, expressions of genes encoding their corresponding receptors were not
-different in line 6, but upregulated in line 7.  This could also reflect the
-ongoing immune responses in line 7.
-
-\subsection*{Functional analysis of differential-expressed genes indicates
-inactive adaptive immune responses in line 6}
-
-To determine pathways that were perturbed during the infection, data were
-analyzed by GOSeq, which accounted for gene lengths bias unique for RNA-Seq
-data~\cite{young2010method}.  Significantly perturbed pathways (FDR $< 0.1$)
-from both lines that involved in immune response include TLR signaling pathway,
-cytokine-cytokine receptor interaction, intestinal immune network for IgA
-production, and cell adhesion molecules (CAMs)(figure~\ref{line67_kegg}).  Some
-other pathways important in response to viral infection and only significantly
-enriched in line 7 include phagosome, apoptosis, RIG-I-like receptor signaling
-pathway, NOD-like receptor signaling pathway, and lysosome.
-Figure~\ref{kegg_phagosome} is a pictorial example of differentially expressed
-genes in the phagosome pathway.  Although MHC I gene ({\em BF1}) was
-differentially expressed in both lines, other genes involved in expressing newly
-synthesized MHC I were only upregulated in line 7 suggesting that new MHC I
-molecules were actively produced.  Furthermore, gene ontology analysis of
-biological processes (GO:BP) shows that categories involved in both adaptive and
-innate immune responses were enriched in line 7 (supplementary material).  On
-the other hand, only categories involved in innate immune responses were
-enriched in line 6.  In addition, enrichment of apoptosis pathway in line 7
-indicates that the programmed cell death could be induced by CTL response to
-eliminate the viruses.
-
-At this stage of infection, our results suggest that lytic infection of MDV
-stimulates both innate and adaptive immune responses, which leads to activation
-of T cells in line 7.  Only activated T cells are infected by MDV, therefore,
-the lytic phase could facilitate the spread of the viruses by enhancing
-expansion of activated T cells.  Due to cell-associate nature of MDV, the
-viruses transfer to T cells via cell-to-cell contact between B cells and T cells
-during antigen presentation or B cells activation by T helper cells.  Therefore,
-it is beneficial for the host to restrain such contact.  However, it is not
-clear how chickens in line 6 control the lytic infection of MDV.  Two mechanisms
-have been speculated to contribute to MD resistance.  First innate immune
-responses could be highly effective and could activate strong adaptive immune
-responses that rapidly control viral replication and force the viruses to
-undergo latent phase.  Second, innate immune resposnes itself could be highly
-effective in limiting viral replication~\cite{smith2011systems}.
-
-\subsection*{Genes with differential exon usage (DEU) in response to MDV
-infection can be divided into four groups based on their patterns of expression}
-
-The immune system is isoform-rich and many genes express different isoforms with
-distinctive functions in response to stimuli such as stress, chemicals and
-infection.  Changes in expression of splice forms of immune related genes have
-been reported to be associated with increase susceptibility and poor prognosis
-of diseases~\cite{lynch2004consequences}.  Studying differential isoform
-expression could therefore shed lights into inherent differences between lines 6
-and 7 that confer resistance or susceptibility to MD.
-
-In the past decades, microarray technology has been used to study gene and
-isoform expressions in many studies, but its sensitivity for detection of
-structurally similar isoforms is low, and known or predicted annotations are
-required to design probes~\cite{kane2000assessment}.  Although RNA-Seq method
-can provide a reliable estimate of an exon expression compared to
-microarray~\cite{pan2008deep} and is not constrained to those limitations,
-studying isoform expressions using RNA-Seq is still not straightforward because
-of the short read lengths.  Reads from current Illumina technology are generally
-not long enough to span across all exons in an isoform.  In most cases, only
-exons in close proximity are covered by the same read, which makes it difficult
-to accurately predict a full structure of the isoform.  In addition, some genes
-are fused due to overlapping untranslated regions (UTRs), which can also result
-in erroneous predicted isoform structures.
-
-Due to those issues, it is not feasible to accurately estimate expression of
-isoforms, especially when gene annotation is constructed from {\em de novo}
-assembly~\cite{trapnell2013differential}.  To avoid these issues, we chose to
-study exon expression instead of isoform expression.  Using MISO with
-exon-centric method, only reads spanning across a few exons are used and only
-exons involved in a splicing event are examined.  The expression of exon
-inclusion is calculated as Percent Spliced In (Psi or $\Psi$), which can be used
-to infer the portion of transcripts that include the exon in each
-sample~\cite{Katz:2010iv}.  In this study, we investigated the three most common
-alternative splicing events in vertebrates, which are skipped exons (SE), an
-alternative $3\prime$ (A3SS) and $5\prime$ (A5SS) splice site.
-Lists of DEU genes from line 6 that show difference in $\Psi$ greater than 0.20
-when compared to line 7 in infected chickens are shown in
+To extensively study gene and isoform expression, we incorporated
+ENSEMBL gene models release 73 with {\em de novo} and a
+reference-guided transcriptome assembly to build custom gene
+models.  The models, therefore, include both ENSEMBL annotated
+transcripts and putative genes and isoforms.  The advantage of
+using custom gene models is it allows an investigation of
+unannotated genes and isoforms, which is necessary for in-depth
+study of an immune system.
+
+Some DE genes have been reported from previous microarray studies
+were also found to be differentially expressed in this study.
+For example, B6.1 (Bu-1) is known to be down-regulated
+approximately 2.27 fold in susceptible chickens with the MHC
+allele B\textsuperscript{19} at 4 days post infection
+(d.p.i)~\cite{sarson2008transcriptional}.  It was also found to
+be down-regulated about 3-fold in the susceptible line in our
+study.  Similarly, {\em GMZA} reported to be upregulated across
+genetically different chickens (B\textsuperscript{19},
+B\textsuperscript{21} alleles), was also found to be highly
+upregulated.  In contrast, some genes that have been reported to
+be highly expressed in resistant chickens were downregulated in
+both lines. Those genes are {\em AMIGO2, MMP13} and {\em CLEC3B},
+which were found to be downregulated more than
+2-fold~\cite{sarson2008transcriptional}.
+% Several genes involved in innate immune response were
+% differentially expressed in resistant chickens: \textit{DDT,
+% NMU, VIPR2} and \textit{HSP5}. Of all those genes, only HSP5
+% was upregulated.
+Other immune genes reported to be highly expressed in susceptible
+chickens including {\em AVD, ART1, NOS2, CXCL13L2, MX1} and {\em
+SOCS1}~\cite{smith2011systems} were also found to be highly
+upregulated in both lines.  However, our results show the similar
+expression patterns for {\em IL6} and {\em IL18}, which were only
+upregulated in the susceptible line in early stage of infection
+($3-5$ d.p.i).
+
+In contrast, {\em IL15} has been reported non
+differential-expressed between control and infected chickens in
+both lines~\cite{kaiser2003differential}; however, it was only
+upregulated in the susceptible line.  Expression of {\em IL15} is
+induced by {\em TLR9}, which binds to non-methylated CpG residues
+present in genomes of many DNA viruses, including herpes simplex
+virus.  This cytokine auto-regulates the expression of {\em
+CD40}, which is a transmembrane receptor required for activation
+of macrophages by CD4 T cells.  Consequently, {\em CD40} was only
+upregulated in the susceptible line (data not shown).
+
+\subsection*{Differential gene expression indicates active immune
+responses to ongoing lytic infection in the susceptible line}
+
+Many genes were found to be differentially expressed (DE) between
+control and infected chickens in both lines.  While the number of
+unique downregulated genes in both lines was approximately equal,
+the number of unique upregulated genes in the susceptible line
+was much greater compared to the resistant line
+(Figure~\ref{degenes_venn}).  Interestingly, some genes that were
+differentially expressed in both lines were regulated in the
+opposite direction (Table~\ref{tab:opposite}).  Among genes
+downregulated in the resistant line but upregulated in the
+susceptible line were {\em LL} (lung lectin) and {\em SFTPA1},
+which encode a calcium-dependent C-type lectin and a lung
+surfactant protein respectively.  Both molecules are important in
+innate immunity~\cite{hogenkamp2008chicken,kingma2006defense}.
+{\em LIMS1} is involved in cell differentiation and proliferation
+and {\em PPARG} is a suppressor of {\em NF$\kappa$B}-mediated
+proinflamatory response.  On the other hand, nearly all genes
+upregulated in the resistant line but downregulated in the
+susceptible line are involved in cell survival such as mRNAs
+splicing, cell growth, and protein synthesis, except CD7 whose
+function is involved in T cell-B cell interaction.  This
+difference suggests that even at this stage of infection in the
+resistant line, the lytic phase could be subsided, therefore,
+only genes important for survival of cell are upregulated in the
+resistant line.  In comparison, the lytic phase in the
+susceptible line may still continue and as a result, genes
+involved in immune responses are still upregulated.  In addition,
+type I interferon ({\em IFN-$\gamma$} and {\em IFN-$\beta$}) as
+well as {\em INF-$\alpha$3} were found to be highly upregulated
+in infected chickens in both lines  (Table~\ref{tab:cytokines}).
+However, expressions of genes encoding their corresponding
+receptors were not different in the resistant line, but
+upregulated in the susceptible line.  This could also reflect the
+ongoing immune responses in the susceptible line.
+
+\subsection*{Functional analysis of differential-expressed genes
+indicates inactive adaptive immune responses in the resistant
+line}
+
+To determine pathways that were perturbed during the infection,
+data were analyzed by GOSeq, which accounted for gene lengths
+bias unique for RNA-Seq data~\cite{young2010method}.
+Significantly perturbed pathways (FDR $< 0.1$) from both lines
+that involved in immune response include TLR signaling pathway,
+cytokine-cytokine receptor interaction, intestinal immune network
+for IgA production, and cell adhesion molecules
+(CAMs)(figure~\ref{line67_kegg}).  Some other pathways important
+in response to viral infection and only significantly enriched in
+the susceptible line include phagosome, apoptosis, RIG-I-like
+receptor signaling pathway, NOD-like receptor signaling pathway,
+and lysosome.  Figure~\ref{kegg_phagosome} is a pictorial example
+of genes in the phagosome pathway.  Although MHC I ({\em BF1})
+was differentially expressed in both lines, other genes involved
+in expressing newly synthesized MHC I were only upregulated in
+the susceptible line suggesting that new MHC I molecules were
+actively produced.  Furthermore, gene ontology analysis of
+biological processes (GO:BP) shows that categories involved in
+both adaptive and innate immune responses were enriched in the
+susceptible line (supplementary material).  On the other hand,
+only categories involved in innate immune responses were enriched
+in the resistant line.  In addition, enrichment of apoptosis
+pathway in the susceptible line indicates that the programmed
+cell death could be induced by CTL response to eliminate the
+viruses.
+
+At this stage of infection, our results suggest that lytic
+infection of MDV stimulates both innate and adaptive immune
+responses, which leads to activation of T cells in the
+susceptible line.  Only activated T cells are believed to be
+infected by MDV, therefore, the lytic phase could facilitate the
+spread of the viruses by enhancing expansion of activated T
+cells.  Due to cell-associate nature of MDV, the viruses transfer
+to T cells via cell-to-cell contact between B cells and T cells
+during antigen presentation or B cells activation by T helper
+cells.  Therefore, it is beneficial for the host to restrain such
+contact.  However, it is not clear how chickens in the resistant
+line control the lytic infection of MDV.  Two mechanisms have
+been speculated to contribute to MD resistance.  First innate
+immune responses could be highly effective and could activate
+strong adaptive immune responses that rapidly control viral
+replication and force the viruses to undergo latent phase.
+Second, innate immune resposnes itself could be highly effective
+in limiting viral replication~\cite{smith2011systems}.
+
+\subsection*{Genes with differential exon usage (DEU) in response
+to MDV infection can be divided into four groups based on their
+patterns of expression}
+
+The immune system is isoform-rich and many genes express
+different isoforms with distinctive functions in response to
+stimuli such as stress, chemicals and infection.  Changes in
+expression of splice forms of immune related genes have been
+reported to be associated with increase susceptibility and poor
+prognosis of diseases~\cite{lynch2004consequences}.  Studying
+differential isoform expression could therefore shed lights into
+inherent differences between lines that confer resistance or
+susceptibility to MD.
+
+In the past decades, microarray technology has been used to study
+gene and isoform expressions in many studies, but its sensitivity
+for detection of structurally similar isoforms is low, and known
+or predicted annotations are required to design
+probes~\cite{kane2000assessment}.  Although RNA-Seq method can
+provide a reliable estimate of an exon expression compared to
+microarray~\cite{pan2008deep} and is not constrained to those
+limitations, studying isoform expressions using RNA-Seq is still
+not straightforward because of the short read lengths.  Reads
+from current Illumina technology are generally not long enough to
+span across all exons in an isoform.  In most cases, only exons
+in close proximity are covered by the same read, which makes it
+difficult to accurately predict a full structure of the isoform.
+In addition, some genes are fused due to overlapping untranslated
+regions (UTRs), which can also result in erroneous predicted
+isoform structures.
+
+Due to those issues, it is not feasible to accurately estimate
+expression of isoforms, especially when gene annotation is
+constructed from {\em de novo}
+assembly~\cite{trapnell2013differential}.  To avoid these issues,
+we chose to study exon expression instead of isoform expression.
+Using MISO with exon-centric method, only reads spanning across a
+few exons are used and only exons involved in a splicing event
+are examined.  The expression of exon inclusion is calculated as
+Percent Spliced In (Psi or $\Psi$), which can be used to infer
+the portion of transcripts that include the exon in each
+sample~\cite{Katz:2010iv}.  In this study, we investigated the
+three most common alternative splicing events in vertebrates,
+which are skipped exons (SE), an alternative $3\prime$ (A3SS) and
+$5\prime$ (A5SS) splice site.  Lists of DEU genes from the
+resistant line that show difference in $\Psi$ greater than 0.20
+when compared to the susceptible line in infected chickens are
+shown in
 Tables~\ref{tab:line67i_diff_line67u_one},~\ref{tab:line67i_diff_line67u_two}
-and~\ref{tab:line67i_diff_line67u_three}.  Genes can be categorized roughly into
-four groups based on the pattern of $\Psi$ across control and infected birds in
-both lines.
-
-Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes with $\Psi$s
-that were up- or down-regulated in infected chickens in line 6 only.  This group
-includes {\em BCL11B} (B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell
-lymphoma associated C2H2-type zinc finger protein encoding gene, which functions
-as a tumour-suppressor in T-cell lymphoma in human.  According to homologous
-alignments on UCSC genome browser, a splice form with the skipped exon is
-similar to mouse {\em BCL11B isoform b}.  The skipped exon was expressed 30\% in
-infected line 6 chickens; whereas it was rarely expressed (4-7\%) in the control
-line 6 and both groups in line 7.  The skipped exon was not found to encode any
-known protein domain, however, it is in the middle of two adjacent C2H2-type
-finger protein domains.  Another gene related to B and T cell lymphomas is {\em
-SIK2} (salt-inducible kinase 2).  This gene has been reported to have a negative
-effect on T cell lymphomas by limiting the transcription of HTLV-1
-virus~\cite{tang2013lkb1}.  Some other genes are important in pre-mRNA splicing
-including {\em TRA2A}, {\em SRSF6} and {\em GEMIN6}.  {\em TRA2A} (transformer-2
-protein homolog alpha) encodes RNA recognition motif (RRM), the skipped exon is
-not found to encode any known protein domain.  {\em SRSF6} (SR splicing factor
-6) encodes a nuclear protein that belongs to the splicing factor protein family.
-{\em GEMIN6} plays a role in the assembly of spliceosomal snRNP in cytoplasm.
-These genes could play a significant role in regulating inclusion of alternative
-exons.  A gene in this group that might be important for innate immune responses
-is {\em RAC3} (Ras-related C3 botulinum toxin substrate 3).  This gene encodes
-small GTPases, belonging to Ras family, that regulates a wide variety of
-cellular events including cell growth, cytoskeletal reorganization, and the
-activation of protein kinases.  The role of small GTPases in immune responses is
-discussed further below.
-
-In group II (Table~\ref{tab:line67i_diff_line67u_two}), $\Psi$ values were
-relatively stable in control and infected chickens within line, but not between
-lines.  Genes that could play an important role in immune responses are {\em
-ITGB2} and {\em HCK}.  {\em HCK} transmits signal from cell surface receptors
-such as {\em FCGR1A, FCGR2A, IL2, IL6, IL18}, and integrins ({\em ITGB1,
-ITGB2}).  {\em ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin of {\em LFA-1}
-and {\em CR3} receptors.  {\em LFA-1} plays an important role in adhesion of
-lymphocytes with other cells.  {\em CR3} binds to a vast array of ligands and
-molecules including complement C3bi, microbial proteins, ICAM-1 and -2, ECM
-proteins, and coagulation proteins~\cite{}.  It plays a significant role in
-neutrophils and monocytes activation including phagocytosis, adhesion and
-migration~\cite{}.  The role of {\em ITGB2} in immune responses is discussed
-further in the next section.  {\em DYNLT1,DYNLL2, SEPT11, PFN2}, and {\em
-ZDHHC7} are also involved in cell rearrangement and cytokinesis.  In particular,
-{\em DYNLT1} and {\em DYNLL2} are dynein proteins that have been demonstrated to
-regulate T cell activation by driving T cell receptor microclusters (TCR-MCs)
-toward the center of immune synapse~\cite{hashimoto2011dynein}.
-
-Group III (Table~\ref{tab:line67i_diff_line67u_three}) includes genes that
-exhibit differential isoform expression only in infected line 7.  A number of
-genes in this group encode proteins that are parts of spliceosome: {\em SRSF3},
-{\em HNRNPDL}, {\em SFSWAP}, {\em THOC1}, {\em RNPC3} and {\em SRSF5}.  Two
-genes involved in cell-cell contact regulated by integrins are also in this
-group.  {\em PPP1R12A} is a myosin phosphatase that regulates the interaction of
-actin and myosin downstream of GTPase Rho proteins.  {\em PODXL} encodes
-PODX-like protein that functions in integrin-dependent manner as both
-pro-adhesive and anti-adhesive molecules.  This protein is involved in
+and~\ref{tab:line67i_diff_line67u_three}.  Genes can be
+categorized roughly into four groups based on the pattern of
+$\Psi$ across control and infected birds in both lines.
+
+Group I (Table~\ref{tab:line67i_diff_line67u_one}) includes genes
+with $\Psi$s that were up- or down-regulated in infected chickens
+in the resistant line only.  This group includes {\em BCL11B}
+(B-cell CLL/lymphoma 11B zinc finger proteins), a B-cell lymphoma
+associated C2H2-type zinc finger protein encoding gene, which
+functions as a tumour-suppressor in T-cell lymphoma in human.
+According to homologous alignments on UCSC genome browser, a
+splice form with the skipped exon is similar to mouse {\em BCL11B
+isoform b}.  The skipped exon was expressed 30\% in infected the
+resistant line chickens; whereas it was rarely expressed (4-7\%)
+in the control the resistant line and both groups in the
+susceptible line.  The skipped exon was not found to encode any
+known protein domain, however, it is in the middle of two
+adjacent C2H2-type finger protein domains.  Another gene related
+to B and T cell lymphomas is {\em SIK2} (salt-inducible kinase
+2).  This gene has been reported to have a negative effect on T
+cell lymphomas by limiting the transcription of HTLV-1
+virus~\cite{tang2013lkb1}.  Some other genes are important in
+pre-mRNA splicing including {\em TRA2A}, {\em SRSF6} and {\em
+GEMIN6}.  {\em TRA2A} (transformer-2 protein homolog alpha)
+encodes RNA recognition motif (RRM), the skipped exon is not
+found to encode any known protein domain.  {\em SRSF6} (SR
+splicing factor 6) encodes a nuclear protein that belongs to the
+splicing factor protein family.  {\em GEMIN6} plays a role in the
+assembly of spliceosomal snRNP in cytoplasm.  These genes could
+play a significant role in regulating inclusion of alternative
+exons.  A gene in this group that might be important for innate
+immune responses is {\em RAC3} (Ras-related C3 botulinum toxin
+substrate 3).  This gene encodes small GTPases, belonging to Ras
+family, that regulates a wide variety of cellular events
+including cell growth, cytoskeletal reorganization, and the
+activation of protein kinases.  The role of small GTPases in
+immune responses is discussed further below.
+
+In group II (Table~\ref{tab:line67i_diff_line67u_two}), $\Psi$
+values were relatively stable in control and infected chickens
+within line, but not between lines.  Genes that could play an
+important role in immune responses are {\em ITGB2} and {\em HCK}.
+{\em HCK} transmits signal from cell surface receptors such as
+{\em FCGR1A, FCGR2A, IL2, IL6, IL18}, and integrins ({\em ITGB1,
+ITGB2}).  {\em ITGB2} (CD18) encodes subunit $\beta_{2}$ integrin
+of {\em LFA-1} and {\em CR3} receptors.  {\em LFA-1} plays an
+important role in adhesion of lymphocytes with other cells.  {\em
+CR3} binds to a vast array of ligands and molecules including
+complement C3bi, microbial proteins, ICAM-1 and -2, ECM proteins,
+and coagulation proteins~\cite{}.  It plays a significant role in
+neutrophils and monocytes activation including phagocytosis,
+adhesion and migration~\cite{}.  The role of {\em ITGB2} in
+immune responses is discussed further in the next section.  {\em
+DYNLT1,DYNLL2, SEPT11, PFN2}, and {\em ZDHHC7} are also involved
+in cell rearrangement and cytokinesis.  In particular, {\em
+DYNLT1} and {\em DYNLL2} are dynein proteins that have been
+demonstrated to regulate T cell activation by driving T cell
+receptor microclusters (TCR-MCs) toward the center of immune
+synapse~\cite{hashimoto2011dynein}.
+
+Group III (Table~\ref{tab:line67i_diff_line67u_three}) includes
+genes that exhibit differential isoform expression only in
+infected the susceptible line.  A number of genes in this group
+encode proteins that are parts of spliceosome: {\em SRSF3}, {\em
+HNRNPDL}, {\em SFSWAP}, {\em THOC1}, {\em RNPC3} and {\em SRSF5}.
+Two genes involved in cell-cell contact regulated by integrins
+are also in this group.  {\em PPP1R12A} is a myosin phosphatase
+that regulates the interaction of actin and myosin downstream of
+GTPase Rho proteins.  {\em PODXL} encodes PODX-like protein that
+functions in integrin-dependent manner as both pro-adhesive and
+anti-adhesive molecules.  This protein is involved in
 cell-to-cell contact, cell trafficking, and cancer
@@
 
-The last group (Group IV, Table~\ref{tab:line67i_diff_line67u_three}) only has
-one gene, {\em GOSR1}.  The $\Psi$ value of this gene was less than 0.20 cutoff
-in control and infected chickens in line 6, but it is significantly different
-between infected chickens in line 6 and 7.  Also, there is a significant
-difference between control and infected chickens in line 7.  This gene encodes a
-trafficking membrane protein important for transporting proteins from {\em cis-}
-to {\em trans-}golgi network.
-
-\subsection*{Roles of {\em LFA-1} and actin cytoskeleton in T cells
-activation}
-
-By grouping genes based on patterns of $\Psi$s, we found that many genes in
-group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1, SEPT11}, and {\em ZDHHC7}, are
-involved in cytokinesis or cell synapse, which are important for T cell
-activation.  As described above, {\em ITGB2} encodes $\beta$-subunit of
-integrins including LFA-1, which exclusively expressed in lymphocytes and plays
-a major role in lymphoproliferation, antigen presentation, T cells activation,
-and cytotoxicity.  Integrins are a special kind of receptors that transmit
-signals bidirectionally across the cell membrane.  They are heterodimeric
-composed of an $\alpha$ (large) and a $\beta$ (small)
-subunits~\cite{wang2010immunopathologies}.  $\beta_{2}$ (CD18) subunit encoded
-by {\em ITGB2} is expressed on lymphocytes and antigen presenting cells (APCs)
-as a component of {\em LFA-1} and {\em CR3} receptors.  LFA-1 binds to its
-ligand ICAM-1 to help form a synapse that bring APC and T cell together to
-initiate antigen presentation leading to T cell
+The last group (Group IV,
+Table~\ref{tab:line67i_diff_line67u_three}) only has one gene,
+{\em GOSR1}.  The $\Psi$ value of this gene was less than 0.20
+cutoff in control and infected chickens in the resistant line,
+but it is significantly different between infected chickens in
+the resistant and susceptible lines.  Also, there is a significant
+difference between control and infected chickens in the
+susceptible line.  This gene encodes a trafficking membrane
+protein important for transporting proteins from {\em cis-} to
+{\em trans-}golgi network.
+
+\subsection*{Roles of {\em LFA-1} and actin cytoskeleton in T
+cells activation}
+
+By grouping genes based on patterns of $\Psi$s, we found that
+many genes in group II: {\em ITGB2, PFN2, DYNLL2, DYNLT1,
+SEPT11}, and {\em ZDHHC7}, are involved in cytokinesis or cell
+synapse, which are important for T cell activation.  As described
+above, {\em ITGB2} encodes $\beta$-subunit of integrins including
+LFA-1, which exclusively expressed in lymphocytes and plays a
+major role in lymphoproliferation, antigen presentation, T cells
+activation, and cytotoxicity.  Integrins are a special kind of
+receptors that transmit signals bidirectionally across the cell
+membrane.  They are heterodimeric composed of an $\alpha$ (large)
+and a $\beta$ (small) subunits~\cite{wang2010immunopathologies}.
+$\beta_{2}$ (CD18) subunit encoded by {\em ITGB2} is expressed on
+lymphocytes and antigen presenting cells (APCs) as a component of
+{\em LFA-1} and {\em CR3} receptors.  LFA-1 binds to its ligand
+ICAM-1 to help form a synapse that bring APC and T cell together
+to initiate antigen presentation leading to T cell
 activation~\cite{dustin2000immunological}.
 
-Absence of LFA-1 leads to impaired functions of lymphocytes in proliferation and
-tumor rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.  Mutations in
-{\em ITGB2} gene has been associated with type 1 leukocyte adhesion deficiency
-(LAD-1), an autosomal-recessive inherited disease found in few families.  The
-disease is characterized by impair of lymphocytes in adherent-dependent
-functions, lack of accumulation to the site of infection and recurrent bacterial
-and fungal infection~\cite{springer1987lymphocyte}.  In addition, the response
-of lymphocytes to mitogens is decreased in patients with
-LAD~\cite{springer1987lymphocyte}.  The decrease in responsiveness to mitogens
-have been shown to correlate with resistance to MD by Lee and
-Bacon~\cite{lee1983ontogeny}, who illustrated that resistant birds (line 6 and
-N) were less responsive to phytohemagglutinin (PHA) than susceptible birds (line
-7 and P).
-
-Actin cytoskeleton is very important in T cell activation because it enhances
-activity of LFA-1 by increasing its avidity and recruiting signaling molecules
-necessary for downstream signaling~\cite{dustin2000immunological,
-van2000avidity}.  Cytoskeleton proteins binding to cytoplasmic domain of LFA-1
-are thought to play an important role in driving LFA-1 to aggregation on the
-cell surface, resulting in increased avidity.  Aggregation of LFA-1 has been
-demonstrated to be essential for lymphocytes to bind to the
-ligand~\cite{van1994extracellular}.  Interestingly, {\em RAC3, PFN2,} and {\em
-PPP1R12A}, which are involved in actin cytoskeleton pathway
-(Figure~\ref{kegg_actin}), also expressed different ratios of alternative splice
-forms between lines 6 and 7.  Some of these genes are also co-present in three
-other pathways that involved in immune responses (Table~\ref{tab:integrin}).
-It could be speculated that pre-mRNA splicing of these genes is co-regulated
-by splicing regulators or some genetic factors.
-
-\subsection*{Prediction of functional domains of splice forms of genes in the
-actin cytoskeleton pathway}
-
-To predict the function of the alternative splice forms of genes in actin
-cytoskeleton pathway, transcript sequences were translated to protein sequences
-by ESTscan~\cite{iseli1999estscan}.  Protein sequences were then searched for
-annotated protein domains using the standalone version of InterPro
-Scan~\cite{quevillon2005interproscan}.  Besides {\em ITGB2}, other genes have
-alternative exons located in coding regions and could potentially affect
-functional protein domains in some ways.  The exon with alternative 3$\prime$
-splice site of {\em RAC3} encode part of a protein domain identified as small
-GTPase of Ras subfamily (ProSiteProfiles:PS5142 and SMART:SM00173).  Rac3 is
-highly homologous to Rac1 and has been reported to possess the ability of
-promoting membrane ruffling, transformation, activation of c-Jun transcriptional
-activity and a co-activator of NF$\kappa$B~\cite{werbajh2000rac}.  Activated Rac
-also regulates production of superoxide in neutrophils and macrophages.
-
-The alternative exon of {\em PFN2} seems to disrupt the coding sequence that
-encode profilin domain (Pfam:PF00235).  The profilin domain is essential for
-almost all organisms and its functions include regulating actin polymerization,
-controlling complex network of molecular interaction and transmitting signals
-from small-GTPase pathway.  It also binds to Rac effector molecules and a number
-of other ligands~\cite{witke2004role}.  The skipped exon of {\em PPP1R12A}
-encodes part of a protein domain annotated as protein phosphatase 1
-(PIRSF:PIRSF038141).  The functional effects of alternative splicing on these
-functional domains have to be further investigated by simulation or an
-experiment.
-
-Even though the exact mechanism is not known, incompetency of T cells in
-response to stimuli appears to benefit resistant birds because MD virus could
-not induce T cells to proliferate and cause them to undergo neoplastic
-transformation.  It has also been suggested that the mechanism that controls
-both lymphocyte proliferation induced by MD virus and lymphocyte proliferation
-induced by immune response is the same~\cite{pazderka1975histocompatibility}.
-Therefore, it may be useful to consider a link between deficiency of lymphocytes
-in line 6 to the alternative splice form of {\em ITGB2} that is only expressed
-in line 6.  Although the exon included in the alternative splice form is
-non-coding, it could serve important functions in translation or
-posttranscriptional regulation.
-
-% Based on our gene model, the skipped exon of \textit{ITGB2} is located in the
-% 5$\prime$ UTR; whereas, the skipped exon of \textit{HCK}, encodes protein
-% tyrosine kinase (Pfam:PF07714).
-
-\subsection*{Prediction of {\em cis}-regulatory elements in alternative
-splicing exons of genes in group II}
-
-Although alternative splicing of genes in all groups could be regulated by {\em
-cis-} and {\em trans-}splicing factors, genes in group II are more likely to be
-regulated by {\em cis-}regulatory factors.  The ratios of isoform expression in
-this group were relatively stable within line, but were significantly different
-between lines.  Investigation of nucleotide differences within exons of lines 6
-and 7 could reveal a possible role of SNPs in regulating alternative splicing in
-this group.  We obtained a sequence of an alternative exon from line 6 and used
-Human Splicing Finder (HSF) to determine whether SNPs from line 7 could alter
-predicted ESEs or ESSs.  Results from some genes involved in cytokinesis are
-discussed in this section.  Exonic SNPs from line 6 and 7 are listed in
-Table~\ref{tab:deu_snps}.
-
-For {\em ITGB2}, a SNP (T) at position 26 of the cDNA from line 6, which
-corresponds to position 7,183,696 on chromosome 7 is located in a predicted
-binding site for SC35, which is an exon enhancer.  Although the exon is not
-expressed in line 7, we found that there is no polymorphisms between lines 6 and
-7 according to SNPs data from the genome resequencing.  Therefore, this SNP may
-not be accounted for exclusion of the exon in line 7 (Figure~\ref{itgb2}).  Exon
-sequences of {\em PFN2} from line 6 and 7 differ at position 2,072 on cDNA or
-23,221,934 on chromosome 9.  A small insertion of two AA nucleotides is
-predicted to create a new binding site for Tra2-$\beta$ splicing regulator,
-which serves as a stabilizer of an enhancer complex~\cite{lopez1998alternative}.
-From exon expression data, $\Psi$ of an exon with alternative splicing increases
-from 0.20-0.30 to about 0.50 in line 7.  Tra2 could possibly increase inclusion
-of the exon with alternative 3$\prime$ splice site via ESE-dependent 3$\prime$
-splice site activation (Figure~\ref{pfn2}).  Even though the linear distance
-between Tra2-$\beta$ binding site and the alternative 3$\prime$ splice site is
-greater than 1kb, Tra2 could possibly get close to the 3$\prime$ splice site in
-the secondary structure of the mRNA.
-
-Replacement of A with G nucleotide in the skipped exon of {\em DYNLL2} from line
-6 is predicted to slightly alter the binding site of several ESEs as well as to
-create a new binding site for 9G8.  This exon is upregulated in line 7 compared
-to line 6, therefore, the present of the new binding site for 9G8 exon enhancer
-helps support the expression results.  In addition, A nucleotide is this
-position matches the reference nucleotide, therefore, we could expect this exon
-to be expressed in other datasets.  According to EST tags on the UCSC genome
-browser, the exon has been found and sequenced from chicken eyes (15d
-post-hatched, EST sequence:DR424100).  For {\em DYNLT1}, the skipped exon was
-absent in line 7, therefore, we could not determine the SNPs from RNA-Seq data.
-However, data from DNA resequencing showed that there was a polymorphism at
-position 51,357,865 on chromosome 3.  Replacing G with C from line 7 is
-predicted to slightly alter predicted binding sites for 9G8 as well as SRp55
-exon enhancer.  However, whether the alteration down regulates the exon in line
-7 is unclear.
-
-There are too many SNPs in the exon of {\em SEPT11} and {\em ZDHHC7}, making it
-unfeasible to predict which SNP regulates the exon expression.  Therefore, we do
-not discuss these two genes in this section.  However, results from HSF analysis
-of these two genes and other genes in this group are provided in the
-supplementary materials.  Experimental validation of exonic SNPs provided by
-this study could shed some lights on underlying polymorphisms that contribute to
+Absence of LFA-1 leads to impaired functions of lymphocytes in
+proliferation and tumor
+rejection~\cite{scharffetter1998spontaneous,schmits1996lfa}.
+Mutations in {\em ITGB2} gene has been associated with type 1
+leukocyte adhesion deficiency (LAD-1), an autosomal-recessive
+inherited disease found in few families.  The disease is
+characterized by impair of lymphocytes in adherent-dependent
+functions, lack of accumulation to the site of infection and
+recurrent bacterial and fungal
+infection~\cite{springer1987lymphocyte}.  In addition, the
+response of lymphocytes to mitogens is decreased in patients with
+LAD~\cite{springer1987lymphocyte}.  The decrease in
+responsiveness to mitogens have been shown to correlate with
+resistance to MD by Lee and Bacon~\cite{lee1983ontogeny}, who
+illustrated that resistant birds (the resistant line and N) were
+less responsive to phytohemagglutinin (PHA) than susceptible
+birds (line 7 and P).
+
+Actin cytoskeleton is very important in T cell activation because
+it enhances activity of LFA-1 by increasing its avidity and
+recruiting signaling molecules necessary for downstream
+signaling~\cite{dustin2000immunological, van2000avidity}.
+Cytoskeleton proteins binding to cytoplasmic domain of LFA-1 are
+thought to play an important role in driving LFA-1 to aggregation
+on the cell surface, resulting in increased avidity.  Aggregation
+of LFA-1 has been demonstrated to be essential for lymphocytes to
+bind to the ligand~\cite{van1994extracellular}.  Interestingly,
+{\em RAC3, PFN2,} and {\em PPP1R12A}, which are involved in actin
+cytoskeleton pathway (Figure~\ref{kegg_actin}), also expressed
+different ratios of alternative splice forms between lines .
+Some of these genes are also co-present in three other pathways
+that involved in immune responses (Table~\ref{tab:integrin}).  It
+could be speculated that pre-mRNA splicing of these genes is
+co-regulated by splicing regulators or some genetic factors.
+
+\subsection*{Prediction of functional domains of splice forms of
+genes in the actin cytoskeleton pathway}
+
+To predict the function of the alternative splice forms of genes
+in actin cytoskeleton pathway, transcript sequences were
+translated to protein sequences by
+ESTscan~\cite{iseli1999estscan}.  Protein sequences were then
+searched for annotated protein domains using the standalone
+version of InterPro Scan~\cite{quevillon2005interproscan}.
+Besides {\em ITGB2}, other genes have alternative exons located
+in coding regions and could potentially affect functional protein
+domains in some ways.  The exon with alternative 3$\prime$ splice
+site of {\em RAC3} encode part of a protein domain identified as
+small GTPase of Ras subfamily (ProSiteProfiles:PS5142 and
+SMART:SM00173).  Rac3 is highly homologous to Rac1 and has been
+reported to possess the ability of promoting membrane ruffling,
+transformation, activation of c-Jun transcriptional activity and
+a co-activator of NF$\kappa$B~\cite{werbajh2000rac}.  Activated
+Rac also regulates production of superoxide in neutrophils and
+macrophages.
+
+The alternative exon of {\em PFN2} seems to disrupt the coding
+sequence that encode profilin domain (Pfam:PF00235).  The
+profilin domain is essential for almost all organisms and its
+functions include regulating actin polymerization, controlling
+complex network of molecular interaction and transmitting signals
+from small-GTPase pathway.  It also binds to Rac effector
+molecules and a number of other ligands~\cite{witke2004role}.
+The skipped exon of {\em PPP1R12A} encodes part of a protein
+domain annotated as protein phosphatase 1 (PIRSF:PIRSF038141).
+The functional effects of alternative splicing on these
+functional domains have to be further investigated by simulation
+or an experiment.
+
+Even though the exact mechanism is not known, incompetency of T
+cells in response to stimuli appears to benefit resistant birds
+because MDV could not induce T cells to proliferate and cause
+them to undergo neoplastic transformation.  It has also been
+suggested that the mechanism that controls both lymphocyte
+proliferation induced by MDV and lymphocyte proliferation induced
+by immune response is the
+same~\cite{pazderka1975histocompatibility}.  Therefore, it may be
+useful to consider a link between deficiency of lymphocytes in
+the resistant line to the alternative splice form of {\em ITGB2}
+that is only expressed in the resistant line.  Although the exon
+included in the alternative splice form is non-coding, it could
+serve important functions in translation or posttranscriptional
+regulation.
+
+% Based on our gene model, the skipped exon of \textit{ITGB2} is
+% located in the 5$\prime$ UTR; whereas, the skipped exon of
+% \textit{HCK}, encodes protein tyrosine kinase (Pfam:PF07714).
+
+\subsection*{Prediction of {\em cis}-regulatory elements in
+alternative splicing exons of genes in group II}
+
+Although alternative splicing of genes in all groups could be
+regulated by {\em cis-} and {\em trans-}splicing factors, genes
+in group II are more likely to be regulated by {\em
+cis-}regulatory factors.  The ratios of isoform expression in
+this group were relatively stable within line, but were
+significantly different between lines.  Investigation of
+nucleotide differences within exons of both lines could reveal a
+possible role of SNPs in regulating alternative splicing in this
+group.  We obtained a sequence of an alternative exon from the
+resistant line and used Human Splicing Finder (HSF) to determine
+whether SNPs from the susceptible line could alter predicted ESEs
+or ESSs.  Results from some genes involved in cytokinesis are
+discussed in this section.  Exonic SNPs from the resistant and
+susceptible lines are listed in Table~\ref{tab:deu_snps}.
+
+For {\em ITGB2}, a SNP (T) at position 26 of the cDNA from the
+resistant line, which corresponds to position 7,183,696 on
+chromosome 7 is located in a predicted binding site for SC35,
+which is an exon enhancer.  Although the exon is not expressed in
+the susceptible line, we found that there is no polymorphisms
+between lines according to SNPs data from the genome
+resequencing.  Therefore, this SNP may not be accounted for
+exclusion of the exon in the susceptible line
+(Figure~\ref{itgb2}).  Exon sequences of {\em PFN2} from the
+resistant line and 7 differ at position 23,221,934 on chromosome
+9.  A small insertion of two AA nucleotides is predicted to
+create a new binding site for Tra2-$\beta$ splicing regulator,
+which serves as a stabilizer of an enhancer
+complex~\cite{lopez1998alternative}.  From exon expression data,
+$\Psi$ of an exon with alternative splicing increases from
+0.20-0.30 to about 0.50 in the susceptible line.  Tra2 could
+possibly increase inclusion of the exon with alternative
+3$\prime$ splice site via ESE-dependent 3$\prime$ splice site
+activation (Figure~\ref{pfn2}).  Even though the linear distance
+between Tra2-$\beta$ binding site and the alternative 3$\prime$
+splice site is greater than 1kb, Tra2 could possibly get close to
+the 3$\prime$ splice site in the secondary structure of the mRNA.
+
+Replacement of A with G nucleotide in the skipped exon of {\em
+DYNLL2} from line 6 is predicted to slightly alter the binding
+site of several ESEs as well as to create a new binding site for
+9G8.  This exon is upregulated in the susceptible line compared
+to the resistant line, therefore, the present of the new binding
+site for 9G8 exon enhancer helps support the expression results.
+In addition, A nucleotide is this position matches the reference
+nucleotide, therefore, we could expect this exon to be expressed
+in other datasets.  According to EST tags on the UCSC genome
+browser, the exon has been found and sequenced from chicken eyes
+(15d post-hatched, EST sequence:DR424100).  For {\em DYNLT1}, the
+skipped exon was absent in the susceptible line, therefore, we
+could not determine the SNPs from RNA-Seq data.  However, data
+from DNA resequencing showed that there was a polymorphism at
+position 51,357,865 on chromosome 3.  Replacing G with C from the
+susceptible line is predicted to slightly alter predicted binding
+sites for 9G8 as well as SRp55 exon enhancer.  However, whether
+the alteration down regulates the exon in line 7 is unclear.
+
+There are too many SNPs in the exon of {\em SEPT11} and {\em
+ZDHHC7}, making it unfeasible to predict which SNP regulates the
+exon expression.  Therefore, we do not discuss these two genes in
+this section.  However, results from HSF analysis of these two
+genes and other genes in this group are provided in the
+supplementary materials.  Experimental validation of exonic SNPs
+provided by this study could shed some lights on underlying
+polymorphisms that contribute to
 
@@
 
-Custom gene models built from combination of gene models from {\em de novo}
-assembly, reference-based assembly, and ENSEMBL had allowed us to identify genes
-and isoforms that might play an important role in resistance to MD.  Results
-from gene expression analysis indicated that adaptive immune responses were
-active during lytic infection in line 7, but not in line 6.  Because only
-activated T cells are infected by MDV, we speculated that elicitation of
-adaptive immune responses could help spread the viruses by recruiting and
-activating more T cells and.  In contrast, the delay of adaptive immune
-responses could benefit the host by limiting infection of activated T cells.
-
-To elucidate the molecular mechanism of MD resistance, we investigated
-differential isoforms expression between lines 6 and 7 and identified a number
-of genes that could be responsible for difference in immune responses. Notably,
-several genes are involved in actin cytoskeleton and cytokinesis, which are
-important for the functions of lymphocytes and immune cells but have not been of
-great interest in the filed of MD research.  Even though we mainly discussed the
-possible role of {\em ITGB2} in MD resistance, other genes cannot be precluded
-and should be a candidate for further investigation and experimental studies.
-Moreover, a full mechanism of MD resistance is highly complex and more data from
-different stages of infection as well as a greater sequencing depth will be
-required to identify all genes and isoforms involved.  To enable the study of
-unannotated gene and isoform expression, our approach of constructing gene
-models from RNA-Seq should be iteratively used to extend the ENSEMBL gene models
-to construct more complete gene models.
+Custom gene models built from combination of gene models from
+{\em de novo} assembly, reference-based assembly, and ENSEMBL had
+allowed us to identify genes and isoforms that might play an
+important role in resistance to MD.  Results from gene expression
+analysis indicated that adaptive immune responses were active
+during lytic infection in the susceptible line, but not in the
+resistant line.  Because only activated T cells are infected by
+MDV, we speculated that elicitation of adaptive immune responses
+could help spread the viruses by recruiting and activating more T
+cells and.  In contrast, the delay of adaptive immune responses
+could benefit the host by limiting infection of activated T
+cells.
+
+To elucidate the molecular mechanism of MD resistance, we
+investigated differential isoforms expression between lines  and
+identified a number of genes that could be responsible for
+difference in immune responses. Notably, several genes are
+involved in actin cytoskeleton and cytokinesis, which are
+important for the functions of lymphocytes and immune cells but
+have not been of great interest in the filed of MD research.
+Even though we mainly discussed the possible role of {\em ITGB2}
+in MD resistance, other genes cannot be precluded and should be a
+candidate for further investigation and experimental studies.
+Moreover, a full mechanism of MD resistance is highly complex and
+more data from different stages of infection as well as a greater
+sequencing depth will be required to identify all genes and
+isoforms involved.  To enable the study of unannotated gene and
+isoform expression, our approach of constructing gene models from
+RNA-Seq should be iteratively used to extend the ENSEMBL gene
+models to construct more complete gene models.
 
@@
 
-\subsection{Sequences and quality trimming}
-
-mRNAs were extracted from spleens of control and infected chickens lines 6 and 7
-(4 d.p.i).  Sequence libraries were prepared by standard Illumina unstranded
-single- and paired-end protocols.  Library size of the paired-end datasets is
-approximately 175 bp.  Read lengths are 75 bp in both single- and paired-end
-libraries.  Reads were quality trimmed by Condetri 2.1~\cite{smeds2011condetri}
-with quality score cutoff of 30.  The first 10 bases were removed due to
-non-uniform distribution of nucleotides.
+\subsection*{Sequences and quality trimming}
+
+mRNAs were extracted from spleens of control and infected
+chickens lines 6 and 7 (4 d.p.i).  Sequence libraries were
+prepared by standard Illumina unstranded single- and paired-end
+protocols.  Library size of the paired-end datasets is
+approximately 175 bp.  Read lengths are 75 bp in both single- and
+paired-end libraries.  Reads were quality trimmed by Condetri
+2.1~\cite{smeds2011condetri} with quality score cutoff of 30.
+The first 10 bases were removed due to non-uniform distribution
+of nucleotides.
 
@@
 
-Due to lack of complete gene models for chickens, we employed two methods to
-construct gene models from RNA-Seq reads.  First, short reads were assembled
-using Velvet/1.2.03~\cite{Zerbino:2008vu} and Oases/0.2.06~\cite{Schulz:2012je}
-to obtain long transcripts.  Assembly was done with hash lengths range from 21
-to 31 for both local and global assembly (described in Gimme paper).  Poly-A
+Due to lack of complete gene models for chickens, we employed two
+methods to construct gene models from RNA-Seq reads.  First,
+short reads were assembled using
+Velvet/1.2.03~\cite{Zerbino:2008vu} and
+Oases/0.2.06~\cite{Schulz:2012je} to obtain long transcripts.
+Assembly was done with hash lengths range from 21 to 31 for both
+local and global assembly (described in Gimme paper).  Poly-A
 tails were trimmed and low complexity transcripts were removed by
-Seqclean~\cite{seqclean}.  All transcripts were then aligned to chicken
-reference genome (galGal4, with unplaced and random chromosomes removed) with
-BLAT~\cite{Kent:2002tv}.  Second, reads were aligned to reference genome using
-Tophat/2.0.9~\cite{Trapnell:2009dp} and ENSEMBLE gene models release 73 was used
-to guide reference-based assembly by Cufflinks2~\cite{Trapnell:2010kd}.
-Alignments from BLAT and models from Cufflinks were then combined to construct
-gene models by Gimme (manuscript in preparation).
-
-\subsection*{Differential gene expression and gene ontology}
+Seqclean~\cite{seqclean}.  All transcripts were then aligned to
+chicken reference genome (galGal4, with unplaced and random
+chromosomes removed) with BLAT~\cite{Kent:2002tv}.  Second, reads
+were aligned to reference genome using
+Tophat/2.0.9~\cite{Trapnell:2009dp} and ENSEMBLE gene models
+release 73 was used to guide reference-based assembly by
+Cufflinks2~\cite{Trapnell:2010kd}.  Alignments from BLAT and
+models from Cufflinks were then combined to construct gene models
+by Gimme (manuscript in preparation).
+
+\subsection*{Differential gene expression analysis and gene
+ontology}
 
 To identify DE genes, reads were mapped to transcripts by RSEM
-v.1.10.1~\cite{li2011rsem}, which is also used to estimate genes expression and
-identify DE genes.  Data from single- and paired-end datasets from the same line
-were treated as biological replicates.  To identify enriched pathways and
-ontology terms, a list of DE genes was analysed by GOSeq v.1.10.0 based on
-chicken KEGG annotations.  P-values were corrected by Benjamini-Hochberg
-multiple testing correction.  Genes, patways and GO terms with corrected P-value
-$<0.1$ were considered significant.  Pathview~\cite{luo2013pathview} was used to
-create a KEGG pathway diagram with colors representing relative level of gene
+v.1.2.7~\cite{li2011rsem}, which is also used to estimate genes
+expression and identify DE genes.  Data from single- and
+paired-end datasets from the same line were treated as biological
+replicates.  To identify enriched pathways and ontology terms, a
+list of DE genes was analysed by GOSeq v.1.10.0 based on chicken
+KEGG annotations.  P-values were corrected by Benjamini-Hochberg
+multiple testing correction.  Genes, patways and GO terms with
+corrected P-value $<0.1$ were considered significant.
+Pathview~\cite{luo2013pathview} was used to create a KEGG pathway
+diagram with colors representing relative level of gene
 expressions.
 
-\subsection*{Differential splicing event}
-
-Gene models were converted to alternative splicing models using a Python script.
-In order to increase sensitivity, read counts from single- and paired-end
-samples were combined and treated as single-end reads for splicing event
-analysis with MISO/0.4.9~\cite{Katz:2010iv}.  Splicing events with Bayes factor
-$>10$ and $\Delta\Psi>0.20$ were considered significant.  Read coverages and
-$\Psi$ distributions were plotted using Sashimi plot~\cite{Katz:2013vx}.
-
-\subsection*{Variant calling and {\em in silico} splicing analysis}
+\subsection*{Differential exon usage analysis}
+
+Gene models were converted to alternative splicing models using a
+Python script.  In order to increase sensitivity, read counts
+from single- and paired-end samples were combined and treated as
+single-end reads for splicing event analysis with
+MISO/0.4.9~\cite{Katz:2010iv}.  Splicing events with Bayes factor
+$>10$ and $\Delta\Psi>0.20$ were considered significant.  Read
+coverages and $\Psi$ distributions were plotted using Sashimi
+plot~\cite{Katz:2013vx}.
+
+\subsection*{Variant calling and {\em in silico} splicing
+analysis}
 
 Variants were called using mpileup command from
-SAMTools/0.1.18~\cite{li2009sequence} and BCFTools~\cite{bcftools}.  Only
-variants with quality score $\ge20$ were used for mutation analyses.  Exon
-enhancers and suppressors were predicted using Human Splicing Finder web
-portal~\cite{desmet2009human}.  Human default parameter settings were used in
-all analyses.  The following regulatory sequences were used to determined the
-effect of variants: HSF integrated matrices for serine/arginine-rich proteins
-(SRp40, SC35, SF2/ASF, SF2/ASF, IgM/BRCA1, and SRp55), exonic splicing enhancer
-(ESE), RESCUE-ESE hexamer (RESCUE-ESE), putative 8-mer ESEs (PESEs) and putative
-8-mers exonic splicing silencers (PESSs), exon-identity and intron-identiry
-elements (EIEs and IIEs), hetronuclear ribonucleoprotein-binding motifs, and Fas
-exonic splicing silencers.
+SAMTools/0.1.18~\cite{li2009sequence} and
+BCFTools~\cite{bcftools}.  Only variants with quality score
+$\ge20$ were used for mutation analyses.  Exon enhancers and
+suppressors were predicted using Human Splicing Finder web
+portal~\cite{desmet2009human}.  Human default parameter settings
+were used in all analyses.
 
 \subsection*{Protein domains search}
+
+Transcripts were translated to protein sequences using ESTScan
+3.0.3~\cite{iseli1999estscan} with chicken HMM matrices built
+from chicken cDNAs and refSeq sequences. Protein sequences were
+searched against InterPro database using
+InterProScan/5.44.0~\cite{quevillon2005interproscan}.
 
@@
         {\bf ITGB2 exon expression.}
-        A single nuclotide substition (T$>$C) at position 7,183,696 is predicted to disrupt a putative ESE
-        motif which binds SC35 on the skipped exon of IRGB2 gene shown in this figure.
-        The prediction correlate with exclusion of the exon in line 7 ($\Psi \leq 0.02$).
-        A new binding site for a silencer is also predicted and could also promote the exclusion of the exon. 
-    }
+        A single nuclotide substition (T$>$C) at position
+        7,183,696 is predicted to disrupt a putative ESE motif
+        which binds SC35 on the skipped exon of IRGB2 gene shown
+        in this figure.  The prediction correlate with exclusion
+        of the exon in the susceptible line ($\Psi \leq 0.02$).
+    A new binding site for a silencer is also predicted and could
+also promote the exclusion of the exon.  }
     \label{itgb2}
@@
         {\bf PFN2 exon expression.}
-        A small insertion of AA nucleotides at position 23,221,934 is predicted to create a putative ESE motif
-        which binds to Tra2 protein.
-        Tra2 could promote inclusion of exon with alternative 3$\prime$ splice site via ESE-dependent
-        3$\prime$ splice site activation resulting in higher expression of the exon in line 7.
-        Even though the predicted ESE motif is located at a substantial distance from the alternative 3$\prime$ splice site,
-        Tra2 could still regulate the splicing if a secondary structure of the mRNA moves the ESE motif closer to the
-        alternative 3$\prime$ splice site.
-    }
+        A small insertion of AA nucleotides at position
+        23,221,934 is predicted to create a putative ESE motif
+        which binds to Tra2 protein.  Tra2 could promote
+        inclusion of exon with alternative 3$\prime$ splice site
+        via ESE-dependent 3$\prime$ splice site activation
+        resulting in higher expression of the exon in the
+        susceptible line.  Even though the predicted ESE motif is
+        located at a substantial distance from the alternative
+        3$\prime$ splice site, Tra2 could still regulate the
+    splicing if a secondary structure of the mRNA moves the ESE
+motif closer to the alternative 3$\prime$ splice site.  }
     \label{pfn2}
@@
 \section*{Tables}
+
+\begin{table}[!ht]
+\caption{
+\bf{Genes regulated in opposite directions}}
+\begin{tabular}{cccc}
+        \hline
+        Gene & Description & $log_{2}$FC & \\
+         & & Resistant & Susceptible\\
+        \hline
+        LL & Lung lectin & -3.36 & 8.70 \\
+        GIF & Gastric intrinsic factor & -2.15 & 3.11 \\
+        SFTPA1 & Surfactant protein A1 & -4.84 & 3.73 \\
+        SCAF8 & SR-related CTD-associated factor 8 & -8.70 & 8.24 \\
+        LIMS1 & LIM and senescent cell antigen-like domains 1 & -2.26 & 1.33 \\
+        PPARG & Peroxisome proliferator-activated receptor gamma & -6.99 & 2.06 \\
+        C14ORF1 & Chromosome 14 open reading frame 1 & -3.11 & 2.71 \\
+        Unknown & Unknown & -3.08 & 1.62 \\
+        \hline
+        ATP8A2 & ATPase, aminophospholipid transporter, class I, type 8A, member 2 & 7.66 & -7.80 \\
+        S1PR1 & Sphingosine-1-phosphate receptor 1 & 1.31 & -6.95 \\
+        MED9 & Mediator complex subunit 9 & 8.22 & -3.15 \\
+        DNAJA2 & DnaJ (Hsp40) homolog, subfamily A, member 1 & 2.34 & -6.02 \\
+        RAD17 & RAD17 homolog (S. pombe) & 1.45 & -1.80 \\
+        PSMG3 & Proteosome assembly chaperone 3 & 1.52 & -1.26 \\
+        PNISR & PNN-interacting serine/argining-rich protein & 5.58 & -1.81 \\
+        THOC7 & THO complex 7 homolog (Drosophila) & 7.00 & -6.90 \\
+        YTHDC2 & YTH domain containing 2 & 1.00 & -1.74 \\
+        RPL39 & Ribosomal protein L39 & 3.31 & -1.59 \\
+        CD7 & CD7 molecule & 2.88 & -1.38 \\
+        NDUFB3 & NADH dehydrogenase (uqiquinone) 1 beta subcomplex 3 & 1.03 & -1.34 \\
+        LOC100858785 & Unknown & 1.26 & -1.75 \\
+        Unknown & Unknown & 6.30 & -5.22 \\
+        \hline
+    \end{tabular}
+    \begin{flushleft}
+        (-) down-regulated, (+) up-regulated
+    \end{flushleft}
+    \label{tab:opposite}
+\end{table}
+
+\begin{table}[!ht]
+\caption{
+\bf{Cytokine-related gene expression in response to MDV infection}}
+    \begin{tabular}{cccccc}
+        \hline
+        & & $log_{2}$FC & \\
+        Symbol & Description & Resistant & Susceptible \\
+        \hline
+        IL2RG & Interleukin 2 receptor, $\gamma$ & -- & 0.55 \\
+        IL6 & Interleukin 6 (interferon, $\beta$ 2) & -- & 5.15 \\
+        IL6ST & Interleukin 6 signal transducer (gp130, oncostatin M receptor) & -- & 1.36 \\
+        IL8L1 & Interleukin 8-like 1 & 1.90 & -- \\
+        IL15 & Interleukin 15 & -- & 1.06 \\
+        IL18 & Interleukin 18 (interferon-$\gamma$ inducing factor) & 1.92 & 4.06 \\
+        IL18R1 & Interleukin 18 receptor 1 & 1.94 & 1.64 \\
+        IFNG & Interferon-$\gamma$ & 5.14 & 4.90 \\
+        IFNB & Interferon-$\beta$ & 4.83 & 5.64 \\
+        IFNA3 & Interferon-$\alpha$ 3 & 4.09 & 5.48 \\
+        IFNGR1 & Interferon-$\gamma$ receptor 1 & -- & 2.05 \\
+        IFNGR2 & Interferon-$\gamma$ receptor 2& -- & 0.50 \\
+        IFNAR1 & Interferon-$\alpha$,$\beta$ receptor 1 & -- & 1.46 \\
+        IFNAR2 & Interferon-$\alpha$,$\beta$ receptor 2 & -- & 0.58 \\
+        \hline
+    \end{tabular}
+    \begin{flushleft}
+    \end{flushleft}
+    \label{tab:cytokines}
+\end{table}
 
@@
 \caption{
-\bf{Genes regulated in opposite directions}}
-    \begin{tabular}{cccc}
-        \hline
-        Gene & Description & $log_{2}$FC & \\
-         & & Line 6 & Line 7 \\
-        \hline
-        LL & Lung lectin & -3.36 & 8.70 \\
-        GIF & Gastric intrinsic factor & -2.15 & 3.11 \\
-        SFTPA1 & Surfactant protein A1 & -4.84 & 3.73 \\
-        SCAF8 & SR-related CTD-associated factor 8 & -8.70 & 8.24 \\
-        LIMS1 & LIM and senescent cell antigen-like domains 1 & -2.26 & 1.33 \\
-        PPARG & Peroxisome proliferator-activated receptor gamma & -6.99 & 2.06 \\
-        C14ORF1 & Chromosome 14 open reading frame 1 & -3.11 & 2.71 \\
-        Unknown & Unknown & -3.08 & 1.62 \\
-        \hline
-        ATP8A2 & ATPase, aminophospholipid transporter, class I, type 8A, member 2 & 7.66 & -7.80 \\
-        S1PR1 & Sphingosine-1-phosphate receptor 1 & 1.31 & -6.95 \\
-        MED9 & Mediator complex subunit 9 & 8.22 & -3.15 \\
-        DNAJA2 & DnaJ (Hsp40) homolog, subfamily A, member 1 & 2.34 & -6.02 \\
-        RAD17 & RAD17 homolog (S. pombe) & 1.45 & -1.80 \\
-        PSMG3 & Proteosome assembly chaperone 3 & 1.52 & -1.26 \\
-        PNISR & PNN-interacting serine/argining-rich protein & 5.58 & -1.81 \\
-        THOC7 & THO complex 7 homolog (Drosophila) & 7.00 & -6.90 \\
-        YTHDC2 & YTH domain containing 2 & 1.00 & -1.74 \\
-        RPL39 & Ribosomal protein L39 & 3.31 & -1.59 \\
-        CD7 & CD7 molecule & 2.88 & -1.38 \\
-        NDUFB3 & NADH dehydrogenase (uqiquinone) 1 beta subcomplex 3 & 1.03 & -1.34 \\
-        LOC100858785 & Unknown & 1.26 & -1.75 \\
-        Unknown & Unknown & 6.30 & -5.22 \\
-        \hline
-    \end{tabular}
-    \begin{flushleft}
-        (-) down-regulated, (+) up-regulated
-    \end{flushleft}
-    \label{tab:opposite}
-\end{table}
-
-\begin{table}[!ht]
-\caption{
-\bf{Cytokine-related gene expression in response to MDV infection}}
-    \begin{tabular}{cccccc}
-        \hline
-        & & $log_{2}$FC & \\
-        Symbol & Description & Line 6 & Line 7 \\
-        \hline
-        IL2RG & Interleukin 2 receptor, $\gamma$ & -- & 0.55 \\
-        IL6 & Interleukin 6 (interferon, $\beta$ 2) & -- & 5.15 \\
-        IL6ST & Interleukin 6 signal transducer (gp130, oncostatin M receptor) & -- & 1.36 \\
-        IL8L1 & Interleukin 8-like 1 & 1.90 & -- \\
-        IL15 & Interleukin 15 & -- & 1.06 \\
-        IL18 & Interleukin 18 (interferon-$\gamma$ inducing factor) & 1.92 & 4.06 \\
-        IL18R1 & Interleukin 18 receptor 1 & 1.94 & 1.64 \\
-        IFNG & Interferon-$\gamma$ & 5.14 & 4.90 \\
-        IFNB & Interferon-$\beta$ & 4.83 & 5.64 \\
-        IFNA3 & Interferon-$\alpha$ 3 & 4.09 & 5.48 \\
-        IFNGR1 & Interferon-$\gamma$ receptor 1 & -- & 2.05 \\
-        IFNGR2 & Interferon-$\gamma$ receptor 2& -- & 0.50 \\
-        IFNAR1 & Interferon-$\alpha$,$\beta$ receptor 1 & -- & 1.46 \\
-        IFNAR2 & Interferon-$\alpha$,$\beta$ receptor 2 & -- & 0.58 \\
-        \hline
-    \end{tabular}
-    \begin{flushleft}
-    \end{flushleft}
-    \label{tab:cytokines}
-\end{table}
-
-\begin{table}[!ht]
-\caption{
-\bf{DEU between Line 6 and Line 7 in infected birds group I}}
+\bf{DEU between the resistant line and the susceptible line in infected birds group I}}
 \begin{tabular}{cccccccc}
 \hline
-& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
+& & & & Resistant ($\Psi$) & & Susceptible ($\Psi$) & \\
 Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
@@
     *human homologs, Un=uninfected, Inf=infected.
-    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
+    Bold face indicates that there is a SNP between the resistant line and 7 within an alternative exon.
 \end{flushleft}
@@
 \caption{
-\bf{DEU between Line 6 and Line 7 in infected birds group II}}
+\bf{DEU between the resistant line and the susceptible line in infected birds group II}}
 \begin{tabular}{cccccccc}
 \hline
-& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
+& & & & Resistant ($\Psi$) & & Susceptible ($\Psi$) & \\
 Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
@@
 \caption{
-\bf{DEU between Line 6 and Line 7 in infected birds group III and IV}}
-\begin{tabular}{cccccccc}
-\hline
-& & & & Line 6 ($\Psi$) & & Line 7 ($\Psi$) & \\
+\bf{DEU between the resistant line and the susceptible line in infected birds group III and IV}}
+\begin{tabular}{cccp{2cm}cccc}
+\hline
+& & & & Resistant ($\Psi$) & & Susceptible ($\Psi$) & \\
 Type & Event ID & Ensembl & Symbol  & Un & Inf & Un & Inf \\
@@
     *human homologs, Un=uninfected, Inf=infected.
-    Bold face indicates that there is a SNP between line 6 and 7 within an alternative exon.
+    Bold face indicates that there is a SNP between the resistant
+    line and 7 within an alternative exon.
 \end{flushleft}
@@
 \hline
-Gene &  Chromosome & Position & Reference & Line 6 & Line 7 & Strand \\
+Gene &  Chromosome & Position & Reference & Resistant & Susceptible & Strand \\
 \hline
@@
 \bf{Results from human splicing finder}}
-\begin{tabular}{|l|p{1cm}|p{3cm}|l|l|l|l|}
+\begin{tabular}{lp{1cm}p{3cm}llll}
 \hline
